### 2025-04-05 doc/proposal/proposal.tex
Proposal

\documentclass[a4paper,12pt]{article}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{fancyhdr}

\title{D.I.A.G.R.A.M.: Development of Image Analysis for Graph Recognition And Modeling}
\author{Napolitano Saverio, Ricciardi Nicola, Garagnani Filippo}
\date{\today}

\pagestyle{fancy}
\fancyhead[L]{D.I.A.G.R.A.M. Proposal}
\fancyhead[R]{Napolitano, Ricciardi, Garagnani}

\begin{document}

\maketitle

\begin{abstract}
    This project proposes the development of a system for analyzing different types of handwritten diagrams and converting them into well-rendered images through a textual syntax. The goal is to create a tool capable of analyzing scanned or photographed sketches of diagrams and automatically generating code that can be rendered into the same diagrams digitally, to eventually integrate or modify them.
\end{abstract}

\section{Introduction}
    Diagrams are a powerful means of representing complex systems, ideas, and relationships. Writing diagrams in code format is extremely efficient to easily edit the diagram and to display it in a better-formed format. This project aims to automate the conversion of handwritten diagrams into code, streamlining the process of diagram creation and visualization.

\section{Objective}
    The main objective of the D.I.A.G.R.A.M. system is to:
    \begin{itemize}
        \item Develop a computer vision model to recognize handwritten diagrams.
        \item Extract relevant features from the diagrams (nodes, edges, labels).
        \item Generate equivalent code in supported languages for the recognized diagrams.
        \item Automatically parse the code to generate the formatted image of the input diagram.
    \end{itemize}

\section{System Architecture}
    The project will be composed of different modules that will be stacked together in an uniform pipeline [see Figure 1]. These components are:
    \begin{itemize}
    		\item \textbf{Preprocessor}: Applies some operations (like denoising, rotation...) in order to obtain a more suitable input image for the following modules.
    		\item \textbf{Classifier}: In order to manage more than one type of diagrams, a classifier that identifies the category of the input diagram is needed.
    		\item \textbf{Extractor}: By employing feature extraction, image segmentation and static analysis over the recognized components, the extractor will output a robust structure to contain all the relevant information about the diagram. This key component takes the image as input and will output the entities found, their relations and more useful information.
    		\item \textbf{Transducer}: From the data output by the extractor, a certain transducer will give out a formatted code relative to the input diagram. We envision a single transducer for every supported language.
    		\item \textbf{Compiler}: Given the code, the compiler will convert it into an output image - hopefully, as close as possible as the input diagram.
    \end{itemize}

\begin{figure}[h]
    \centering
    \includegraphics[width=1.2\textwidth]{../assets/images/overview.png} % Change the path to your image
    \caption{The D.I.A.G.R.A.M. pipeline.}
\end{figure}

\end{document}

### 2025-04-07 src/transducer/d2/flowchart_transducer.py
Refactor

@@ -1,12 +1,14 @@
 from typing import List, Type
-from py_d2 import D2Diagram, D2Shape, D2Connection, D2Style
 
 from src.wellknown_diagram import WellKnownDiagram
 from core.representation.representation import DiagramRepresentation
 from src.representation.flowchart_representation.flowchart_representation import FlowchartRepresentation
+from src.representation.flowchart_representation.element import FlowchartElementCategory, Element
+from src.representation.flowchart_representation.relation import FlowchartRelationCategory, Relation
 from core.transducer.outcome import TransducerOutcome
 from core.transducer.transducer import Transducer
 from src.wellknown_markuplang import WellKnownMarkupLanguage
+
 
 class FlowchartToD2Transducer(Transducer):
 
@@ -22,26 +24,67 @@ class FlowchartToD2Transducer(Transducer):
     def compatible_representations(self) -> List[Type[DiagramRepresentation]]:
         return [FlowchartRepresentation]
 
+    @staticmethod
+    def wrap_element(category: str, label: str, elem_id: str) -> str:
+        match category:
+            case FlowchartElementCategory.CIRCLE.value:
+                return f"{elem_id}: {label}\n" \
+                       f"{elem_id}.shape: circle\n"
+            case FlowchartElementCategory.TERMINAL.value:
+                return f"{elem_id}: {label}\n" \
+                       f"{elem_id}.shape: oval\n"
+            case FlowchartElementCategory.PROCESS.value:
+                return f"{elem_id}: {label}\n" \
+                       f"{elem_id}.style.border-radius: 8\n"
+            case FlowchartElementCategory.DECISION.value:
+                return f"{elem_id}: {label}\n" \
+                       f"{elem_id}.shape: diamond\n"
+            case FlowchartElementCategory.INPUT_OUTPUT.value:
+                return f"{elem_id}: {label}\n" \
+                       f"{elem_id}.shape: parallelogram\n"
+            case FlowchartElementCategory.SUBROUTINE.value:
+                return f"{elem_id}: {label}\n"
+            case _:
+                raise ValueError(f"Unknown flowchart element category: {category}")
+
+    @staticmethod
+    def wrap_relation(category: str, label: str, target_id: str) -> str:
+        match category:
+            case FlowchartRelationCategory.ARROW.value:
+                if label == "":
+                    return f"->{target_id}\n"
+                return f"->{target_id}: {label}\n"
+            case FlowchartRelationCategory.OPEN_LINK.value:
+                if label == "":
+                    return f"--{target_id}\n"
[…]

### 2025-04-07 src/transducer/mermaid/flowchart_transducer.py

@@ -1,13 +1,15 @@
 from typing import List, Type
-from python_mermaid.diagram import MermaidDiagram, Link, Node
-from python_mermaid.utils import snake_case
+import sys, os
+
+from src.representation.flowchart_representation.element import FlowchartElementCategory
+from src.representation.flowchart_representation.relation import FlowchartRelationCategory
 
 from src.wellknown_diagram import WellKnownDiagram
 from core.representation.representation import DiagramRepresentation
 from src.representation.flowchart_representation.flowchart_representation import FlowchartRepresentation
 from core.transducer.outcome import TransducerOutcome
 from core.transducer.transducer import Transducer
-from src.wellknown_markuplang import WellKnownMarkupLanguage
+
 
 class FlowchartToMermaidTransducer(Transducer):
 
@@ -26,33 +28,53 @@ class FlowchartToMermaidTransducer(Transducer):
 
         ]
 
+    @staticmethod
+    def wrap_element(category: str, label: str) -> str:
+        match category:
+            case FlowchartElementCategory.CIRCLE.value:
+                return f"(({label}))"
+            case FlowchartElementCategory.TERMINAL.value:
+                return f"([{label}])"
+            case FlowchartElementCategory.PROCESS.value:
+                return f"({label})"
+            case FlowchartElementCategory.DECISION.value:
+                return "{" + label + "}"
+            case FlowchartElementCategory.INPUT_OUTPUT.value:
+                return f"[/{label}/]"
+            case FlowchartElementCategory.SUBROUTINE.value:
+                return f"[{label}]"
+            case _:
+                raise ValueError(f"Unknown flowchart element category: {category}")
+
+    @staticmethod
+    def wrap_relation(category: str, label: str) -> str:
+        match category:
+            case FlowchartRelationCategory.ARROW.value:
+                if label == "":
+                    return "-->"
+                return f"-->|{label}|"
+            case FlowchartRelationCategory.OPEN_LINK.value:
+                if label == "":
+                    return " --- "
+                return f"---|{label}|"
+            case FlowchartRelationCategory.DOTTED_ARROW.value:
+                if label == "":
+                    return "-.->"
+                return f" -. {label} .->"
+            case _:
+                raise ValueError(f"Unknown flowchart relation category: {category}")
+
     def transduce(self, diagram_id: str, diagram_representation: DiagramRepresentation) -> TransducerOutcome:
         assert isinstance(diagram_representation, FlowchartRepresentation)
-
-        nodes: list[Node] = [
-            Node(
[…]

### 2025-04-07 doc/proposal/proposal.tex
Proposal edit

@@
 
+\renewcommand{\abstractname}{Summary}
+
 \title{D.I.A.G.R.A.M.: Development of Image Analysis for Graph Recognition And Modeling}
-\author{Napolitano Saverio, Ricciardi Nicola, Garagnani Filippo}
+\author{Filippo Garagnani, Saverio Napolitano, Nicola Ricciardi}
 \date{\today}
@@
 \pagestyle{fancy}
-\fancyhead[L]{D.I.A.G.R.A.M. Proposal}
-\fancyhead[R]{Napolitano, Ricciardi, Garagnani}
+\fancyhead[L]{D.I.A.G.R.A.M. - Project Proposal}
+\fancyhead[R]{\small Garagnani, Napolitano, Ricciardi}
 
@@
     Diagrams are a powerful means of representing complex systems, ideas, and relationships.
-Writing diagrams in code format is extremely efficient to easily edit the diagram and to display it in a better-formed format.
+Writing diagrams in a textual format is extremely efficient to easily edit the diagram and to display it in a better-formed format.
 This project aims to automate the conversion of handwritten diagrams into code, streamlining the process of diagram creation and visualization.
@@
 \section{System Architecture}
-    The project will be composed of different modules that will be stacked together in an uniform pipeline [see Figure 1]. These components are:
+    The project will be composed of different modules that will be stacked together in an uniform pipeline [\textbf{Figure 1}]. These components are:
     \begin{itemize}
@@
 
+\newpage
+
+\section{Constraint}
+While the whole project idea could be applied to a wide variety of diagrams and languages, in order to manage the complexity we decided to impose a few theoretical constraints:
+
+	\begin{enumerate}
+
+	\item \textbf{Types of diagrams}: We restrict the types of diagrams that the system is able to correctly recognize and reconstruct (\textit{e.g.} Graph diagrams, Flowchart diagrams...)
+	\item \textbf{Languages}: Even if there are various languages in which the diagrams could be formatted, we limit the languages that the system could translate the diagrams to (\textit{e.g.} Mermaid, D2...)
+
+	\end{enumerate}
+
 \end{document}

### 2025-04-07 doc/proposal/proposal.tex
Proposal edit 2

@@
         \item Develop a computer vision model to recognize handwritten diagrams.
-        \item Extract relevant features from the diagrams (nodes, edges, labels).
+        \item Extract relevant features from the diagrams (\textit{e.g.} nodes, edges, labels).
         \item Generate equivalent code in supported languages for the recognized diagrams.
@@
 
-\section{Constraint}
+\section{Constraints}
 While the whole project idea could be applied to a wide variety of diagrams and languages, in order to manage the complexity we decided to impose a few theoretical constraints:

### 2025-04-07 src/representation/flowchart_representation/element.py
Revert element.py

@@ -6,12 +6,12 @@ For reference: https://en.wikipedia.org/wiki/Flowchart#Building_blocks
 """
 
 class FlowchartElementCategory(Enum):
-    PROCESS: str = "normal"
-    TERMINAL: str = "round-edge"
-    DECISION: str = "rhombus"
-    INPUT_OUTPUT: str = "parallelogram"
-    CIRCLE: str = "circle"
-    SUBROUTINE: str = "subroutine-shape"
+    TERMINAL: str = "TerminalNode"
+    PROCESS: str = "ProcessNode"
+    DECISION: str = "DecisionNode"
+    INPUT_OUTPUT: str = "InputOutputNode"
+    CIRCLE: str = "CircleNode"
+    SUBROUTINE: str = "SubroutineNode"
 
 
 @dataclass(frozen=True, slots=True)

### 2025-04-07 src/representation/flowchart_representation/relation.py
Revert relation.py

@@ -7,8 +7,9 @@ For reference: https://en.wikipedia.org/wiki/Flowchart#Building_blocks
 
 
 class FlowchartRelationCategory(Enum):
-    ARROW: str = "normal"
-    DOTTED_ARROW: str = "dotted"
+    ARROW: str = "Arrowhead"
+    OPEN_LINK: str = "OpenLink"
+    DOTTED_ARROW: str = "DottedArrowhead"
 
 
 @dataclass(frozen=True, slots=True)

### 2025-04-07 src/transducer/mermaid/flowchart_transducer.py
revert flowchart_transducer.py

@@ -11,6 +11,10 @@ from core.transducer.outcome import TransducerOutcome
 from core.transducer.transducer import Transducer
 
 
+class FlowchartRelation:
+    pass
+
+
 class FlowchartToMermaidTransducer(Transducer):
 
     def __init__(self, identifier: str):
@@ -66,15 +70,15 @@ class FlowchartToMermaidTransducer(Transducer):
 
     def transduce(self, diagram_id: str, diagram_representation: DiagramRepresentation) -> TransducerOutcome:
         assert isinstance(diagram_representation, FlowchartRepresentation)
-        
+
         body: str = "Flowchart TD\n"
         for id, element in diagram_representation.elements.items():
             body += f"\t{id}{self.wrap_element(element.category, element.label)}\n"
-            
+
         body += "\n"
         for relation in diagram_representation.relations:
             body += f"\t{relation.source_id}"
             body += f"{self.wrap_relation(relation.category, relation.label)}{relation.target_id}\n"
-        
+
         outcome: TransducerOutcome = TransducerOutcome(diagram_id=diagram_id, payload=body, markup_language="mermaid")
         return outcome

### 2025-04-07 src/transducer/mermaid/flowchart_transducer.py
Fix tests

@@ -71,9 +71,9 @@ class FlowchartToMermaidTransducer(Transducer):
     def transduce(self, diagram_id: str, diagram_representation: DiagramRepresentation) -> TransducerOutcome:
         assert isinstance(diagram_representation, FlowchartRepresentation)
 
-        body: str = "Flowchart TD\n"
-        for id, element in diagram_representation.elements.items():
-            body += f"\t{id}{self.wrap_element(element.category, element.label)}\n"
+        body: str = "flowchart TD\n"
+        for identifier, element in diagram_representation.elements.items():
+            body += f"\t{identifier}{self.wrap_element(element.category, element.label)}\n"
 
         body += "\n"
         for relation in diagram_representation.relations:

### 2025-04-07 src/transducer/mermaid/test.py

@@ -42,15 +42,14 @@ class TestFlowchartToMermaidTransducer(unittest.TestCase):
 
         representation = FlowchartRepresentation(elements, relations)
         expected_outcome = TransducerOutcome(diagram_id="test_diagram", markup_language=WellKnownMarkupLanguage.MERMAID.value,
-                                             payload="""---
-title: test_diagram
----
-graph 
-a(("Start_Node"))
-b["i++"]
-c{"if i > 5")
-a --->|int i = 0| b
-b -.-> c""")
+                                             payload="""flowchart TD
+	A((Start_Node))
+	B(i++)
+	C{if i > 5}
+
+	A-->|int i = 0|B
+	B-.->C
+""")
 
         assert expected_outcome == self.transducer.transduce("test_diagram", representation), \
             "Elaborated outcome should match expected outcome"

### 2025-04-07 src/representation/flowchart_representation/flowchart_representation.py
revert flowchart_transducer.py

@@ -16,7 +16,7 @@ class FlowchartRepresentation(DiagramRepresentation):
 
     """
         The syntax is: first all elements - with id, category and label separated by ;; -
-        then a new line and all relations - with category, source_id, target_id and label separeted by ;;
+        then a new line and all relations - with category, source_id, target_id and label separated by ;;
     """
     def dump(self, output_path: str):
         

### 2025-04-07 src/transducer/d2/flowchart_transducer.py

@@ -9,7 +9,6 @@ from core.transducer.outcome import TransducerOutcome
 from core.transducer.transducer import Transducer
 from src.wellknown_markuplang import WellKnownMarkupLanguage
 
-
 class FlowchartToD2Transducer(Transducer):
 
     def __init__(self, identifier: str):
@@ -87,4 +86,3 @@ class FlowchartToD2Transducer(Transducer):
 
         outcome: TransducerOutcome = TransducerOutcome(diagram_id, WellKnownMarkupLanguage.D2_LANG.value, body)
         return outcome
-

### 2025-05-03 doc/article/article.tex
Article draft

\documentclass[conference]{IEEEtran}

\usepackage{graphicx}
\usepackage{amsmath, amssymb}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{cite}
\usepackage{float}

\title{D.I.A.G.R.A.M.: Development of Image Analysis for Graph Recognition And Modeling}

\author{
    \IEEEauthorblockN{Filippo Garagnani, Saverio Napolitano, Nicola Ricciardi}
    \IEEEauthorblockA{
        'Computer Vision and Cognitive System' course \\
        \textit{Università di Modena e Reggio Emilia}
    }
}

\begin{document}

\maketitle

\begin{abstract}
This report describes D.I.A.G.R.A.M., a system for diagram recognition and generation. We present its architecture, the logic behind deterministic algorithms and the training of its deep learning components. The project aims to transform visual diagram input into structured representations.
\end{abstract}

\section{Introduction}
Diagrams are crucial in education, documentation, and many other fields. Developing a system that automatically understands and generates diagrams may reveal helpful for having high-quality, easily editable representations. This, however, poses challenges involving classification, shape detection, structure interpretation and symbolic representation. Our project proposes a modular architecture to tackle these tasks.

\section{System Architecture}
Figure~\ref{fig:pipeline} illustrates the full pipeline, composed of four main modules: the Classifier, Extractor, Transducer, and Compiler. Each module is designed to process and transform the handwritten diagram image progressively toward a structured output.

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{example-image}
\caption{Overview of the D.I.A.G.R.A.M. system architecture.}
\label{fig:pipeline}
\end{figure}

The Classifier is able to recognize the category of the hand-written diagram that the user submitted as input. This is necessary in order for the system to know to which Extractor pass the data. This module is able to apply object detection and semantic recognition to represent the diagram in a unified way. After that, the representation is sent to the Transducer, which converts the data in a Markup language of choice. The Markup language content is therefore sent to the Compiler, thus generating the high-quality and editable diagram in .png format.

% Classifier
\section{Classifier Module}
TODO
\subsection{Preprocessing}
TODO

\subsection{Model}
TODO

% Extractor
\section{Extractor Module}
TODO
\subsection{Preprocessing}
TODO
\subsection{Bounding Box Detection}
TODO
\subsection{Content Recognition}
TODO

% Transducer
\section{Transducer Module}
TODO

% Compiler
\section{Compiler Module}
TODO

\section{Experiments}
\subsection{Dataset}
TODO

\subsection{Metrics}
TODO

\subsection{Results}
TODO

\section{Discussion}
TODO

\section{Conclusion and Future Work}
TODO

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

### 2025-05-03 doc/article/article.tex
Draft

@@
 \section{Extractor Module}
-TODO
+The Extractor is the key component of the D.I.A.G.R.A.M. system, able to transform the raw diagram image into a structured, category-agnostic representation.
+The Extractor recognizes the diagram's components, such as shapes, lines, and text, through an object detection network, and organizes them into a unified format.
+This representation serves as the foundation for the subsequent Transducer module, which converts the structured data into a domain-specific markup language.
+
 \subsection{Preprocessing}
@@
 \section{Transducer Module}
-TODO
+The Transducer is responsible for converting the unified, agnostic representation of a diagram into a domain-specific markup language. 
+This transformation enables the subsequent Compiler module to generate high-quality visual outputs.
 
@@
 \section{Compiler Module}
-TODO
+The Compiler module is the final stage of the D.I.A.G.R.A.M. pipeline. 
+Its primary role is to take the structured representation of the diagram, expressed in a markup language (e.g., Mermaid.js), and generate a high-quality, editable diagram in a visual format such as PNG.
 

### 2025-05-05 doc/article/article.tex
Draft

@@
 % Classifier
-\section{Classifier Module}
+\subsection{Classifier Module}
 TODO
-\subsection{Preprocessing}
-TODO
-
-\subsection{Model}
-TODO
-
-% Extractor
-\section{Extractor Module}
+\subsection{Extractor Module}
 The Extractor is the key component of the D.I.A.G.R.A.M. system, able to transform the raw diagram image into a structured, category-agnostic representation.
@@
 This representation serves as the foundation for the subsequent Transducer module, which converts the structured data into a domain-specific markup language.
+\subsection{Transducer Module}
+The Transducer is responsible for converting the unified, agnostic representation of a diagram into a domain-specific markup language. 
+This transformation enables the subsequent Compiler module to generate high-quality visual outputs.
+\subsection{Compiler Module}
+The Compiler module is the final stage of the D.I.A.G.R.A.M. pipeline. 
+Its primary role is to take the structured representation of the diagram, expressed in a markup language (e.g., Mermaid.js), and generate a high-quality, editable diagram in a visual format such as PNG.
 
-\subsection{Preprocessing}
+\section{Graph Diagram Recognition}
+
+\subsection{Classifier Module}
+\subsubsection{Preprocessing}
 TODO
-\subsection{Bounding Box Detection}
-TODO
-\subsection{Content Recognition}
+\subsubsection{Model}
 TODO
 
-% Transducer
-\section{Transducer Module}
-The Transducer is responsible for converting the unified, agnostic representation of a diagram into a domain-specific markup language. 
-This transformation enables the subsequent Compiler module to generate high-quality visual outputs.
+\subsection{Extractor Module}
+\subsubsection{Preprocessing}
+TODO
+\subsubsection{Bounding Box Detection}
+TODO
+\subsubsection{Content Recognition}
 
-% Compiler
-\section{Compiler Module}
-The Compiler module is the final stage of the D.I.A.G.R.A.M. pipeline. 
-Its primary role is to take the structured representation of the diagram, expressed in a markup language (e.g., Mermaid.js), and generate a high-quality, editable diagram in a visual format such as PNG.
+\subsection{Transducer Module}
+\subsection{Compiler Module}
 

### 2025-05-05 doc/article/article.tex
Draft

@@
 
+% System Architecture
 \section{System Architecture}
@@
 \centering
-\includegraphics[width=\linewidth]{example-image}
+\includegraphics[width=\linewidth]{overview.png}
 \caption{Overview of the D.I.A.G.R.A.M. system architecture.}
@@
 
-% Classifier
 \subsection{Classifier Module}
-TODO
+The Classifier is the first component of the pipeline.
+It is able to tell to which category a given image of an handwritten diagram belongs to.
+This is necessary in order to later know the Extractors that can process the input image.
 \subsection{Extractor Module}
@@
 The Compiler module is the final stage of the D.I.A.G.R.A.M. pipeline. 
-Its primary role is to take the structured representation of the diagram, expressed in a markup language (e.g., Mermaid.js), and generate a high-quality, editable diagram in a visual format such as PNG.
+Its primary role is to take the structured representation of the diagram, expressed in a markup language (e.g., Mermaid.js), and generate a high-quality diagram in a visual format such as PNG.
 

### 2025-05-08 doc/article/article.tex
Draft update

@@
 
+% 
+% Graph Diagram
+%
 \section{Graph Diagram Recognition}
+
+\subsection{Internal Representation}
+In order to pass data through the components of the pipeline in a unified way, it was necessary to decide upon an internal representation of the graph diagram.
+It consists of \textbf{elements} and \textbf{relations}.
+The latter are linked to at least one element.
+Every element has a \textit{category} and possibly some \textit{inner} and \textit{outer text}.
+Every relation has a \textit{category} too, either a \textit{source} or \textit{target element} or both, and possibly \textit{text} related to it.
+
+This internal representation is generated from the image provided by the user by the Extractor module, while it is parsed and converted to a markup language of choice by the Transducer.
 
@@
 \subsubsection{Preprocessing}
-TODO
+TODO \\
+
 \subsubsection{Model}
-TODO
+TODO \\
 
@@
 \subsubsection{Preprocessing}
-TODO
+TODO\\
 \subsubsection{Bounding Box Detection}
-TODO
+TODO\\
 \subsubsection{Content Recognition}
+After having determined all the bounding boxes of the diagram image, in order to correctly create the diagram internal representation, there's the need to link together text, elements and relations.\\
 
-\subsection{Transducer Module}
-\subsection{Compiler Module}
+\subsubsection{Text Digitization}
+Since the task of converting an image of text into the relative string has already been solved in various ways, we settled upon the Microsoft \textit{Trocr Handwritten} Transformer model; more precisely, the \textit{small} version of it.
+[TODO: Fare esperimenti e provarne altri].
+Every bounding box containing text that has been found is later passed to the model, and the resulting string generated by the model is linked to it.\\
+
+\subsubsection{Text Association}
+It can be assumed that every piece of text is linked to one and only one element or relation.
+So, every bounding box containing text is linked to the nearest element - it being either a relation or an element.
+The associations that require too long a distance are eliminated further in the algorithm, and the relevant text is discarded.\\
+
+\subsubsection{Element-Text Association}
+After that a text box has been assigned to an element, there's the need to understand if the text is either inside or outside the element and whether the text is actually referring to the given element or not.
+In order to do this, the overlap between the two bounding boxes is computed; if the overlap is over a certain threshold, heuristically decided, the association is kept.
+Otherwise, the distance between the two bounding boxes is computed - as the distance between their central point.
+If the distance is over a certain threshold, the association is kept; otherwise, it's discarded.\\
+
+\subsubsection{Relation-Text Association}
+After a text box is assigned to a relation, some computation is needed to fully understand how the text is associated to the relation itself.
+[TODO: Come trova source e target].
+The relation bounding box is then split into three different sections.
+The overlap between the text bounding box and each of the three sections boxes is computed - then, the section with the highest overlap value is then assigned the text under consideration.
 

### 2025-05-11 src/extractor/text_extraction/test.py
Text Extractor Test

@@ -3,7 +3,7 @@ import sys, os
 sys.path.append(os.path.dirname("."))
 sys.path.append(os.path.dirname("/"))
 
-from src.extractor.text_extraction.text_extractor import TrOCRTextExtractorSmall, TrOCRTextExtractorBase, TrOCRTextExtractorLarge
+from src.extractor.text_extraction.text_extractor import TrOCRTextExtractorSmall, TrOCRTextExtractorBase, TrOCRTextExtractionSmallHandwritten
 from src.extractor.text_extraction.text_extractor import TextExtractor
 from core.image.tensor_image import TensorImage, Image
 from core.image.bbox.bbox2p import ImageBoundingBox2Points, ImageBoundingBox
@@ -11,36 +11,49 @@ import os, json, torch
 
 if __name__ == "__main__":
     # Create instances of the text extractors
-    small_extractor = TrOCRTextExtractorSmall()
+    # small_extractor = TrOCRTextExtractorSmall()
     # base_extractor = TrOCRTextExtractorBase()
-    # large_extractor = TrOCRTextExtractorLarge()
+    small_handwritten = TrOCRTextExtractionSmallHandwritten()
     
     json_path = os.path.join(os.path.dirname(__file__), "../../../dataset/source/fa/test.json")
     with open(json_path, 'r') as json_file:
         json_content = json.load(json_file)
     
-    images: list[Image] = []
-    text_bboxes: list[ImageBoundingBox] = []
-    ground_truth_texts: list[str] = []
+    # small_metrics: list[dict] = []
+    base_metrics: list[dict] = []
+    # large_metrics: list[dict] = []
     dataset_path = os.path.join(os.path.dirname(__file__), "../../../dataset/source/fa/test/")
     for root, dirs, files in os.walk(dataset_path):
-        for file in files:
+        for i, file in enumerate(files):
+            print("Processing file: " + str(i) + "/" + str(len(files)))
             file_path = os.path.join(root, file)
             image_id = next((item for item in json_content["images"] if item["file_name"] == file), None)["id"]
-            annotations = [annotation for annotation in json_content["annotations"] if annotation["image_id"] == image_id and annotation["category"] == "text"]
+            annotations = [annotation for annotation in json_content["annotations"] if annotation["image_id"] == image_id and annotation["category"] == "text" and annotation["text"] != None]
+            image = TensorImage.from_str(file_path)
+            partial_metric = []
             for annotation in annotations:
[…]

### 2025-05-11 src/extractor/text_extraction/text_extractor.py

@@ -24,7 +24,7 @@ class TextExtractor(ABC):
         """
         raise NotImplementedError("Subclasses should implement this method.")
     
-    def compute_metric(self, image: Image, text_bbox: ImageBoundingBox, ground_truth: str, metric: str) -> float:
+    def compute_metric(self, extracted_text: str, ground_truth: str, metric: str) -> float:
         """
         Computes a metric based on the extracted text and ground truth.
 
@@ -36,7 +36,6 @@ class TextExtractor(ABC):
         Returns:
             float: The computed metric value.
         """
-        extracted_text = self.extract_text(image, text_bbox)
 
         if metric == "levenshtein":
             return self.levenshtein_distance(extracted_text, ground_truth)
@@ -47,6 +46,25 @@ class TextExtractor(ABC):
                 else:
                     extracted_text += " " * (len(ground_truth) - len(extracted_text))
             return self.hamming_distance(extracted_text, ground_truth)
+        elif metric == "cosine":
+            similarity = torch.nn.CosineSimilarity(dim=1, eps=1e-6)
+            extracted_embedding = self.compute_embedding(extracted_text).float()
+            ground_truth_embedding = self.compute_embedding(ground_truth).float()
+            if extracted_embedding.shape != ground_truth_embedding.shape:
+                if extracted_embedding.shape[1] > ground_truth_embedding.shape[1]:
+                    ground_truth_embedding = torch.cat([ground_truth_embedding, torch.zeros(1, extracted_embedding.shape[1] - ground_truth_embedding.shape[1])], dim=1)
+                else:
+                    extracted_embedding = torch.cat([extracted_embedding, torch.zeros(1, ground_truth_embedding.shape[1] - extracted_embedding.shape[1])], dim=1)
+            return 1 - similarity(extracted_embedding, ground_truth_embedding).item()
+        elif metric == "euclidean":
+            extracted_embedding = self.compute_embedding(extracted_text).float()
+            ground_truth_embedding = self.compute_embedding(ground_truth).float()
+            if extracted_embedding.shape != ground_truth_embedding.shape:
+                if extracted_embedding.shape[1] > ground_truth_embedding.shape[1]:
+                    ground_truth_embedding = torch.cat([ground_truth_embedding, torch.zeros(1, extracted_embedding.shape[1] - ground_truth_embedding.shape[1])], dim=1)
+                else:
[…]

### 2025-05-11 core/image/tensor_image.py
Add dataset extractor

@@ -1,8 +1,9 @@
-from torchvision.io import decode_image, read_image
+from torchvision.io import read_image
 
 from core.image.image import Image
 from dataclasses import dataclass
 from torch import Tensor
+
 
 @dataclass
 class TensorImage(Image):

### 2025-05-11 src/dataset/classifier/create_annotations.py

@@ -1,0 +1,100 @@
+from os import listdir
+from typing import List, Dict
+import json
+
+
+def annotate(images: List[str], annotation: str) -> Dict[str, str]:
+    json_annotations: Dict[str, str] = {}
+    for image in images:
+        json_annotations[image] = annotation
+    return json_annotations
+
+
+if __name__ == "__main__":
+    fa_val_path = "/dataset/source/fa/val"
+    fa_test_path = "/dataset/source/fa/test"
+    fa_train_path = "/dataset/source/fa/train"
+
+    fca_test_path = "/dataset/source/fca/test"
+    fca_train_path = "/dataset/source/fca/train"
+
+    fcb_val_path = "/dataset/source/fcb/val"
+    fcb_test_path = "/dataset/source/fcb/test"
+    fcb_train_path = "/dataset/source/fcb/train"
+
+    circuit_path = "/dataset/source/circuit"
+
+    class_path = "/dataset/source/class"
+
+    school_path = "/dataset/source/school"
+
+    bpmn_val_path = "/dataset/source/hdBPMN-icdar2021/val"
+    bpmn_test_path = "/dataset/source/hdBPMN-icdar2021/test"
+    bpmn_train_path = "/dataset/source/hdBPMN-icdar2021/train"
+
+    fa_val_file_names = [f for f in listdir(fa_val_path)]
+    fa_test_file_names = [f for f in listdir(fa_test_path)]
+    fa_train_file_names = [f for f in listdir(fa_train_path)]
+
+    fa_json = annotate(fa_val_file_names, 'graph')
+    fa_test_json = annotate(fa_test_file_names, 'graph')
+    fa_train_json = annotate(fa_train_file_names, 'graph')
+
+    fa_json.update(fa_test_json)
+    fa_json.update(fa_train_json)
+
+    fca_test_file_names = [f for f in listdir(fca_test_path)]
+    fca_train_file_names = [f for f in listdir(fca_train_path)]
+
+    fca_json = annotate(fca_test_file_names, 'flowchart')
+    fca_train_json = annotate(fca_train_file_names, 'flowchart')
+
+    fca_json.update(fca_train_json)
+
+    fcb_val_file_names = [f for f in listdir(fcb_val_path)]
+    fcb_test_file_names = [f for f in listdir(fcb_test_path)]
+    fcb_train_file_names = [f for f in listdir(fcb_train_path)]
+
[…]

### 2025-05-11 src/dataset/classifier/test.py

@@ -6,20 +6,20 @@ from src.dataset.classifier.dataset import ClassifierDataset
 import cv2
 
 
-class TestGNRDataset(unittest.TestCase):
+class TestClassifierDataset(unittest.TestCase):
 
     def setUp(self):
         self.dataset = ClassifierDataset("/Users/saverionapolitano/PycharmProjects/diagram/dataset/classifier/labels"
                                          ".json",
-                                         "/dataset/classifier/all")
+                                         "/Users/saverionapolitano/PycharmProjects/diagram/dataset/classifier/all")
 
     def test_len(self):
-        self.assertEqual(self.dataset.__len__(), 1391)
+        self.assertEqual(self.dataset.__len__(), 9216)
 
     def test_get_item(self):
         image, label = self.dataset.__getitem__(0)
         im = cv2.imread(
-            '/dataset/classifier/all/writer013_fa_007.png', cv2.IMREAD_UNCHANGED)
+            '/Users/saverionapolitano/PycharmProjects/diagram/dataset/classifier/all/writer013_fa_007.png', cv2.IMREAD_UNCHANGED)
         tensor_im = torch.from_numpy(im)
         self.assertTrue(torch.equal(tensor_im[torch.newaxis, :, :], image.as_tensor()))
         self.assertEqual(label, 'graph')

### 2025-05-11 src/dataset/create_annotations.py

@@ -1,100 +1,0 @@
-from os import listdir
-from typing import List, Dict
-import json
-
-
-def annotate(images: List[str], annotation: str) -> Dict[str, str]:
-    json_annotations: Dict[str, str] = {}
-    for image in images:
-        json_annotations[image] = annotation
-    return json_annotations
-
-
-if __name__ == "__main__":
-    fa_val_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fa/val"
-    fa_test_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fa/test"
-    fa_train_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fa/train"
-
-    fca_test_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fca/test"
-    fca_train_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fca/train"
-
-    fcb_val_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fcb/val"
-    fcb_test_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fcb/test"
-    fcb_train_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fcb/train"
-
-    circuit_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/circuit"
-
-    class_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/class"
-
-    school_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/school"
-
-    bpmn_val_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/hdBPMN-icdar2021/val"
-    bpmn_test_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/hdBPMN-icdar2021/test"
-    bpmn_train_path = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/hdBPMN-icdar2021/train"
-
-    fa_val_file_names = [f for f in listdir(fa_val_path)]
-    fa_test_file_names = [f for f in listdir(fa_test_path)]
-    fa_train_file_names = [f for f in listdir(fa_train_path)]
-
-    fa_json = annotate(fa_val_file_names, 'graph')
-    fa_test_json = annotate(fa_test_file_names, 'graph')
[…]

### 2025-05-11 src/dataset/extractor/category.py

@@ -1,0 +1,13 @@
+from enum import IntEnum
+
+
+class Category(IntEnum):
+    STATE = 1
+    FINAL_STATE = 2
+    TEXT = 3
+    ARROW = 4
+    CONNECTION = 5
+    DATA = 6
+    DECISION = 7
+    PROCESS = 8
+    TERMINATOR = 9

### 2025-05-11 src/dataset/extractor/dataset.py

@@ -1,0 +1,66 @@
+import json
+import os
+from typing import Tuple, Dict, List
+
+from torch.utils.data import Dataset
+from torchvision import tv_tensors as t
+from torchvision.io import read_image, ImageReadMode
+import torch
+
+
+class ObjectDetectionDataset(Dataset):
+    """
+    Dataset class for fine-tuning the object detection network
+    """
+
+    def __init__(self, annotations_file, img_dir, transform=None, target_transform=None):
+        with open(annotations_file, 'r') as f:
+            self.labels = json.load(f)
+        self.img_dir = img_dir
+        self.transform = transform
+        self.target_transform = target_transform
+
+    def __len__(self):
+        return len(self.labels['images'])
+
+    def __getitem__(self, idx) -> Tuple[t.Image, Dict]:
+        """
+        See https://docs.pytorch.org/tutorials/intermediate/torchvision_tutorial.html
+        """
+        images: List[Dict] = self.labels['images']
+        img_annotations: List[Dict] = [img_annotation for img_annotation in self.labels['annotations'] if
+                                       img_annotation['image_id'] == idx]
+        img_path = os.path.join(self.img_dir, images[idx]['file_name'])
+        image = read_image(str(img_path), ImageReadMode.RGB).float() / 255.0
+        target: Dict = {}
+        # 1st version following the tutorial
+        # boxes: t.BoundingBoxes = t.BoundingBoxes(data=[img_annotation['bbox'] for img_annotation in img_annotations],
+        #                                        format=t.BoundingBoxFormat.XYXY,
+        #                                       canvas_size=(images[idx]['height'], images[idx]['width']))
+
+        # 2nd version (chat)
+        # boxes = torch.tensor([img_annotation['bbox'] for img_annotation in img_annotations])
+
+        # 3rd version (chat)
+        boxes = torch.tensor([ann['bbox'] for ann in img_annotations], dtype=torch.float32)
+        if boxes.numel() == 0:
+            boxes = torch.empty((0, 4), dtype=torch.float32)
+        # Convert from [x, y, width, height] to [x1, y1, x2, y2]
+        boxes[:, 2:] += boxes[:, :2]
+
+        labels: torch.Tensor = torch.tensor([img_annotation['category_id'] for img_annotation in img_annotations], dtype=torch.int64)
+        # label = 0 -> background
+        image_id: int = images[idx]['id']  # must be unique for each image, used during evaluation
[…]

### 2025-05-11 src/dataset/extractor/merge_annotations.py

@@ -1,0 +1,130 @@
+import json
+from typing import List, Dict, Tuple
+
+
+def merge(test_annotations_path: str, train_annotations_path: str, val_annotations_path: str = '') -> Tuple[List, List]:
+    with open(test_annotations_path, 'r') as f:
+        test_annotations_json: Dict = json.load(f)
+
+    with open(train_annotations_path, 'r') as f:
+        train_annotations_json: Dict = json.load(f)
+
+    if val_annotations_path != '':
+        with open(val_annotations_path, 'r') as f:
+            val_annotations_json: Dict = json.load(f)
+
+    test_images: List = test_annotations_json['images']
+    train_images: List = train_annotations_json['images']
+    if val_annotations_path != '':
+        val_images: List = val_annotations_json['images']
+
+    test_annotations: List = test_annotations_json['annotations']
+    train_annotations: List = train_annotations_json['annotations']
+    if val_annotations_path != '':
+        val_annotations: List = val_annotations_json['annotations']
+
+    next_id: int = len(test_images)
+    images = test_images
+    annotations = test_annotations
+    for image in train_images:
+        for annotation in train_annotations:
+            if annotation['image_id'] == image['id']:
+                annotation['image_id'] = next_id
+                annotations.append(annotation)
+        image['id'] = next_id
+        next_id += 1
+        images.append(image)
+
+    if val_annotations_path != '':
+        for image in val_images:
+            for annotation in val_annotations:
+                if annotation['image_id'] == image['id']:
+                    annotation['image_id'] = next_id
+                    annotations.append(annotation)
+            image['id'] = next_id
+            next_id += 1
+            images.append(image)
+
+    return images, annotations
+
+
+def main():
+    fa_images, fa_annotations = merge('/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fa/test.json',
+                                      '/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fa/train.json',
+                                      '/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fa/val.json')
+    fca_images, fca_annotations = merge('/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fca/test.json',
+                                        '/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fca/train.json')
+    fcb_images, fcb_annotations = merge('/Users/saverionapolitano/PycharmProjects/diagram/dataset/source/fcb/test.json',
[…]

### 2025-05-11 src/dataset/extractor/test.py

@@ -1,0 +1,28 @@
+import unittest
+
+import torch
+
+from src.dataset.extractor.dataset import ObjectDetectionDataset
+import cv2
+
+
+class TestObjectDetectionDataset(unittest.TestCase):
+
+    def setUp(self):
+        self.dataset = ObjectDetectionDataset("/Users/saverionapolitano/PycharmProjects/diagram/dataset/extractor/labels.json",
+                                         "/Users/saverionapolitano/PycharmProjects/diagram/dataset/extractor/flow_graph_diagrams")
+
+    def test_len(self):
+        self.assertEqual(self.dataset.__len__(), 1391)
+
+    def test_get_item(self):
+        image, target = self.dataset.__getitem__(0)
+        im = cv2.imread(
+            '/Users/saverionapolitano/PycharmProjects/diagram/dataset/extractor/flow_graph_diagrams/writer018_fa_001.png', cv2.IMREAD_COLOR)
+        tensor_im = torch.from_numpy(im)
+        self.assertTrue(torch.equal(tensor_im.swapaxes(0, 2).swapaxes(1, 2), image))
+        self.assertEqual(target['boxes'].shape[0], 17)
+
+
+if __name__ == '__main__':
+    unittest.main()

### 2025-05-11 doc/article/article.tex

@@
 \section{Experiments}
-\subsection{Dataset}
-TODO
+\subsection{Text Digitization}
+In order to choose an appropriate model, some tests were done over our specific dataset employing some different, pre-made models.
+The average results over the dataset are shown in Table~\ref{tab:text_digitization}.
 
-\subsection{Metrics}
-TODO
-
-\subsection{Results}
-TODO
+\begin{table}[htbp]
+\caption{Text Distance Metrics}
+\centering
+\begin{tabular}{lccc}
+\hline
+\textbf{Model} & \textbf{Hamming} & \textbf{Cosine} & \textbf{Euclidean} \\
+\hline
+microsoft-trocr-small-printed & 1.392 & 0.150 & 7133 \\
+microsoft-trocr-base-printed & 1.286 & 0.081 & 3594 \\
+microsoft-trocr-small-handwritten & 1.346 & 0.100  & 3947 \\
+microsoft-trocr-base-handwritten & --- & --- & --- \\
+% Add more models as needed
+\hline
+\end{tabular}
+\label{tab:text_digitization}
+\end{table}
 

### 2025-05-11 src/extractor/bbox_detection/fine_tuning.py
Add bbox detection for extractor

@@ -1,0 +1,101 @@
+import matplotlib.pyplot as plt
+import torch
+from torch.utils.data import DataLoader
+from torchvision.models.detection import fasterrcnn_resnet50_fpn
+from torchvision.models.detection.faster_rcnn import FastRCNNPredictor
+from torchvision.transforms.functional import to_pil_image
+from torchvision.utils import draw_bounding_boxes
+from torchmetrics.detection.mean_ap import MeanAveragePrecision
+
+from src.dataset.extractor.dataset import ObjectDetectionDataset
+
+import logging
+
+logger = logging.getLogger(__name__)
+
+
+def collate_fn(batch):
+    return tuple(zip(*batch))
+
+
+# Setup paths
+annotations_file = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/extractor/labels.json"
+img_dir = "/Users/saverionapolitano/PycharmProjects/diagram/dataset/extractor/flow_graph_diagrams"
+
+# Create dataset and dataloader
+dataset = ObjectDetectionDataset(annotations_file, img_dir)
+dataloader = DataLoader(dataset, batch_size=4, shuffle=True, collate_fn=collate_fn)
+
+# Load model
+model = fasterrcnn_resnet50_fpn(weights="DEFAULT")  # fine-tuning
+# model = fasterrcnn_resnet50_fpn(weights=None, weights_backbone=None) # from scratch
+in_features = model.roi_heads.box_predictor.cls_score.in_features
+num_classes = 10  # number of classes (state, final state, text, arrow, connection, data, decision, process,
+# terminator) + 1 (for background)
+model.roi_heads.box_predictor = FastRCNNPredictor(in_features, num_classes)
+
+device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
+model.to(device)
+
+# Optimizer
+params = [p for p in model.parameters() if p.requires_grad]
+optimizer = torch.optim.SGD(params, lr=0.005, momentum=0.9, weight_decay=0.0005)
+
+# Training loop
+model.train()
+epochs = 100
+min_loss = 10000000
+for epoch in range(epochs):
+    for images, targets in dataloader:
+        images = [img.to(device) for img in images]
+
+        targets = [{k: v.to(device) for k, v in t.items()} for t in targets]
+
+        loss_dict = model(images, targets)
+        losses = sum(loss for loss in loss_dict.values())
+
+        optimizer.zero_grad()
+        losses.backward()
[…]

### 2025-05-12 doc/article/article.tex
Draft

@@
 microsoft-trocr-small-printed & 1.392 & 0.150 & 7133 \\
-microsoft-trocr-base-printed & 1.286 & 0.081 & 3594 \\
+% microsoft-trocr-base-printed & 1.286 & 0.081 & 3594 \\
 microsoft-trocr-small-handwritten & 1.346 & 0.100  & 3947 \\
-microsoft-trocr-base-handwritten & --- & --- & --- \\
-% Add more models as needed
+microsoft-trocr-base-handwritten & 1.549 & 0.057 & 3003 \\
 \hline
@@
 \end{table}
+
+Some examples of the different models applied to our graph diagram dataset can be seen in Figure~\ref{fig:text_digitization_results_1}.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\linewidth]{text_digitization_results_1.png}
+\caption{Some random results of the text digitization over the proposed models.}
+\label{fig:text_digitization_results_1}
+\end{figure}
 

### 2025-05-12 src/extractor/text_extraction/example.py

@@ -1,0 +1,101 @@
+import sys, os
+
+sys.path.append(os.path.dirname("."))
+sys.path.append(os.path.dirname("/"))
+
+from src.extractor.text_extraction.text_extractor import TrOCRTextExtractorSmall, TrOCRTextExtractorBase, TrOCRTextExtractionSmallHandwritten, TrOCRTextExtractorBaseHandwritten
+from src.extractor.text_extraction.text_extractor import TextExtractor
+from core.image.tensor_image import TensorImage, Image
+from torchvision.transforms.functional import to_pil_image
+from core.image.bbox.bbox2p import ImageBoundingBox2Points, ImageBoundingBox
+import json, torch
+import random
+
+import matplotlib.pyplot as plt
+from PIL import Image
+
+def show_ocr_comparison(images, ground_truths, preds: list[dict], model_names, max_width=200):
+    n = len(images)
+    fig, axes = plt.subplots(nrows=n, ncols=2, figsize=(10, 2.5 * n))
+    
+    if n == 1:
+        axes = [axes]  # Ensure it's iterable
+
+    for idx in range(n):
+        # Load and resize image
+        img = images[idx]
+        ratio = min(max_width / img.width, 1.0)
+        new_size = (int(img.width * ratio), int(img.height * ratio))
+        img = img.resize(new_size)
+
+        # Display image
+        axes[idx][0].imshow(img)
+        axes[idx][0].axis('off')
+
+        # Show texts
+        gt = ground_truths[idx]
+        text_lines = [(
+            f"Ground Truth: {gt}\n", "black"
+        )]
+        for model_name, pred in preds.items():
+            text_lines.append((f"{model_name}: {pred[idx]}\n", "green" if gt == pred[idx] else "red"))
+        
+        ax = axes[idx][1]
+        y = 0.9
+        for line, color in text_lines:
+            ax.text(0, y, line, fontsize=30, verticalalignment='top', fontfamily='monospace', color=color)
+            y -= 0.3  # Line spacing
+        ax.axis('off')
+
+    plt.tight_layout()
+    plt.show()
+
+if __name__ == "__main__":
+    # Create instances of the text extractors
+    small_extractor = TrOCRTextExtractorSmall()
+    # base_extractor = TrOCRTextExtractorBase()
+    small_handwritten = TrOCRTextExtractionSmallHandwritten()
+    base_handwritten = TrOCRTextExtractorBaseHandwritten()
+
[…]

### 2025-05-12 src/extractor/text_extraction/test.py

@@ -3,17 +3,19 @@ import sys, os
 sys.path.append(os.path.dirname("."))
 sys.path.append(os.path.dirname("/"))
 
-from src.extractor.text_extraction.text_extractor import TrOCRTextExtractorSmall, TrOCRTextExtractorBase, TrOCRTextExtractionSmallHandwritten
+from src.extractor.text_extraction.text_extractor import TrOCRTextExtractorSmall, TrOCRTextExtractorBase, TrOCRTextExtractionSmallHandwritten, TrOCRTextExtractorBaseHandwritten
 from src.extractor.text_extraction.text_extractor import TextExtractor
 from core.image.tensor_image import TensorImage, Image
 from core.image.bbox.bbox2p import ImageBoundingBox2Points, ImageBoundingBox
-import os, json, torch
+import json, torch
 
 if __name__ == "__main__":
     # Create instances of the text extractors
+    
     # small_extractor = TrOCRTextExtractorSmall()
     # base_extractor = TrOCRTextExtractorBase()
-    small_handwritten = TrOCRTextExtractionSmallHandwritten()
+    # small_handwritten = TrOCRTextExtractionSmallHandwritten()
+    base_handwritten = TrOCRTextExtractorBaseHandwritten()
     
     json_path = os.path.join(os.path.dirname(__file__), "../../../dataset/source/fa/test.json")
     with open(json_path, 'r') as json_file:
@@ -34,7 +36,7 @@ if __name__ == "__main__":
             for annotation in annotations:
                 x, y, w, h = annotation["bbox"]
                 text_bbox = ImageBoundingBox2Points("text", torch.Tensor([x, y, x + w, y + h]), 1)
-                metric = small_handwritten.compute_metrics([image], [text_bbox], [annotation["text"]])
+                metric = base_handwritten.compute_metrics([image], [text_bbox], [annotation["text"]])
                 partial_metric.append(metric)
             base_metrics.append({
                 "hamming": sum([metric["hamming"] for metric in partial_metric]) / len(partial_metric),
@@ -57,3 +59,5 @@ if __name__ == "__main__":
         "euclidean": sum([metric["euclidean"] for metric in base_metrics]) / len(base_metrics),
     }
     print("Base Extractor Average Metrics:" + str(base_avg_metrics))
+    
+    

### 2025-05-12 src/extractor/text_extraction/text_extractor.py

@@ -259,3 +259,31 @@ class TrOCRTextExtractionSmallHandwritten(TextExtractor):
         
         def compute_embedding(self, text: str) -> torch.Tensor:
             return self.processor.tokenizer(text, return_tensors="pt").input_ids
+        
+class TrOCRTextExtractorBaseHandwritten(TextExtractor):
+    
+            def __init__(self):
+                self.processor = TrOCRProcessor.from_pretrained("microsoft/trocr-base-handwritten")
+                self.model = VisionEncoderDecoderModel.from_pretrained("microsoft/trocr-base-handwritten")
+                self.model.eval()
+                self.model.to("cuda" if torch.cuda.is_available() else "cpu")
+            
+            def extract_text(self, image: Image, text_bbox: ImageBoundingBox) -> str:
+                """
+                Extracts text from the given image using the TrOCR model.
+        
+                Args:
+                    image (Image): The image object containing the file.
+        
+                Returns:
+                    str: The extracted text.
+                """
+                
+                cropped_image = self.crop_image(image, text_bbox)
+                pixel_values = self.processor(images=cropped_image, return_tensors="pt").pixel_values
+                generated_ids = self.model.generate(pixel_values)
+                generated_text: str = self.processor.batch_decode(generated_ids, skip_special_tokens=True)[0]
+                return generated_text.strip()
+            
+            def compute_embedding(self, text: str) -> torch.Tensor:
+                return self.processor.tokenizer(text, return_tensors="pt").input_ids

### 2025-05-12 src/extractor/text_extraction/example.py
TODO

@@ -83,7 +83,7 @@ if __name__ == "__main__":
             annotations = [annotation for annotation in json_content["annotations"] if annotation["image_id"] == image_id and annotation["category"] == "text" and annotation["text"] != None]
             image = TensorImage.from_str(file_path)
             for annotation in annotations:
-                if random.random() * (1 / len(annotation["text"]) ** 3) < 0.025:    
+                if random.random() * (1 / len(annotation["text"])) < 0.02:    
                     if ground_truths_glob.count(annotation["text"]) >= 1:
                         continue
                     x, y, w, h = annotation["bbox"]

### 2025-05-12 doc/article/article.tex
Draft

@@
 \subsubsection{Text Digitization}
-Since the task of converting an image of text into the relative string has already been solved in various ways, we settled upon the Microsoft \textit{Trocr Handwritten} Transformer model; more precisely, the \textit{small} version of it.
-[TODO: Fare esperimenti e provarne altri].
+Since the task of converting an image of text into the relative string has already been solved in various ways, we settled upon the Microsoft \textit{Trocr Handwritten} Transformer model; more precisely, the \textit{base} version of it.
 Every bounding box containing text that has been found is later passed to the model, and the resulting string generated by the model is linked to it.\\
@@
 
-Some examples of the different models applied to our graph diagram dataset can be seen in Figure~\ref{fig:text_digitization_results_1}.
+Some examples of the different models applied to our graph diagram dataset can be seen in Figure~\ref{fig:text_digitization_results_1} and in Figure~\ref{fig:text_digitization_results_2}.
 
@@
 \label{fig:text_digitization_results_1}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\linewidth]{text_digitization_results_2.png}
+\caption{Some results in which the \textit{trocr-base-handwritten} fails.}
+\label{fig:text_digitization_results_2}
 \end{figure}

### 2025-05-12 src/extractor/text_extraction/test.py
Large Handwritten

@@ -3,7 +3,7 @@ import sys, os
 sys.path.append(os.path.dirname("."))
 sys.path.append(os.path.dirname("/"))
 
-from src.extractor.text_extraction.text_extractor import TrOCRTextExtractorSmall, TrOCRTextExtractorBase, TrOCRTextExtractionSmallHandwritten, TrOCRTextExtractorBaseHandwritten
+from src.extractor.text_extraction.text_extractor import TrOCRTextExtractorSmall, TrOCRTextExtractorBase, TrOCRTextExtractorLargeHandwritten, TrOCRTextExtractionSmallHandwritten, TrOCRTextExtractorBaseHandwritten
 from src.extractor.text_extraction.text_extractor import TextExtractor
 from core.image.tensor_image import TensorImage, Image
 from core.image.bbox.bbox2p import ImageBoundingBox2Points, ImageBoundingBox
@@ -15,7 +15,8 @@ if __name__ == "__main__":
     # small_extractor = TrOCRTextExtractorSmall()
     # base_extractor = TrOCRTextExtractorBase()
     # small_handwritten = TrOCRTextExtractionSmallHandwritten()
-    base_handwritten = TrOCRTextExtractorBaseHandwritten()
+    # base_handwritten = TrOCRTextExtractorBaseHandwritten()
+    large_handwritten = TrOCRTextExtractorLargeHandwritten()
     
     json_path = os.path.join(os.path.dirname(__file__), "../../../dataset/source/fa/test.json")
     with open(json_path, 'r') as json_file:
@@ -36,7 +37,7 @@ if __name__ == "__main__":
             for annotation in annotations:
                 x, y, w, h = annotation["bbox"]
                 text_bbox = ImageBoundingBox2Points("text", torch.Tensor([x, y, x + w, y + h]), 1)
-                metric = base_handwritten.compute_metrics([image], [text_bbox], [annotation["text"]])
+                metric = large_handwritten.compute_metrics([image], [text_bbox], [annotation["text"]])
                 partial_metric.append(metric)
             base_metrics.append({
                 "hamming": sum([metric["hamming"] for metric in partial_metric]) / len(partial_metric),
@@ -58,6 +59,6 @@ if __name__ == "__main__":
         "cosine": sum([metric["cosine"] for metric in base_metrics]) / len(base_metrics),
         "euclidean": sum([metric["euclidean"] for metric in base_metrics]) / len(base_metrics),
     }
[…]

### 2025-05-12 src/extractor/text_extraction/text_extractor.py

@@ -287,3 +287,31 @@ class TrOCRTextExtractorBaseHandwritten(TextExtractor):
             
             def compute_embedding(self, text: str) -> torch.Tensor:
                 return self.processor.tokenizer(text, return_tensors="pt").input_ids
+            
+class TrOCRTextExtractorLargeHandwritten(TextExtractor):
+    
+            def __init__(self):
+                self.processor = TrOCRProcessor.from_pretrained("microsoft/trocr-large-handwritten")
+                self.model = VisionEncoderDecoderModel.from_pretrained("microsoft/trocr-large-handwritten")
+                self.model.eval()
+                self.model.to("cuda" if torch.cuda.is_available() else "cpu")
+            
+            def extract_text(self, image: Image, text_bbox: ImageBoundingBox) -> str:
+                """
+                Extracts text from the given image using the TrOCR model.
+        
+                Args:
+                    image (Image): The image object containing the file.
+        
+                Returns:
+                    str: The extracted text.
+                """
+                
+                cropped_image = self.crop_image(image, text_bbox)
+                pixel_values = self.processor(images=cropped_image, return_tensors="pt").pixel_values
+                generated_ids = self.model.generate(pixel_values)
+                generated_text: str = self.processor.batch_decode(generated_ids, skip_special_tokens=True)[0]
+                return generated_text.strip()
+            
+            def compute_embedding(self, text: str) -> torch.Tensor:
+                return self.processor.tokenizer(text, return_tensors="pt").input_ids

### 2025-05-12 doc/article/article.tex
Device

@@
 % 
-% Graph Diagram
+% Flowchart Diagram
 %
-\section{Graph Diagram Recognition}
+\section{Flowchart Diagram Recognition}
 
 \subsection{Internal Representation}
-In order to pass data through the components of the pipeline in a unified way, it was necessary to decide upon an internal representation of the graph diagram.
+In order to pass data through the components of the pipeline in a unified way, it was necessary to decide upon an internal representation of the flowchart diagram.
 It consists of \textbf{elements} and \textbf{relations}.
@@
 % microsoft-trocr-base-printed & 1.286 & 0.081 & 3594 \\
-microsoft-trocr-small-handwritten & 1.346 & 0.100  & 3947 \\
-microsoft-trocr-base-handwritten & 1.549 & 0.057 & 3003 \\
+microsoft-trocr-small-handwritten & \textbf{1.346} & 0.100  & 3947 \\
+microsoft-trocr-base-handwritten & 1.549 & \textbf{0.057} & \textbf{3003} \\
 \hline
@@
 
-Some examples of the different models applied to our graph diagram dataset can be seen in Figure~\ref{fig:text_digitization_results_1} and in Figure~\ref{fig:text_digitization_results_2}.
+Some examples of the different models applied to our flowchart diagram dataset can be seen in Figure~\ref{fig:text_digitization_results_1} and in Figure~\ref{fig:text_digitization_results_2}.
 

### 2025-05-12 src/extractor/text_extraction/test.py

@@ -12,6 +12,8 @@ import json, torch
 if __name__ == "__main__":
     # Create instances of the text extractors
     
+    
+    device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
     # small_extractor = TrOCRTextExtractorSmall()
     # base_extractor = TrOCRTextExtractorBase()
     # small_handwritten = TrOCRTextExtractionSmallHandwritten()

### 2025-05-12 src/extractor/text_extraction/text_extractor.py

@@ -307,8 +307,10 @@ class TrOCRTextExtractorLargeHandwritten(TextExtractor):
                     str: The extracted text.
                 """
                 
+                device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
                 cropped_image = self.crop_image(image, text_bbox)
-                pixel_values = self.processor(images=cropped_image, return_tensors="pt").pixel_values
+                pixel_values = self.processor(images=cropped_image.to(device), return_tensors="pt").pixel_values
+                self.model.to(device)
                 generated_ids = self.model.generate(pixel_values)
                 generated_text: str = self.processor.batch_decode(generated_ids, skip_special_tokens=True)[0]
                 return generated_text.strip()

### 2025-05-12 src/extractor/text_extraction/text_extractor.py
Device

@@ -309,7 +309,7 @@ class TrOCRTextExtractorLargeHandwritten(TextExtractor):
                 
                 device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
                 cropped_image = self.crop_image(image, text_bbox)
-                pixel_values = self.processor(images=cropped_image.to(device), return_tensors="pt").pixel_values
+                pixel_values = self.processor(images=cropped_image, return_tensors="pt").pixel_values.to(device)
                 self.model.to(device)
                 generated_ids = self.model.generate(pixel_values)
                 generated_text: str = self.processor.batch_decode(generated_ids, skip_special_tokens=True)[0]

### 2025-05-26 src/classifier/test.py
Preprocessor for Classification

@@ -9,5 +9,5 @@ import matplotlib.pyplot as plt
 
 if __name__ == "__main__":
     processor = GNRMultiProcessor()
-    image = TensorImage.from_str("test_resources/logo-wikipedia.png")
+    image = TensorImage.from_str("test_resources/handwritten_diagram.jpg")
     image = processor.process(image, debug=True)

### 2025-05-26 src/classifier/model.py
Update model.py

@@ -11,32 +11,50 @@ class ClassifierCNN(nn.Module):
     
     def __init__(self, num_classes: int):
         super(ClassifierCNN, self).__init__()
-        self.conv1 = nn.Conv2d(1, 2, kernel_size=3, stride=1, padding=1)
-        self.conv2 = nn.Conv2d(2, 4, kernel_size=3, stride=1, padding=1)
-        self.conv3 = nn.Conv2d(4, 8, kernel_size=3, stride=1, padding=1)
-        self.pool = nn.MaxPool2d(kernel_size=2, stride=2, padding=0)
-        self.fc1 = nn.Linear(64 * 8 * 8, 512)
-        self.fc2 = nn.Linear(512, 128)
-        self.fc3 = nn.Linear(128, num_classes)
-        self.relu = nn.ReLU()
-        self.dropout = nn.Dropout(p=0.5)
+        
+        class ConvLayer(nn.Module):
+            """
+            A convolutional layer followed by ReLU activation and max pooling.
+            """
+            def __init__(self, in_channels: int, out_channels: int, kernel_size: int = 3, stride: int = 1, padding: int = 1, kernel_pool_size: int = 2, pool_stride: int = 2, pool_padding: int = 0):
+                super(ConvLayer, self).__init__()
+                self.conv = nn.Conv2d(in_channels, out_channels, kernel_size=kernel_size, stride=stride, padding=padding)
+                self.relu = nn.ReLU()
+                self.pool = nn.MaxPool2d(kernel_size=kernel_pool_size, stride=pool_stride, padding=pool_padding)
+            def forward(self, x):
+                x = self.conv(x)
+                x = self.relu(x)
+                x = self.pool(x)
+                return x
+        
+        class LinearLayer(nn.Module):
+            """
+            A linear layer followed by ReLU activation.
+            """
+            def __init__(self, in_features: int, out_features: int, dropout_rate: float = 0.5):
+                super(LinearLayer, self).__init__()
+                self.linear = nn.Linear(in_features, out_features)
+                self.dropout = nn.Dropout(p=dropout_rate)
+                self.relu = nn.ReLU()
+            def forward(self, x):
+                x = self.linear(x)
+                x = self.relu(x)
+                x = self.dropout(x)
+                return x
+        
+        self.conv1 = ConvLayer(1, 2)
+        self.conv2 = ConvLayer(2, 4)
+        self.conv3 = ConvLayer(4, 8)
+        self.fc1 = LinearLayer(64*8*8, 512)
[…]

### 2025-05-26 src/classifier/dataset.py
Made Datatset Classifier

@@ -1,0 +1,41 @@
+from torch.utils.data import Dataset
+from core.image.tensor_image import TensorImage, Image
+import json
+
+class DatasetClassifier(Dataset):
+    """
+    A custom dataset class for image classification tasks.
+    
+    :param images: A list of image tensors.
+    :param labels: A list of labels corresponding to the images.
+    """
+    
+    def __init__(self, json_path: str = "dataset/classifier/labels.json", base_image_path: str = "dataset/classifier/all/"):
+        """
+        Initialize the dataset with images and labels.
+        
+        :param json_path: Path to the JSON file containing image paths and labels.
+        """
+        self.base_image_path = base_image_path
+        with open(json_path, 'r') as f:
+            data = json.load(f)
+        
+        self.images = []
+        self.labels = []
+        for key, value in data:
+            self.images.append(key)
+            self.labels.append(value)
+    
+    def __len__(self):
+        return len(self.images)
+    
+    def __getitem__(self, idx):
+        """
+        Get an item from the dataset.
+        
+        :param idx: The index of the item to retrieve.
+        
+        :return: A tuple containing the image tensor and its corresponding label.
+        """
+        image: Image = TensorImage().from_str(self.base_image_path + self.images[idx])
+        return self.images[idx], self.labels[idx]

### 2025-05-26 src/classifier/model.py
Update Classifier Model

@@ -68,6 +68,7 @@ class ClassifierCNN(nn.Module):
         :return: Output tensor.
         """
         x = self.sequential(x)
+        x = torch.argmax(x, dim=1)
         return x
     
     def load(self, path: str):

### 2025-05-26 src/classifier/training.py
Training classifier script

@@ -1,0 +1,11 @@
+import os, sys
+sys.path.append(os.path.dirname("."))
+sys.path.append(os.path.dirname("/"))
+
+from src.classifier.classifier import GNRClassifier
+
+if __name__ == "__main__":
+    
+    classifier = GNRClassifier()
+    classifier.train()
+    classifier.save_model(path="src/classifier/classifier.pth")

### 2025-05-26 src/classifier/dataset.py
Updated dataloader + model classifier

@@ -1,6 +1,7 @@
 from torch.utils.data import Dataset
 from core.image.tensor_image import TensorImage, Image
 import json
+from src.classifier.preprocessing.processor import Processor, GNRMultiProcessor
 
 class DatasetClassifier(Dataset):
     """
@@ -10,19 +11,20 @@ class DatasetClassifier(Dataset):
     :param labels: A list of labels corresponding to the images.
     """
     
-    def __init__(self, json_path: str = "dataset/classifier/labels.json", base_image_path: str = "dataset/classifier/all/"):
+    def __init__(self, preprocessor: Processor = GNRMultiProcessor(), json_path: str = "dataset/classifier/labels.json", base_image_path: str = "dataset/classifier/all/"):
         """
         Initialize the dataset with images and labels.
         
         :param json_path: Path to the JSON file containing image paths and labels.
         """
         self.base_image_path = base_image_path
+        self.preprocessor = preprocessor
         with open(json_path, 'r') as f:
             data = json.load(f)
         
         self.images = []
         self.labels = []
-        for key, value in data:
+        for key, value in data.items():
             self.images.append(key)
             self.labels.append(value)
     
@@ -37,5 +39,6 @@ class DatasetClassifier(Dataset):
         
         :return: A tuple containing the image tensor and its corresponding label.
         """
-        image: Image = TensorImage().from_str(self.base_image_path + self.images[idx])
-        return self.images[idx], self.labels[idx]
+        image: Image = TensorImage.from_str(self.base_image_path + self.images[idx])
+        image = self.preprocessor.process(image)
+        return image, self.labels[idx]

### 2025-05-26 src/classifier/model.py

@@ -68,7 +68,6 @@ class ClassifierCNN(nn.Module):
         :return: Output tensor.
         """
         x = self.sequential(x)
-        x = torch.argmax(x, dim=1)
         return x
     
     def load(self, path: str):

### 2025-05-26 src/dataset/classifier/test.py
Fix classifier dataset folder

@@ -17,12 +17,12 @@ class TestClassifierDataset(unittest.TestCase):
         self.assertEqual(self.dataset.__len__(), 9216)
 
     def test_get_item(self):
-        image, label = self.dataset.__getitem__(0)
+        image, label = self.dataset.__getitem__(1098)
         im = cv2.imread(
-            '/Users/saverionapolitano/PycharmProjects/diagram/dataset/classifier/all/writer013_fa_007.png', cv2.IMREAD_UNCHANGED)
+            '/Users/saverionapolitano/PycharmProjects/diagram/dataset/classifier/all/writer017_fc_008b.png', cv2.IMREAD_UNCHANGED)
         tensor_im = torch.from_numpy(im)
         self.assertTrue(torch.equal(tensor_im[torch.newaxis, :, :], image.as_tensor()))
-        self.assertEqual(label, 'graph')
+        self.assertEqual(label, 'flowchart')
 
 
 if __name__ == '__main__':

### 2025-05-26 doc/article/article.tex
Device

@@
 
+\subsubsection{Arrow Orientation}
+To later correctly assign arrows to their source and target elements, the need to determine where the tip of an arrow lies.
+This task involves a few problems, among which determining exactly the arrow shape and its end - because sometimes other pen signs may be found in the bounding box of the arrow itself.
+[TODO: Dire l'approccio finale usato].
+\\
+
 \subsubsection{Text Digitization}
@@
 \section{Experiments}
+
+\subsection{Classifier}
+TODO
+
+\subsection{Object Detection}
+TODO
+
+\subsection{Arrow Orientation}
+As we've seen in the previous section, the need to find the exact position of an arrow's head arises.
+The first algorithmic approach proposed to solve this task - called \textit{Double Clustering} - was as follows: 
+\begin{enumerate}
+	\item Apply a clustering algorithm to the bounding box.
+	\item Suppose that the biggest cluster belongs to the arrow; the other clusters are thought to contain useless elements for this task, and can be discarded.
+	\item Find the keypoints of the arrow using SIFT.
+	\item Find the two keypoints inside the arrow with maximum distance. Assume that these two points are closely related to the two ends of the arrow.
+	\item Run a template matching procedure inside the two clusters to find the head of the arrow.
+\end{enumerate}
+
+For Step (1), we tried to employ both Spectral Clustering (see Figure~\ref{fig:spectral_clustering}) and DBSCan (see Figure~\ref{fig:arrow_double_clustering_clusters}).
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=150pt]{spectral_clustering.png}
+\caption{The result of Spectral Clustering to one instance of an arrow bounding box in our dataset.}
+\label{fig:spectral_clustering}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=150pt]{arrow_double_clustering_clusters.png}
+\caption{The result of DBScan to one instance of an arrow bounding box in our dataset.}
+\label{fig:arrow_double_clustering_clusters}
+\end{figure}
+
+For Step (3), we used SIFT in order to find keypoints and some descriptors for them - necessary for the later template matching.
+During Step (4), we were able to correctly deduce the overall direction of the arrow (see Figure~\ref{fig:arrow_double_clustering_direction}).
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=150pt]{arrow_double_clustering_direction.png}
+\caption{The estimated direction of the arrow, finding the points that maximize the distance inside the cluster.}
+\label{fig:arrow_double_clustering_direction}
+\end{figure}
+
+For Step (5), we had previously hand-picked some templates (see Figure~\ref{fig:templates}).
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=50pt]{templates.png}
+\caption{An example of hand-picked template with SIFT descriptors.}
+\label{fig:templates}
+\end{figure}
+
+However, this approach didn't work as expected (see Figure~\ref{fig:template_matching_fail}). [TODO: Spiegare perchè].
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=150pt]{template_matching_fail.png}
+\caption{The overall result of the algorithm.}
+\label{fig:template_matching_fail}
+\end{figure}
+
 \subsection{Text Digitization}
@@
 % microsoft-trocr-base-printed & 1.286 & 0.081 & 3594 \\
-microsoft-trocr-small-handwritten & \textbf{1.346} & 0.100  & 3947 \\
-microsoft-trocr-base-handwritten & 1.549 & \textbf{0.057} & \textbf{3003} \\
+microsoft-trocr-small-handwritten & \textbf{1.346} & 0.100  & 3947 \\ \hline \textbf{microsoft-trocr-base-handwritten} & 1.549 & \textbf{0.057} & \textbf{3003} \\
 \hline

### 2025-06-13 doc/article/article.tex
Update

@@
 \subsection{Classifier}
-TODO
+The first architecture explored for the Classifier module was a Convolutional Neural Network (CNN) following AlexNet's architecture, but smaller in size.
+It had two Convolutional layers, each followed by a batch normalization and then by a max pooling layers. The first layer had only 4 filters, while the second one had 8.
+The output of the second layer was then flattened and passed through a fully connected layer which downsized the features to 128, followed by a ReLU activation function.
+Other two fully connected layer brought the number of output features down to the number of classes.
+Then, a softmax activation function was applied to the output layer, in order to obtain the probabilities of each class.
+The model was trained on our own dataset, with only three possible output classes: 'flowchart', 'graph' and 'other'; it was trained for 10 epochs, with an AdamW optimizer (with a learning rate of 1e-3) and a standard CrossEntropy Loss function.
+
+We first tried such a small network because the features that had to be recognized were very simple, down to the basic shapes of the diagram elements.
+However, the results showed a non-decreasing value of the loss function of 0.9 - better than $-ln(3) \seq 1.1$, which is random guessing, but still not satisfactory.
+
+Judging this to be a problem in architecture size, we then tried adding another CNN layer with 16 filters; this one too followed by a batch normalization and a max pooling layer.
+The rest of the net was untouched; we also decreased the learning rate to 1e-4, in order to avoid overshooting the minimum of the loss function.
+The results, however, didn't show any kind of improvement, with the loss function still being around 0.9.
+
+While analyzing the predictions, we came up with what was actually the problem: the dataset was heavily unbalanced, with the 'other' class being the most represented one, shadowing the other two classes.
+In order to solve this problem, we decided to use a weighted CrossEntropy Loss function, with the weights being inversely proportional to the number of samples in each class.
+
+TODO: Scrivi l'algoritmo
+
+This final approach showed a significant improvement in the results, with the loss function decreasing to around 0.6 after 10 epochs of training, with a global accuracy score of 97\% in the validation set.
 

### 2025-06-13 src/classifier/classifier.py

@@ -145,23 +145,24 @@ class GNRClassifier(Classifier):
         self.model.load(path)
 
     
-    def evaluate(self, dataset: DatasetClassifier = DatasetClassifier(), visual_output: bool = False, verbose: bool = False) -> float:
+    def evaluate(self, dataset: DatasetClassifier = DatasetClassifier(), visual_output: bool = False, subset_percentage: float = 0.1) -> float:
         """
         Evaluate the model on the provided dataset.
         
         :param dataset: The dataset to evaluate the model on.
         :param visual_output: If True, will visualize some predictions.
-        :param verbose: If True, will print detailed evaluation information.
+        :param subset_percentage: The percentage of the dataset to use for evaluation. Defaults to 10%.
         :return: The accuracy of the model on the dataset.
         """
 
-        subset_size = int(0.1 * len(dataset))
+        subset_size = int(subset_percentage * len(dataset))
         _, subset = random_split(dataset, [len(dataset) - subset_size, subset_size])
         dataloader = DataLoader(subset, batch_size=32, shuffle=True)
 
         device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
         self.model.to(device)
 
+        # Computw the accuracy of the model on the dataset
         self.model.eval()
         correct = 0
         total = 0
@@ -185,6 +186,8 @@ class GNRClassifier(Classifier):
         accuracy = 100 * correct / total
         classes_names = {k: v for v, k in self.classes.items()}
         print(f"Accuracy of the model on the dataset: {accuracy:.2f}%")
+
+        # If visual_output is True, plot the confusion matrix
         if visual_output:
             import matplotlib.pyplot as plt
 
@@ -194,4 +197,5 @@ class GNRClassifier(Classifier):
             disp.plot(ax=ax, cmap=plt.cm.Blues, values_format='d')
             plt.title("Confusion Matrix")
             plt.show()
+            plt.waitforbuttonpress()
         return accuracy

### 2025-06-13 src/classifier/dataset.py

@@ -1,6 +1,6 @@
 from torch.utils.data import Dataset
 from core.image.tensor_image import TensorImage, Image
-import json
+import json, random
 from src.classifier.preprocessing.processor import Processor, GNRMultiProcessor
 
 class DatasetClassifier(Dataset):
@@ -11,7 +11,7 @@ class DatasetClassifier(Dataset):
     :param labels: A list of labels corresponding to the images.
     """
     
-    def __init__(self, preprocessor: Processor = GNRMultiProcessor(), json_path: str = "dataset/classifier/labels.json", base_image_path: str = "dataset/classifier/all/"):
+    def __init__(self, preprocessor: Processor = GNRMultiProcessor(), json_path: str = "dataset/classifier/labels.json", base_image_path: str = "dataset/classifier/all/", undersample_classes: list[str] = []):
         """
         Initialize the dataset with images and labels.
         
@@ -28,6 +28,8 @@ class DatasetClassifier(Dataset):
         self.labels = []
         idx = 0
         for key, value in data.items():
+            if value in undersample_classes and random.random() < 0.4:
+                continue
             self.images.append(key)
             self.labels.append(value)
             if value not in self.classes:

### 2025-06-13 src/classifier/training.py

@@ -8,14 +8,26 @@ from src.classifier.dataset import DatasetClassifier
 if __name__ == "__main__":
     
     """
-    Training script for the GNRClassifier.
+    # Training script for the GNRClassifier.
 
     dataset = DatasetClassifier()
     classifier = GNRClassifier(classes=dataset.classes)
     classifier.train(dataset, verbose=True, batch_size=64, epochs=10)
     classifier.save_model(path="src/classifier/classifier.pth")
-    """
+    
+    # Evaluation script for the GNRClassifier.
 
     dataset = DatasetClassifier()
     classifier = GNRClassifier(dataset.classes, model_path="src/classifier/classifier.pth")
     classifier.evaluate(visual_output=True, verbose=True)
+
+    # Finetune the GNRClassifier without the 'other' class.
+    dataset = DatasetClassifier(undersample_classes=["other"])
+    classifier = GNRClassifier(classes=dataset.classes, model_path="src/classifier/classifier.pth")
+    classifier.train(dataset, verbose=True, batch_size=64, epochs=3, learning_rate=5e-5)
+    classifier.save_model(path="src/classifier/classifier.pth")
+    """
+
+    dataset = DatasetClassifier()
+    classifier = GNRClassifier(dataset.classes, model_path="src/classifier/classifier.pth")
+    classifier.evaluate(visual_output=True, subset_percentage=0.25)

### 2025-06-15 doc/article/article.tex

@@
 \usepackage{float}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 
@@
 The model was trained on our own dataset, with only three possible output classes: 'flowchart', 'graph' and 'other'; it was trained for 10 epochs, with an AdamW optimizer (with a learning rate of 1e-3) and a standard CrossEntropy Loss function.
+\\
 
 We first tried such a small network because the features that had to be recognized were very simple, down to the basic shapes of the diagram elements.
-However, the results showed a non-decreasing value of the loss function of 0.9 - better than $-ln(3) \seq 1.1$, which is random guessing, but still not satisfactory.
+However, the results showed a non-decreasing value of the loss function of 0.9 - better than $-ln(3) \simeq 1.1$, which is random guessing, but still not satisfactory.
+\\
 
@@
 The results, however, didn't show any kind of improvement, with the loss function still being around 0.9.
+\\
 
@@
 
-TODO: Scrivi l'algoritmo
+\begin{algorithm}
+	\caption{Cross-Entropy weight computation}
+	\begin{algorithmic}[1]
+		\REQUIRE The classes' frequencies $f_1$, $f_2$, ..., $f_c$
+		\STATE $S \gets \sum_{i = 1}^c f_i$
+		\STATE $f_i \gets (f_i / S)^{-1}$
+		\STATE $S \gets \sum_{i = 1}^c f_i$
+		\STATE $f_i \gets f_i / S$
+		\RETURN $f_1$, $f_2$, ..., $f_c$
+	\end{algorithmic}
+\end{algorithm}
 
@@
 % microsoft-trocr-base-printed & 1.286 & 0.081 & 3594 \\
-microsoft-trocr-small-handwritten & \textbf{1.346} & 0.100  & 3947 \\ \hline \textbf{microsoft-trocr-base-handwritten} & 1.549 & \textbf{0.057} & \textbf{3003} \\
+microsoft-trocr-small-handwritten & \textbf{1.346} & 0.100  & 3947 \\ 
+\hline 
+\textbf{microsoft-trocr-base-handwritten} & 1.549 & \textbf{0.057} & \textbf{3003} \\
 \hline

### 2025-06-17 doc/article/article.tex
Article update

@@
 \subsubsection{Preprocessing}
-TODO \\
+Due to the fact that various kinds of diagrams' images could be submitted to the system, it was decided to streamline a unified preprocessing pipeline in order to reduce the differences that our classifier model had to handle.
+A few problems were analyzed and the final preprocessing pipeline for the classifier module is composed of the following operators, in order:
+\begin{itemize}
+	\item \textbf{Gray Scaling}: In order to remove colors from images, which was deemed an useless information for classifying diagrams.
+	\item \textbf{Otsu Thresholding}: To binarize the image and reducing the information to handle.
+	\item \textbf{Median Filtering}: Some images in our dataset - and in the use-case - were taken on checkered notebooks.
+The application of a median filter, with kernel size equal to three, totally removed these.
+	\item \textbf{Perspective Correction}: The diagrams' images, sometimes, didn't occupy the full photo frame.
+Finding four keypoints (if possible) and stretching the image helped to fully employ the frame.
+	\item \textbf{Padding}: This was necessary in order to keep the same net for every image. Changes in resolution would have implied changes in the number of paramters of the MLP layers.
+\end{itemize}
+
+Figure~\ref{fig:classifier_preprocessing} shows the step-by-step process of preparing an input image for the network.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=250pt]{classifier_preprocessing.png}
+	\caption{
+		The operations applied by the preprocessing module.
+		\textbf{(a)}: The raw image, still colored.
+		\textbf{(b)}: The image in gray scale.
+		\textbf{(c)}: The binarized image, thanks to Otsu.
+		\textbf{(d)}: The median filter result. Some previous thin lines are now no more.
+		\textbf{(e)}: The image with a perspective correction.
+		\textbf{(f)}: The image in a standard size, ready to enter the network.
+	}
+	\label{fig:classifier_preprocessing}
+\end{figure}
 
@@
 \subsection{Arrow Orientation}
-As we've seen in the previous section, the need to find the exact position of an arrow's head arises.
+As we've seen in the previous section, the need to find the exact position of an arrow's head arises (see \textbf{Appendix A} for a full overview of the first approach).
+
+TODO: Current approach.
+
+\subsection{Text Digitization}
+In order to choose an appropriate model, some tests were done over our specific dataset employing some different, pre-made models.
+The average results over the dataset are shown in Table~\ref{tab:text_digitization}.
+
+\begin{table}[htbp]
+\caption{Text Distance Metrics}
+\centering
+\begin{tabular}{lccc}
+\hline
+\textbf{Model} & \textbf{Hamming} & \textbf{Cosine} & \textbf{Euclidean} \\
+\hline
+microsoft-trocr-small-printed & 1.392 & 0.150 & 7133 \\
+% microsoft-trocr-base-printed & 1.286 & 0.081 & 3594 \\
+microsoft-trocr-small-handwritten & \textbf{1.346} & 0.100  & 3947 \\ 
+\hline 
+\textbf{microsoft-trocr-base-handwritten} & 1.549 & \textbf{0.057} & \textbf{3003} \\
+\hline
+\end{tabular}
+\label{tab:text_digitization}
+\end{table}
+
+Some examples of the different models applied to our flowchart diagram dataset can be seen in Figure~\ref{fig:text_digitization_results_1} and in Figure~\ref{fig:text_digitization_results_2}.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\linewidth]{text_digitization_results_1.png}
+\caption{Some random results of the text digitization over the proposed models.}
+\label{fig:text_digitization_results_1}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\linewidth]{text_digitization_results_2.png}
+\caption{Some results in which the \textit{trocr-base-handwritten} fails.}
+\label{fig:text_digitization_results_2}
+\end{figure}
+
+\section{Discussion}
+TODO
+
+\section{Conclusion and Future Work}
+TODO
+
+\bibliographystyle{IEEEtran}
+\bibliography{references}
+
+\begin{appendices}
+
+\section{Double Clustering}
 The first algorithmic approach proposed to solve this task - called \textit{Double Clustering} - was as follows: 
@@
 
-\subsection{Text Digitization}
-In order to choose an appropriate model, some tests were done over our specific dataset employing some different, pre-made models.
-The average results over the dataset are shown in Table~\ref{tab:text_digitization}.
-
-\begin{table}[htbp]
-\caption{Text Distance Metrics}
-\centering
-\begin{tabular}{lccc}
-\hline
-\textbf{Model} & \textbf{Hamming} & \textbf{Cosine} & \textbf{Euclidean} \\
-\hline
-microsoft-trocr-small-printed & 1.392 & 0.150 & 7133 \\
-% microsoft-trocr-base-printed & 1.286 & 0.081 & 3594 \\
-microsoft-trocr-small-handwritten & \textbf{1.346} & 0.100  & 3947 \\ 
-\hline 
-\textbf{microsoft-trocr-base-handwritten} & 1.549 & \textbf{0.057} & \textbf{3003} \\
-\hline
-\end{tabular}
-\label{tab:text_digitization}
-\end{table}
-
-Some examples of the different models applied to our flowchart diagram dataset can be seen in Figure~\ref{fig:text_digitization_results_1} and in Figure~\ref{fig:text_digitization_results_2}.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\linewidth]{text_digitization_results_1.png}
-\caption{Some random results of the text digitization over the proposed models.}
-\label{fig:text_digitization_results_1}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\linewidth]{text_digitization_results_2.png}
-\caption{Some results in which the \textit{trocr-base-handwritten} fails.}
-\label{fig:text_digitization_results_2}
-\end{figure}
-
-\section{Discussion}
-TODO
-
-\section{Conclusion and Future Work}
-TODO
-
-\bibliographystyle{IEEEtran}
-\bibliography{references}
+\end{appendices}
 

### 2025-06-17 core/preprocessing/processor.py

@@ -1,6 +1,7 @@
 from abc import ABC, abstractmethod
 from core.image.image import Image
 import matplotlib.pyplot as plt
+import matplotlib.patches as patches
 
 class Processor(ABC):
     """
@@ -36,17 +37,45 @@ class MultiProcessor(Processor):
         Process the input image using all processors in the list.
 
         Args:
-            debug:
+            debug: If True, display the image after each processing step.
             image: The input image to be processed.
 
         Returns:
             The processed image.
         """
+        steps = []
+        if debug:
+            steps.append(image.tensor.permute(1, 2, 0).squeeze())
+
         for processor in self.processors:
             image = processor.process(image)
             if debug:
-                plt.imshow(image.tensor.squeeze(), cmap="gray")
+                steps.append(image.tensor.squeeze())
+                plt.imshow(steps[-1], cmap="gray")
                 plt.axis("off")
                 plt.show()
+
+        if debug:
+            labels = [f"({chr(97 + i)})" for i in range(len(steps))]
+            fig, axes = plt.subplots((len(steps) + 1) // 2, 2)
+            axes = axes.flatten()
+            for ax, img, label in zip(axes, steps, labels):
+                ax.imshow(img, cmap="gray")
+                ax.axis("off")
+                h, w = img.shape[:2]
+                rect = patches.Rectangle((0.5, 0.5), w - 1, h - 1, linewidth=2, edgecolor='black', facecolor='none', transform=ax.transData)
+                ax.add_patch(rect)
+                ax.text(
+                    0.95, 0.05, label,
+                    transform=ax.transAxes,
+                    fontsize=10,
+                    fontweight='bold',
+                    ha='right', va='top',
+                    bbox=dict(boxstyle='round,pad=0.1', facecolor='white', edgecolor='none', alpha=0.6)
+                )
+            for ax in axes[len(steps):]:
+                ax.axis("off")
+            plt.tight_layout()
+            plt.show()
         return image
         

### 2025-06-21 src/classifier/preprocessing/processor.py
Added reverseProcessor

@@ -6,6 +6,25 @@ import numpy as np
 import cv2
 import torch
 
+class ReverseProcessor(Processor):
+    def __init__(self):
+        super().__init__()
+
+    def process(self, image: Image) -> Image:
+        """
+        Reverse the image tensor.
+        
+        :param image: The input image to be processed.
+        :return: The processed image with reversed tensor.
+        """
+        image_np = image.as_tensor().detach().cpu().numpy()
+        if len(image_np.shape) == 2:
+            return TensorImage(torch.from_numpy(image_np[::-1, ::-1]))
+        elif len(image_np.shape) == 3:
+            return TensorImage(torch.from_numpy(image_np[:, ::-1, ::-1]))
+        else:
+            raise ValueError(f"Unsupported image shape: {image_np.shape}")
+
 class GrayScaleProcessor(Processor):
     def __init__(self):
         super().__init__()

### 2025-06-21 src/classifier/classifier.py
Under/oversampling

@@ -14,10 +14,16 @@ import torch
 from sklearn.metrics import confusion_matrix, ConfusionMatrixDisplay
 
 from src.wellknown_diagram import WellKnownDiagram
-import logging
+import logging, sys
 
 logger = logging.getLogger(__name__)
-recognizable_diagrams = [WellKnownDiagram.GRAPH_DIAGRAM, WellKnownDiagram.FLOW_CHART,  WellKnownDiagram.OTHER]
+handler = logging.StreamHandler(sys.stdout)
+formatter = logging.Formatter('%(asctime)s - %(levelname)s - %(message)s')
+handler.setFormatter(formatter)
+
+logger.handlers.clear()
+logger.addHandler(handler)
+recognizable_diagrams = [WellKnownDiagram.FLOW_CHART, WellKnownDiagram.GRAPH_DIAGRAM,  WellKnownDiagram.OTHER]
 
 class GNRClassifier(Classifier):
     
@@ -102,7 +108,7 @@ class GNRClassifier(Classifier):
             running_loss = 0.0
             for i, (images, labels) in enumerate(train_dataloader):
                 if (i + 1) % 20 == 0 and verbose:
-                    logger.info(f"Epoch [{epoch + 1}/{epochs}], Step [{i + 1}/{len(train_dataloader)}]")
+                    print(f"Epoch [{epoch + 1}/{epochs}], Step [{i + 1}/{len(train_dataloader)}]")
 
                 optimizer.zero_grad()
                 
@@ -116,7 +122,7 @@ class GNRClassifier(Classifier):
                 running_loss += loss.item()
                 
             if verbose:
-                logger.info(f"Epoch [{epoch + 1}/{epochs}], Step [{i + 1}/{len(train_dataloader)}], Loss: {running_loss / len(train_dataloader) :.4f}")
+                print(f"Epoch [{epoch + 1}/{epochs}], Step [{i + 1}/{len(train_dataloader)}], Loss: {running_loss / len(train_dataloader) :.4f}")
                 
         ### Testing loop
         self.model.eval()
@@ -125,7 +131,7 @@ class GNRClassifier(Classifier):
         with torch.no_grad():
             for images, labels in val_dataloader:
                 outputs = self.model(images.float().to(device))
-                labels = torch.tensor([label_to_index[label] for label in labels], dtype=torch.long).to(device)
+                labels = torch.tensor([label_to_index.index(label) for label in labels], dtype=torch.long).to(device)
 
[…]

### 2025-06-21 src/classifier/dataset.py

@@ -1,9 +1,35 @@
 from torch.utils.data import Dataset
 from core.image.tensor_image import TensorImage, Image
 import json, random
-from src.classifier.preprocessing.processor import Processor, GNRMultiProcessor
+from src.classifier.preprocessing.processor import Processor, GNRMultiProcessor, ReverseProcessor
 
 from src.wellknown_diagram import WellKnownDiagram
+from dataclasses import dataclass, field
+from typing import List
+
+class ImagePath:
+
+        def __init__(self, name: str, reversed: bool = False):
+            """
+            Initialize the ImagePath with a path and an optional reversed flag.
+            
+            :param name: The name of the image.
+            :param path: The path to the image.
+            :param reversed: Whether the image should be reversed (default is False).
+            """
+            self.name = name
+            self.reversed = reversed
+        
+        def get(self, base_path: str) -> Image:
+            """
+            Get the image from the path.
+            
+            :return: An Image object loaded from the path.
+            """
+            image: Image = TensorImage.from_str(base_path + self.name)
+            if self.reversed:
+                image = ReverseProcessor().process(image)
+            return image
 
 class DatasetClassifier(Dataset):
     """
@@ -12,8 +38,9 @@ class DatasetClassifier(Dataset):
     :param images: A list of image tensors.
     :param labels: A list of labels corresponding to the images.
     """
-    
-    def __init__(self, preprocessor: Processor = GNRMultiProcessor(), json_path: str = "dataset/classifier/labels.json", base_image_path: str = "dataset/classifier/all/", undersample_classes: list[str] = []):
+
+
+    def __init__(self, preprocessor: Processor = GNRMultiProcessor(), json_path: str = "dataset/classifier/labels.json", base_image_path: str = "dataset/classifier/all/", undersample_classes: list[str] = [], upsample_classes: list[str] = []):
         """
         Initialize the dataset with images and labels.
         
@@ -26,13 +53,16 @@ class DatasetClassifier(Dataset):
         
         freqs = {}
         self.classes = {}
-        self.images = []
+        self.images: list[ImagePath] = []
         self.labels = []
         idx = 0
         for key, value in data.items():
-            if value in undersample_classes and random.random() < 0.4:
+            if value in undersample_classes and random.random() < 0.3:
                 continue
-            self.images.append(key)
+            if value in upsample_classes and random.random() < 0.3:
+                self.images.append(ImagePath(key, reversed=True))
[…]

### 2025-06-21 src/classifier/preprocessing/processor.py

@@ -19,9 +19,9 @@ class ReverseProcessor(Processor):
         """
         image_np = image.as_tensor().detach().cpu().numpy()
         if len(image_np.shape) == 2:
-            return TensorImage(torch.from_numpy(image_np[::-1, ::-1]))
+            return TensorImage(torch.from_numpy(image_np[::-1, ::-1].copy()))
         elif len(image_np.shape) == 3:
-            return TensorImage(torch.from_numpy(image_np[:, ::-1, ::-1]))
+            return TensorImage(torch.from_numpy(image_np[:, ::-1, ::-1].copy()))
         else:
             raise ValueError(f"Unsupported image shape: {image_np.shape}")
 

### 2025-06-21 src/classifier/training.py

@@ -7,23 +7,24 @@ from src.classifier.dataset import DatasetClassifier
 
 if __name__ == "__main__":
     
-    """
+    """""
     # Training script for the GNRClassifier.
 
     dataset = DatasetClassifier()
     classifier = GNRClassifier(classes=dataset.classes)
-    classifier.train(dataset, verbose=True, batch_size=64, epochs=10)
+    classifier.train(dataset, verbose=True, batch_size=64, epochs=5)
     classifier.save_model(path="src/classifier/classifier.pth")
     
+
     # Evaluation script for the GNRClassifier.
 
     dataset = DatasetClassifier()
     classifier = GNRClassifier(dataset.classes, model_path="src/classifier/classifier.pth")
-    classifier.evaluate(visual_output=True, verbose=True)
-
+    classifier.evaluate(visual_output=True, subset_percentage=0.1)
     """
-    # Finetune the GNRClassifier without the 'other' class.
-    dataset = DatasetClassifier(undersample_classes=["other"])
+    
+    # Finetune the GNRClassifier without any under/oversampling.
+    dataset = DatasetClassifier(undersample_classes=["other"], upsample_classes=["flowchart", "graph"])
     classifier = GNRClassifier(classes=dataset.classes, model_path="src/classifier/classifier.pth")
     classifier.train(dataset, verbose=True, batch_size=64, epochs=3, learning_rate=5e-5)
     classifier.save_model(path="src/classifier/classifier.pth")

### 2025-06-21 src/classifier/classifier.py
update bbox_detection

@@ -220,6 +220,6 @@ class GNRClassifier(Classifier):
     def get_default() -> 'GNRClassifier':
         return GNRClassifier(
             classes=recognizable_diagrams,
-            model_path="src/classifier/model.pth",
+            model_path="src/classifier/model_10.pth",
             processor=GNRMultiProcessor()
         )

### 2025-06-21 src/extractor/bbox_detection/fine_tuning.py

@@ -3,6 +3,7 @@ import torch
 from torch.utils.data import DataLoader
 from torchvision.models.detection import fasterrcnn_resnet50_fpn
 from torchvision.models.detection.faster_rcnn import FastRCNNPredictor
+from torchmetrics.detection.mean_ap import MeanAveragePrecision
 
 from src.dataset.extractor.dataset import ObjectDetectionDataset
 from torch.utils.data import random_split
@@ -40,8 +41,8 @@ def fine_tune():
 
     train_dataset, val_dataset = random_split(dataset, [train_size, val_size])
 
-    train_dataloader = DataLoader(train_dataset, batch_size=32, shuffle=True, collate_fn=collate_fn)
-    val_dataloader = DataLoader(val_dataset, batch_size=32, shuffle=True, collate_fn=collate_fn)
+    train_dataloader = DataLoader(train_dataset, batch_size=4, shuffle=True, collate_fn=collate_fn)
+    val_dataloader = DataLoader(val_dataset, batch_size=4, shuffle=True, collate_fn=collate_fn)
     # test_dataloader = DataLoader(test_dataset, batch_size=4, shuffle=True, collate_fn=collate_fn)
 
 
@@ -64,7 +65,7 @@ def fine_tune():
 
     # Training loop
     model.train()
-    epochs = 10
+    epochs = 100
     final_loss: float = 0.0
     for epoch in range(epochs):
         running_loss = 0.0
@@ -86,11 +87,56 @@ def fine_tune():
         final_loss = running_loss
         #logger.debug(f"Loss: {running_loss:.4f} at step {epoch + 1}")
         print(f"Loss: {running_loss:.4f} at step {epoch + 1}")
-        #torch.save(model.state_dict(), "model.pth")
+        #torch.save(model.state_dict(), "model_10.pth")
 
-    #logger.debug(f"Final loss: {final_loss:.4f}")
-    print(f"Final loss: {final_loss:.4f}")
     torch.save(model.state_dict(), "model.pth")  # save the final weights
+
+    # === Evaluation ===
+    model.eval()
+
+    metric = MeanAveragePrecision()
+    with torch.no_grad():
+        for images, targets in val_dataloader:
+            images = [img.to(device) for img in images]
+            targets = [{k: v.to(device) for k, v in t.items()} for t in targets]
+
+            outputs = model(images)
+
+            # Format predictions and targets for torchmetrics
+            preds = [{k: v.detach().cpu() for k, v in o.items()} for o in outputs]
[…]

### 2025-06-21 doc/article/article.tex

@@
 \subsubsection{Model}
-TODO \\
+Since it was assumed that classifying two types of diagrams - with a third 'other' class - was a quite simple task, a small network was deemed enough.
+The model consists of three convolutional blocks - each containing a convolutional layer, followed by a batch normalization and a max-pooling layer - then followed by three linear layers (for a more detailed overview, see Appendix~\ref{classification_net}).
+The output vector of the MLP contains three features - one for each class that this specific classifier needs to address.
+In order to determine the prediction, the softmax function is then applied.
+As later seen in section~\ref{exp:classifier}, a weighted Cross-Entropy loss is then used to train the model.
+\\
 
@@
 \subsection{Classifier}
+\label{exp:classifier}
 The first architecture explored for the Classifier module was a Convolutional Neural Network (CNN) following AlexNet's architecture, but smaller in size.
@@
 \subsection{Arrow Orientation}
-As we've seen in the previous section, the need to find the exact position of an arrow's head arises (see \textbf{Appendix A} for a full overview of the first approach).
+As we've seen in the previous section, the need to find the exact position of an arrow's head arises (see \textbf{Appendix~\ref{double_clustering}} for a full overview of the first approach).
 
@@
 \section{Double Clustering}
+\label{double_clustering}
 The first algorithmic approach proposed to solve this task - called \textit{Double Clustering} - was as follows: 
@@
 
+\section{The Classification Network}
+\label{classification_net}
+
+\begin{table}[h]
+	\centering
+	\caption{Classification Network Details}
+	\label{tab:classification-architecture}
+	\begin{tabular}{|c|c|c|c|}
+		\hline
+		\textbf{Conv Layer} & \textbf{Output Channels} & \textbf{Kernel Size} & \textbf{Pool Kernel Size} \\
+		\hline 
+		1 & 8 & 2 & 3 \\
+		2 & 16 & 2  & 3 \\
+		3 & 24 & 3  & 2 \\
+		\hline
+		\textbf{Linear Layer} & \textbf{Output Features} & \textbf{Droput Rate} & \textbf{Activation Function} \\
+		\hline 
+		1 & 2048 & 0.2 & ReLU \\
+		2 & 128 & 0.2 & ReLU \\
+		3 & 3 & 0.2 & ReLU \\
+		\hline
+	\end{tabular}
+\end{table}
+
 \end{appendices}

### 2025-06-22 src/classifier/classifier.py
Update paper

@@ -226,6 +226,6 @@ class GNRClassifier(Classifier):
     def get_default() -> 'GNRClassifier':
         return GNRClassifier(
             classes=recognizable_diagrams,
-            model_path="src/classifier/model_10.pth",
+            model_path="src/classifier/classifier.pth",
             processor=GNRMultiProcessor()
         )

### 2025-06-22 src/utils/bbox_utils.py
Added IoU func

@@ -233,3 +233,30 @@ def bbox_relative_position(first_bbox: ImageBoundingBox, second_bbox: ImageBound
             return "down"
         else:
             return "up"
+
+def IoU(first_bbox: ImageBoundingBox, second_bbox: ImageBoundingBox) -> float:
+    """
+    Computes the Intersection over Union (IoU) of two bounding boxes.
+    Args:
+        first_bbox (ImageBoundingBox): The first bounding box.
+        second_bbox (ImageBoundingBox): The second bounding box.
+    Returns:
+        float: The IoU value, which is the ratio of the area of intersection to the area of union.
+    """
+    x_left = max(first_bbox.top_left_x, second_bbox.top_left_x)
+    y_top = min(first_bbox.top_left_y, second_bbox.top_left_y)
+    x_right = min(first_bbox.bottom_right_x, second_bbox.bottom_right_x)
+    y_bottom = max(first_bbox.bottom_right_y, second_bbox.bottom_right_y)
+
+    inter_width = max(0.0, x_right - x_left)
+    inter_height = max(0.0, y_top - y_bottom)
+    intersection = inter_width * inter_height
+
+    area1 = (first_bbox.bottom_right_x - first_bbox.top_left_x) * (first_bbox.top_left_y - first_bbox.bottom_right_y)
+    area2 = (second_bbox.bottom_right_x - second_bbox.top_left_x) * (second_bbox.top_left_y - second_bbox.bottom_right_y)
+    union = area1 + area2 - intersection
+
+    if union <= 0:
+        return 0.0
+
+    return intersection / union

### 2025-06-22 src/extractor/arrow/arrow.py
Filled compute_arrows func

@@ -6,6 +6,7 @@ from matplotlib.image import BboxImage
 
 from core.image.bbox.bbox import ImageBoundingBox
 from core.image.bbox.bbox2p import ImageBoundingBox2Points
+from src.utils.bbox_utils import bbox_overlap, IoU
 
 
 @dataclass(frozen=True)
@@ -28,10 +29,58 @@ def compute_arrows(arrow_bboxes: List[ImageBoundingBox], head_bboxes: List[Image
     Returns: |arrow_bboxes| arrows
     """
 
-    # TODO
+    arrows: List[Arrow] = []
 
+    # Sort bboxes based on the number of overlaps with arrow bboxes
+    head_bboxes = sorted(
+        head_bboxes,
+        key=lambda bbox: sum([
+            1 if bbox_overlap(bbox, arrow_bbox) > 0 else 0 for arrow_bbox in arrow_bboxes
+        ])
+    )
+    tail_bboxes = sorted(
+        tail_bboxes,
+        key=lambda bbox: sum([
+            1 if bbox_overlap(bbox, arrow_bbox) > 0 else 0 for arrow_bbox in arrow_bboxes
+        ])
+    )
+    arrow_bboxes = sorted(
+        arrow_bboxes,
+        key=lambda bbox: len([
+            1 if bbox_overlap(bbox, head_bbox) > 0 else 0 for head_bbox in head_bboxes
+        ] + [
+            1 if bbox_overlap(bbox, tail_bbox) > 0 else 0 for tail_bbox in tail_bboxes
+        ])
+    )
+    
+    def get_most_certain_bbox(bboxes_part: List[ImageBoundingBox], arrow_bbox: ImageBoundingBox) -> ImageBoundingBox:
+        """
+        Get the most certain bbox from bboxes_part that overlaps with arrow_bbox.
+        """
+        max_overlap = 0
+        most_certain_bbox = None
+        
+        for bbox in bboxes_part:
+            overlap = bbox_overlap(bbox, arrow_bbox)
+            if overlap > max_overlap:
+                max_overlap = overlap
+                most_certain_bbox = bbox
+        
+        return most_certain_bbox
 
+    for arrow in arrow_bboxes:
+        head_bbox = get_most_certain_bbox(head_bboxes, arrow)
+        tail_bbox = get_most_certain_bbox(tail_bboxes, arrow)
 
+        if head_bbox is None or tail_bbox is None:
+            continue
 
[…]

### 2025-06-22 src/utils/bbox_utils.py
distance bbox/point func

@@ -260,3 +260,16 @@ def IoU(first_bbox: ImageBoundingBox, second_bbox: ImageBoundingBox) -> float:
         return 0.0
 
     return intersection / union
+
+def distance_bbox_point(bbox: ImageBoundingBox, point_x: float, point_y: float) -> float:
+    """
+    Calculates the shortest distance from a point to a bounding box.
+    Args:
+        bbox (ImageBoundingBox): The bounding box.
+        point (Tuple[float, float]): The point as a tuple of (x, y) coordinates.
+    Returns:
+        float: The shortest distance from the point to the bounding box.
+    """
+    x = max(bbox.top_left_x, min(point_x, bbox.bottom_right_x))
+    y = max(bbox.bottom_right_y, min(point_y, bbox.top_left_y))
+    return math.sqrt((x - point_x) ** 2 + (y - point_y) ** 2)

### 2025-06-22 src/extractor/flowchart/gnr_extractor.py
_compute_relations func

@@ -12,7 +12,7 @@ from src.extractor.arrow.arrow import Arrow
 from src.extractor.flowchart.multistage_extractor import MultistageFlowchartExtractor, ArrowTextTypeOutcome, \
     ElementTextTypeOutcome, ObjectRelation
 from src.flowchart_element_category import FlowchartElementCategory
-from src.utils.bbox_utils import bbox_overlap, bbox_distance, bbox_vertices, bbox_split, bbox_relative_position
+from src.utils.bbox_utils import bbox_overlap, bbox_distance, bbox_vertices, bbox_split, bbox_relative_position, distance_bbox_point
 from src.wellknown_diagram import WellKnownDiagram
 from src.extractor.text_extraction.text_extractor import TrOCRTextExtractorSmall
 
@@ -151,73 +151,35 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
                 element_bboxes (List[ImageBoundingBox]): A list of bounding boxes representing the elements in the diagram.
                 arrows (List[Arrow]): A list of arrows in the diagram.
             Returns:
-                ist[ObjectRelation]: A list of object relations, where each relation specifies the category of the arrow
+                List[ObjectRelation]: A list of object relations, where each relation specifies the category of the arrow
                                         and the indices of the source and target elements it connects.
-            Notes:
-                - The function calculates overlaps and distances between arrows and elements to determine relationships.
-                - Arrows are assumed to point in one of four directions: "right", "left", "up", or "down".
-                - Relationships are determined based on overlap scores, distance thresholds, and relative positions.
-                - If multiple overlaps exist, the function prioritizes overlaps with distinct relative positions.
-                - If only one overlap exists, the source and target indices may point to the same element.
         """
 
-        # TODO: si passa da arrow_bbox a 2 punti a testa e coda della freccia
-        # TODO: non ci sarà più right left, bisogna ottenere testa e coda di ogni freccia e
-
         ret: List[ObjectRelation] = []
-        for arrow in arrow_bboxes:
-            overlaps = []
-            arrow_pointing = "right"  # TODO: "right", "left", "up" or "down"
-            for idx, elem in enumerate(element_bboxes):
-                overlap_score = bbox_overlap(bbox1=elem, bbox2=arrow)
-                if overlap_score > self.element_arrow_overlap_threshold:
-                    relative_position = bbox_relative_position(first_bbox=arrow, second_bbox=elem)
[…]

### 2025-06-29 src/main.py
Dummy

@@ -238,7 +238,8 @@ if __name__ == '__main__':
 
 
     classifier = GNRClassifier("gnr-classifier", model_path=args.classifier)
-    text_digitizer = TrOCRTextExtractorBaseHandwritten()    # less expensive version, but with lower performance: TrOCRTextExtractorSmall()
+    text_digitizer = TrOCRTextExtractorBaseHandwritten()
+    # less expensive version, but with lower performance: TrOCRTextExtractorSmall()
 
     extractors = [
         GNRFlowchartExtractor(

### 2025-06-29 src/main.py
Update main.py

@@ -238,7 +238,7 @@ if __name__ == '__main__':
 
 
     classifier = GNRClassifier("gnr-classifier", model_path=args.classifier)
-    text_digitizer = TrOCRTextExtractorBaseHandwritten()
+    text_digitizer = TrOCRTextExtractorSmall()
     # less expensive version, but with lower performance: TrOCRTextExtractorSmall()
 
     extractors = [

### 2025-06-29 src/extractor/flowchart/gnr_extractor.py
Horror Refactor

@@ -185,26 +185,29 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
 
         for arrow in arrows:
 
-            target = source = None
+            target: Optional[ImageBoundingBox] = None
+            source: Optional[ImageBoundingBox] = None
+
             distance_head_min = distance_tail_min = None
+
             for i, element in enumerate(element_bboxes):
                 element_bbox: ImageBoundingBox = element
                 distance_element_head: float = distance_bbox_point(bbox=element_bbox, point_x=arrow.x_head, point_y=arrow.y_head)
                 if distance_element_head < self.element_arrow_distance_threshold and (distance_head_min is None or distance_element_head < distance_head_min):
-                    target = i
+                    target = element_bbox
                     distance_head_min = distance_element_head
 
                 distance_element_tail: float = distance_bbox_point(bbox=element_bbox, point_x=arrow.x_tail, point_y=arrow.y_tail)
                 if distance_element_tail < self.element_arrow_distance_threshold and (distance_tail_min is None or distance_element_tail < distance_tail_min):
-                    source = i
+                    source = element_bbox
                     distance_tail_min = distance_element_tail
 
             ret.append(
                 ObjectRelation(
                     category=Lookup.table_target_int_to_str_by_diagram_id[diagram_id][FlowchartElementCategoryIndex.ARROW.value],
                     arrow=arrow,
-                    source_index=source,
-                    target_index=target,
+                    source=source,
+                    target=target,
                 )
             )
 

### 2025-06-29 src/extractor/flowchart/multistage_extractor.py

@@ -19,8 +19,8 @@ logger = logging.getLogger(__name__)
 class ObjectRelation:
     category: str
     arrow: Arrow
-    source_index: Optional[int]
-    target_index: Optional[int]
+    source: Optional[ImageBoundingBox]
+    target: Optional[ImageBoundingBox]
 
 
 class ElementTextTypeOutcome(Enum):
@@ -97,9 +97,9 @@ class MultistageFlowchartExtractor(MultiStageExtractor, ABC):
 
         objects_relations: List[ObjectRelation] = self._compute_relations(diagram_id, element_bboxes, arrows)
 
-        elements: List[Element] = self._build_elements(diagram_id, image, elements_texts_associations)
+        elements, element_bboxes = self._build_elements(diagram_id, image, elements_texts_associations)
 
-        relations: List[Relation] = self._build_relations(diagram_id, image, objects_relations, arrows_texts_associations)
+        relations: List[Relation] = self._build_relations(diagram_id, image, objects_relations, element_bboxes, arrows_texts_associations)
 
         return FlowchartRepresentation(
             elements=elements,
@@ -184,7 +184,7 @@ class MultistageFlowchartExtractor(MultiStageExtractor, ABC):
         Return if text of arrow is inner, outer or must be discarded
         """
 
-    def _build_elements(self, diagram_id: str, image: Image, elements_texts_associations: Dict[ImageBoundingBox, List[ImageBoundingBox]]) -> List[Element]:
+    def _build_elements(self, diagram_id: str, image: Image, elements_texts_associations: Dict[ImageBoundingBox, List[ImageBoundingBox]]) -> Tuple[List[Element], List[ImageBoundingBox]]:
         """
         Constructs a list of Element objects by associating each detected diagram element with its related texts.
 
@@ -206,6 +206,7 @@ class MultistageFlowchartExtractor(MultiStageExtractor, ABC):
 
         bucket_of_discarded_texts: List[ImageBoundingBox] = []      # kept for future use
         elements: List[Element] = []
+        elements_bboxes: List[ImageBoundingBox] = []
         for element_bbox, associated_text_bboxes in elements_texts_associations.items():
             inner_text: List[str] = []
             outer_text: List[str] = []
@@ -230,11 +231,12 @@ class MultistageFlowchartExtractor(MultiStageExtractor, ABC):
 
             element: Element = Element(element_bbox.category, inner_text, outer_text)
             elements.append(element)
[…]

### 2025-06-29 src/representation/flowchart_representation/relation.py

@@ -19,8 +19,8 @@ class Relation:
     A class representing a relation in a flowchart representation.
     Attributes:
         category (str): The category of the relation.
-        source_id (Optional[int]): The ID of the source node in the relation.
-        target_id (Optional[int]): The ID of the target node in the relation.
+        source_index (Optional[int]): The ID of the source node in the relation.
+        target_index (Optional[int]): The ID of the target node in the relation.
         inner_text (List[str]): A list of strings representing the inner text of the relation.
         source_text (List[str]): A list of strings representing the text near the source node.
         target_text (List[str]): A list of strings representing the text near the target node.
@@ -36,8 +36,8 @@ class Relation:
     
 
     category: str
-    source_id: Optional[int]
-    target_id: Optional[int]
+    source_index: Optional[int]
+    target_index: Optional[int]
     inner_text: List[str]
     source_text: List[str]
     target_text: List[str]
@@ -46,8 +46,8 @@ class Relation:
     def to_dict(self) -> dict:
         return {
             "category": self.category,
-            "source_id": self.source_id,
-            "target_id": self.target_id,
+            "source_index": self.source_index,
+            "target_index": self.target_index,
             "inner_text": self.inner_text,
             "source_text": self.source_text,
             "target_text": self.target_text,
@@ -56,8 +56,8 @@ class Relation:
 
     def from_dict(self, data: dict) -> 'Relation':
         object.__setattr__(self, "category", data["category"])
-        object.__setattr__(self, "source_id", data["source_id"])
-        object.__setattr__(self, "target_id", data["target_id"])
+        object.__setattr__(self, "source_index", data["source_index"])
+        object.__setattr__(self, "target_index", data["target_index"])
         object.__setattr__(self, "inner_text", data["inner_text"])
         object.__setattr__(self, "source_text", data["source_text"])
         object.__setattr__(self, "target_text", data["target_text"])

### 2025-06-29 src/representation/flowchart_representation/test.py

@@ -16,9 +16,9 @@ class TestFlowchartRepresentation(unittest.TestCase):
                 Element(category="terminal", inner_text=["Begin"], outer_text=["End"]),
             ],
             relations=[
-                Relation(category="arrow", source_id="1", target_id="2", inner_text=[], source_text=[], target_text=[], middle_text=[]),
-                Relation(category="arrow", source_id="2", target_id="3", inner_text=[], source_text=["Ciao", "Due"], target_text=[], middle_text=[]),
-                Relation(category="arrow", source_id="1", target_id="5", inner_text=["inner"], source_text=[], target_text=[], middle_text=["Input"]),
+                Relation(category="arrow", source_index="1", target_index="2", inner_text=[], source_text=[], target_text=[], middle_text=[]),
+                Relation(category="arrow", source_index="2", target_index="3", inner_text=[], source_text=["Ciao", "Due"], target_text=[], middle_text=[]),
+                Relation(category="arrow", source_index="1", target_index="5", inner_text=["inner"], source_text=[], target_text=[], middle_text=["Input"]),
             ]
         )
         test_file = "test_flowchart.txt"

### 2025-06-29 src/transducer/d2/flowchart_transducer.py

@@ -100,9 +100,9 @@ class FlowchartToD2Transducer(Transducer):
 
         body += "\n"
         for relation in diagram_representation.relations:
-            if relation.source_id is None or relation.target_id is None:
+            if relation.source_index is None or relation.target_index is None:
                 continue
-            body += f"{relation.source_id}"
+            body += f"{relation.source_index}"
 
             relation_text = ""
             if len(relation.source_text) > 0:
@@ -116,7 +116,7 @@ class FlowchartToD2Transducer(Transducer):
 
             relation_text = relation_text.strip()
 
-            body += self.wrap_relation(relation.category, relation_text, relation.target_id)
+            body += self.wrap_relation(relation.category, relation_text, relation.target_index)
 
         outcome: TransducerOutcome = TransducerOutcome(diagram_id, WellKnownMarkupLanguage.D2_LANG.value, body)
         return outcome

### 2025-06-29 src/transducer/mermaid/flowchart_transducer.py

@@ -111,9 +111,9 @@ class FlowchartToMermaidTransducer(Transducer):
 
         body += "\n"
         for relation in diagram_representation.relations:
-            if relation.source_id is None or relation.target_id is None:
+            if relation.source_index is None or relation.target_index is None:
                 continue
-            body += f"\t{relation.source_id}"
+            body += f"\t{relation.source_index}"
 
             relation_text = ""
             if len(relation.source_text) > 0:
@@ -127,7 +127,7 @@ class FlowchartToMermaidTransducer(Transducer):
 
             relation_text = relation_text.strip()
 
-            body += f"{self.wrap_relation(relation.category, relation_text)}{relation.target_id}\n"
+            body += f"{self.wrap_relation(relation.category, relation_text)}{relation.target_index}\n"
 
         outcome: TransducerOutcome = TransducerOutcome(diagram_id=diagram_id, payload=body, markup_language="mermaid")
         return outcome

### 2025-06-29 doc/article/article.tex
Update paper
Edited paper
Edit paper

@@
 Our project proposes a modular architecture to tackle these tasks.
+
+\section{Previous Works}
+\cite{arrowrcnn} \\
+\cite{interactiveUML} \\
+\cite{sketchdiagram}
 
@@
 \subsection{Extractor Module}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\linewidth]{flowchart-extractor.png}
+	\caption{The complete pipeline of the extractor module.}
+	\label{fig:flowchart-extractor}
+\end{figure}
+
 \subsubsection{Preprocessing}
-TODO\\
+After some brief experiments, we found out that the object detection task through bounding boxes worked very well without any preprocessing technique.
+Moreover, the use of certain filters actually \textit{downgraded} the performances - e.g.
+the use of aggressive median filters reduced the capacity of the OCR model for text extraction.
+
+The only preprocessing applied is a gray-scale transformation.
+Other transformations are not deemed necessary - the model employed automatically resize the input image, so it's not needed to do it beforehand.
+\\
+
 \subsubsection{Bounding Box Detection}
-TODO\\
-\subsubsection{Content Recognition}
-After having determined all the bounding boxes of the diagram image, in order to correctly create the diagram internal representation, there's the need to link together text, elements and relations.\\
+\label{extr:bbd}
+TODO: FasterRCNN - backbone ResNet 50FPN~\cite{rcnn}.
+\\ After having determined all the bounding boxes of the diagram image, in order to correctly create the diagram internal representation, there's the need to link together text, elements and relations.\\
 
@@
 This task involves a few problems, among which determining exactly the arrow shape and its end - because sometimes other pen signs may be found in the bounding box of the arrow itself.
-[TODO: Dire l'approccio finale usato].
+After a few trials (and errors; for further insight, refer to section~\ref{exp:arrow_orientation}), the approach yielding the best results was through the use of the object detection network (see section~\ref{extr:bbd}) to recognize both arrow heads and tails.
 \\
@@
 After a text box is assigned to a relation, some computation is needed to fully understand how the text is associated to the relation itself.
-[TODO: Come trova source e target].
 The relation bounding box is then split into three different sections.
 The overlap between the text bounding box and each of the three sections boxes is computed - then, the section with the highest overlap value is then assigned the text under consideration.
+\\
+
+\subsubsection{Relationships Building}
+Once the text has been dealt with, the last step is to effectively compute which elements are connected by which arrow.
+This is done by finding, for each arrow, the elements nearer to the head (which we will assume to be the 'target' of the relation) and to the tail (the 'source' of the relationship exemplified by the arrow).
+These associations are nullified if the distance is above a certain adjustable threshold.
+This last step is done under the hypothesis that an arrow may come from nothing - or may point to nothing, having actually only either the source or the target.
+
+\section{Dataset}
+TODO: Elencare i dataset usati, in che modo erano usati originariamente e come lo usiamo noi (quali dati per quali componenti) \\
 
@@
 \subsection{Arrow Orientation}
-As we've seen in the previous section, the need to find the exact position of an arrow's head arises (see \textbf{Appendix~\ref{double_clustering}} for a full overview of the first approach).
+\label{exp:arrow_orientation}
+As we've seen in the previous section, the need to find the exact position of an arrow's head arises (see \textbf{Appendix~\ref{double_clustering}} for a full overview of the first approaches).
 
-TODO: Current approach.
+The solution, as stated before, was to
 
@@
 \bibliographystyle{IEEEtran}
-\bibliography{references}
+\bibliography{article}
 
@@
 
-However, this approach didn't work as expected (see Figure~\ref{fig:template_matching_fail}). [TODO: Spiegare perchè].
+However, this approach didn't work as expected (see Figure~\ref{fig:template_matching_fail}).
+In particular, it never worked around self arrows: the maximized distance inside one of the two clusters is never between the tail and the head of the arrow.
 
@@
 \end{figure}
+
+\section{Arrow-Net}
+\label{arrow-net}
 

### 2025-06-29 src/extractor/arrow/arrow.py
Update arrow.py

@@ -1,6 +1,6 @@
 import logging
 from dataclasses import dataclass
-from typing import List
+from typing import List, Tuple
 from scipy.spatial import ConvexHull
 from shapely.constructive import convex_hull
 from shapely.geometry import Polygon
@@ -169,13 +169,14 @@ def get_most_certain_bbox(bboxes_part: List[ImageBoundingBox], arrow_bbox: Image
 
     return most_certain_bbox
 
-def compute_arrows(arrow_bboxes: List[ImageBoundingBox], head_bboxes: List[ImageBoundingBox], tail_bboxes: List[ImageBoundingBox]) -> List[Arrow]:
+def compute_arrows(arrow_bboxes: List[ImageBoundingBox], head_bboxes: List[ImageBoundingBox], tail_bboxes: List[ImageBoundingBox]) \
+        -> Tuple[List[Arrow], List[ImageBoundingBox], List[ImageBoundingBox]]:
     """
     arrow_bboxes: bboxes of arrows
     arrow_bboxes: bboxes of arrow heads
     arrow_bboxes: bboxes of arrow tails
 
-    Returns: |arrow_bboxes| arrows
+    :returns: List of Arrow objects, remaining head bboxes, remaining tail bboxes
     """
 
     arrows: List[Arrow] = []
@@ -213,4 +214,4 @@ def compute_arrows(arrow_bboxes: List[ImageBoundingBox], head_bboxes: List[Image
 
         arrows.append(Arrow.from_bboxes(head_bbox=head_bbox, tail_bbox=tail_bbox, arrow=arrow))
 
-    return arrows
+    return arrows, head_bboxes, tail_bboxes

### 2025-06-29 src/compiler/mermaid/flowchart_compiler.py
Fix

@@ -1,5 +1,7 @@
 import base64
 import io, requests
+import logging
+
 import cv2
 import numpy as np
 from typing import List
@@ -12,7 +14,10 @@ class FlowchartToMermaidCompiler(Compiler):
         return [WellKnownMarkupLanguage.MERMAID.value]
 
     def compile(self, payload: str, output_path: str, dump_markuplang_file: bool = False, markuplang_file_path: str | None = None):
-        mm(payload, output_path)
+        try:
+            mm(payload, output_path)
+        except Exception as e:
+            logging.error(f"Mermaid can not be compiled due to error: {e}")
 
 
 def mm(graph: str, output_path: str):

### 2025-06-29 src/extractor/flowchart/multistage_extractor.py
Update multistage_extractor.py

@@ -70,7 +70,7 @@ class MultistageFlowchartExtractor(MultiStageExtractor, ABC):
         arrow_tail_bboxes: List[ImageBoundingBox] = [bbox for bbox in bboxes if self._is_arrow_tail_category(diagram_id, bbox.category)]
         text_bboxes: List[ImageBoundingBox] = [bbox for bbox in bboxes if self._is_text_category(diagram_id, bbox.category)]
 
-        arrows: List[Arrow] = compute_arrows(arrow_bboxes, arrow_head_bboxes, arrow_tail_bboxes)
+        arrows, head_bboxes, tail_bboxes = compute_arrows(arrow_bboxes, arrow_head_bboxes, arrow_tail_bboxes)
 
         arrow_bboxes_to_recover = []
         for arrow_bbox in arrow_bboxes:

### 2025-06-29 src/extractor/flowchart/gnr_extractor.py
Added _manage_unmatched_arrow_tails_and_heads

@@ -372,15 +372,22 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
 
             if head is not None and tail is not None:
                 managed_arrows.append(Arrow.from_bboxes(
-                head_bbox=ImageBoundingBox2Points.from_image(category=Lookup.table_target_int_to_str_by_diagram_id[diagram_id][FlowchartElementCategoryIndex.ARROW_HEAD.value],
-                                                box=head, trust=head_score, image=image),
-                tail_bbox=ImageBoundingBox2Points.from_image(category=Lookup.table_target_int_to_str_by_diagram_id[diagram_id][FlowchartElementCategoryIndex.ARROW_TAIL.value],
-                                                  box=tail, trust=tail_score, image=image)
-                                                  ))
-            else:
-                managed_arrows.append(None)
+                    head_bbox=ImageBoundingBox2Points.from_image(category=Lookup.table_target_int_to_str_by_diagram_id[diagram_id][FlowchartElementCategoryIndex.ARROW_HEAD.value],
+                                                    box=head, trust=head_score, image=image),
+                    tail_bbox=ImageBoundingBox2Points.from_image(
+                        category=Lookup.table_target_int_to_str_by_diagram_id[diagram_id][FlowchartElementCategoryIndex.ARROW_TAIL.value],
+                        box=tail,
+                        trust=tail_score,
+                        image=image
+                    ),
+                    arrow_bbox=arrow_bbox
+                ))
 
         return managed_arrows
+
+    @override
+    def _manage_unmatched_arrow_tails_and_heads(self, diagram_id: str, image: Image, head_bboxes: List[ImageBoundingBox], tail_bboxes: List[ImageBoundingBox]) -> List[Arrow]:
+        pass    # TODO: x Save
 
     @override
     def _extract_diagram_objects(self, diagram_id: str, image: Image) -> List[ImageBoundingBox]:

### 2025-06-29 src/extractor/flowchart/multistage_extractor.py

@@ -48,7 +48,7 @@ class MultistageFlowchartExtractor(MultiStageExtractor, ABC):
         else:
             return self.__seq_build_diagram_representation(diagram_id, image, bboxes)
 
-    def _manage_wrong_computed_arrows(self, diagram_id: str, image: Image, arrow_bboxes: List[ImageBoundingBox]) -> List[Optional[Arrow]]:
+    def _manage_wrong_computed_arrows(self, diagram_id: str, image: Image, arrow_bboxes: List[ImageBoundingBox]) -> List[Arrow]:
         """
 
         Args:
@@ -56,11 +56,26 @@ class MultistageFlowchartExtractor(MultiStageExtractor, ABC):
             image:
             arrow_bboxes:
 
-        Returns: in each position of the list return None if arrow is wrong computed again or a new Arrow
+        Returns: list of recovered arrows
 
         """
 
-        return [None for bbox in arrow_bboxes]
+        return []
+
+    def _manage_unmatched_arrow_tails_and_heads(self, diagram_id: str, image: Image, head_bboxes: List[ImageBoundingBox], tail_bboxes: List[ImageBoundingBox]) -> List[Arrow]:
+        """
+
+        Args:
+            diagram_id:
+            image:
+            head_bboxes:
+            tail_bboxes:
+
+        Returns: list of recovered arrows
+
+        """
+
+        return []
 
     def __seq_build_diagram_representation(self, diagram_id: str, image: Image, bboxes: List[ImageBoundingBox]) -> FlowchartRepresentation:
 
@@ -70,28 +85,16 @@ class MultistageFlowchartExtractor(MultiStageExtractor, ABC):
         arrow_tail_bboxes: List[ImageBoundingBox] = [bbox for bbox in bboxes if self._is_arrow_tail_category(diagram_id, bbox.category)]
         text_bboxes: List[ImageBoundingBox] = [bbox for bbox in bboxes if self._is_text_category(diagram_id, bbox.category)]
 
-        arrows: List[Arrow] = compute_arrows(arrow_bboxes, arrow_head_bboxes, arrow_tail_bboxes)
+        arrows, arrow_bboxes_to_recover, unmatched_head_bboxes, unmatched_tail_bboxes = compute_arrows(arrow_bboxes, arrow_head_bboxes, arrow_tail_bboxes)
 
-        arrow_bboxes_to_recover = []
-        for arrow_bbox in arrow_bboxes:
-            insert: bool = True
-            for arrow in arrows:
[…]

### 2025-06-29 src/classifier/preprocessing/processor.py
Fix

@@ -54,7 +54,10 @@ class GrayScaleProcessor(Processor):
             gray_image = cv2.cvtColor(image_np, cv2.COLOR_RGB2GRAY)
             assert gray_image.shape == image_np.shape[:2], f"Expected shape {image_np.shape[:2]}, but got {gray_image.shape}"
         
-        gray_image = 255 - gray_image
+        avg_brightness = gray_image.mean()
+        if avg_brightness < 127:
+            gray_image = 255 - gray_image
+
         tensor = torch.from_numpy(gray_image).to(device)
         return TensorImage(tensor)
 

### 2025-06-29 doc/article/article.tex
Updated paper

@@
 \subsubsection{Text Digitization}
-Since the task of converting an image of text into the relative string has already been solved in various ways, we settled upon the Microsoft \textit{Trocr Handwritten} Transformer model; more precisely, the \textit{base} version of it.
+Since the task of converting an image of text into the relative string has already been solved in various ways, we settled upon the Microsoft \textit{Trocr Handwritten} Transformer model; more precisely, the \textit{base} version of it \cite{microsofttrocr}.
 Every bounding box containing text that has been found is later passed to the model, and the resulting string generated by the model is linked to it.\\
@@
 
-The solution, as stated before, was to
+The solution, as stated before, was to finetune the already working RCNN in order for it to also extract the bounding boxes of both tails and heads of the arrow.
 
@@
 \bibliographystyle{IEEEtran}
+
+\newpage
 \bibliography{article}
+\newpage
 
@@
 	\item Suppose that the biggest cluster belongs to the arrow; the other clusters are thought to contain useless elements for this task, and can be discarded.
-	\item Find the keypoints of the arrow using SIFT.
+	\item Find the keypoints of the arrow using SIFT \cite{SIFT}.
 	\item Find the two keypoints inside the arrow with maximum distance. Assume that these two points are closely related to the two ends of the arrow.
@@
 
-For Step (1), we tried to employ both Spectral Clustering (see Figure~\ref{fig:spectral_clustering}) and DBSCan (see Figure~\ref{fig:arrow_double_clustering_clusters}).
+For Step (1), we tried to employ both Spectral Clustering (see Figure~\ref{fig:spectral_clustering}) and DBScan \cite{dbscan} (see Figure~\ref{fig:arrow_double_clustering_clusters}).
 
@@
 
-For Step (3), we used SIFT in order to find keypoints and some descriptors for them - necessary for the later template matching.
+For Step (3), we used SIFT \cite{SIFT} in order to find keypoints and some descriptors for them - necessary for the later template matching.
 During Step (4), we were able to correctly deduce the overall direction of the arrow (see Figure~\ref{fig:arrow_double_clustering_direction}).

### 2025-06-29 src/classifier/preprocessing/processor.py
Delicious refactor
Tasty, yummy

@@ -39,7 +39,8 @@ class GrayScaleProcessor(Processor):
 
         device = image.as_tensor().device
         image_np = image.as_tensor().cpu().detach().numpy()
-        image_np = (image_np * 255).clip(0, 255).astype(np.uint8)
+        if image_np.min() >= 0 and image_np.max() <= 1:
+            image_np = (image_np * 255).clip(0, 255).astype(np.uint8)
         
         if len(image_np.shape) != 3:
             return image

### 2025-06-29 doc/article/article.tex
Updated paper

@@
 
+\subsubsection{Relationships Building}
+Once the text has been digitized, it's necessary to effectively compute which elements are connected by which arrows.
+This is done by finding, for each arrow, the elements nearer to the head (which we will assume to be the 'target' of the relation) and to the tail (the 'source' of the relationship exemplified by the arrow).
+These associations are nullified if the distance is above a certain adjustable threshold.
+This last step is done under the hypothesis that an arrow may come from nothing - or may point to nothing, having actually only either the source or the target.
+\\
+
 \subsubsection{Text Association}
@@
 After a text box is assigned to a relation, some computation is needed to fully understand how the text is associated to the relation itself.
-The relation bounding box is then split into three different sections.
-The overlap between the text bounding box and each of the three sections boxes is computed - then, the section with the highest overlap value is then assigned the text under consideration.
-\\
-
-\subsubsection{Relationships Building}
-Once the text has been dealt with, the last step is to effectively compute which elements are connected by which arrow.
-This is done by finding, for each arrow, the elements nearer to the head (which we will assume to be the 'target' of the relation) and to the tail (the 'source' of the relationship exemplified by the arrow).
-These associations are nullified if the distance is above a certain adjustable threshold.
-This last step is done under the hypothesis that an arrow may come from nothing - or may point to nothing, having actually only either the source or the target.
+The relation's bounding box is then split into three different sections.
+The overlap between the text bounding box and each of the three sections boxes is computed - then, the section with the highest overlap value is then assigned to the text under consideration.
 
@@
 \label{exp:arrow_orientation}
-As we've seen in the previous section, the need to find the exact position of an arrow's head arises (see \textbf{Appendix~\ref{double_clustering}} for a full overview of the first approaches).
+As we've seen in the previous section, the need to find the exact position of an arrow's head arises (see \textbf{Appendix~\ref{double_clustering} and \textbf{Appendix}~\ref{arrow-net}} for a full overview of the first approaches).
 
@@
 
+This second approach for determining an arrow's orientation was inspired by UNet \cite{unet}.
+The core idea was to segment an arrow's bounding box, and to use a specific CNN to try and determine the arrow's head point and the arrow's tail point.
+
+\begin{table}[htbp]
+	\centering
+	\caption{ArrowNet Architecture Details}
+	\label{tab:arrownet}
+	\begin{tabular}{|l|l|l|l|l|}
+		\hline
+		\textbf{Stage} & \textbf{Layer} & \textbf{Input Shape} & \textbf{Output Shape} & \textbf{Operations} \\
+		\hline
+		\multirow{3}{*}{Encoder} 
+		& Encoder1 & $(B, 1, 256, 256)$ & $(B, 32, 256, 256)$ & Double Conv Block \\
+		& MaxPool & $(B, 32, 256, 256)$ & $(B, 32, 128, 128)$ & $2 \times 2$ MaxPooling \\
+		& Encoder2 & $(B, 32, 128, 128)$ & $(B, 64, 128, 128)$ & Double Conv Block \\
+		\hline
+		Bottleneck 
+		& MaxPool + Bottleneck & $(B, 64, 128, 128)$ & $(B, 128, 64, 64)$ & MaxPool + Double Conv \\
+		\hline
+		\multirow{4}{*}{Decoder}
+		& Upsampling2 & $(B, 128, 64, 64)$ & $(B, 64, 128, 128)$ & Transpose Conv Block \\
+		& Skip + Decoder2 & $(B, 128, 128, 128)$ & $(B, 64, 128, 128)$ & Concat + Double Conv \\
+		& Upsampling1 & $(B, 64, 128, 128)$ & $(B, 32, 256, 256)$ & Transpose Conv Block \\
+		& Skip + Decoder1 & $(B, 64, 256, 256)$ & $(B, 32, 256, 256)$ & Concat + Double Conv \\
+		\hline
+		Output & Output Conv & $(B, 32, 256, 256)$ & $(B, 3, 256, 256)$ & $1 \times 1$ Convolution \\
+		\hline
+	\end{tabular}
+\end{table}
+
 \section{The Classification Network}

### 2025-06-29 src/extractor/flowchart/gnr_extractor.py
Massive fix

@@ -16,7 +16,7 @@ from src.extractor.bbox_detection.target import FlowchartElementCategoryIndex, L
 from src.extractor.flowchart.multistage_extractor import MultistageFlowchartExtractor, ArrowTextTypeOutcome, \
     ElementTextTypeOutcome, ObjectRelation
 from src.utils.bbox_utils import bbox_overlap, \
-    distance_bbox_point, split_linestring_by_ratios, bbox_vertices, crop_image
+    distance_bbox_point, split_linestring_by_ratios, bbox_vertices, draw_predictions
 from src.wellknown_diagram import WellKnownDiagram
 from src.extractor.text_extraction.text_extractor import TextExtractor
 
@@ -44,6 +44,8 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
     arrow_text_inner_distance_threshold: float
     arrow_crop_delta_size_x: float
     arrow_crop_delta_size_y: float
+    arrow_head_overlap_threshold: float
+    arrow_tail_overlap_threshold: float
 
     ratios = [0.2, 0.6, 0.2]  # Source, Middle, Target
     bbox_trust_thresholds: Dict[int, Optional[float]] = field(default_factory=dict)
@@ -356,7 +358,7 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
                 delta_x_right -= orig_center_x + delta_x - W
 
             crop_bbox[:, (crop_center_y - delta_y_left):(crop_center_y + delta_y_right), (crop_center_x - delta_x_left):(crop_center_x + delta_x_right)] = \
-                image.as_tensor()[:, orig_y_start:orig_y_end, orig_x_start:orig_x_end]
+                image.as_tensor()[:, orig_y_start:orig_y_end, orig_x_start:orig_x_end] / 255.0
             prediction = self.bbox_detector(crop_bbox.unsqueeze(0))[0]
             head: Optional[torch.Tensor] = None
             tail: Optional[torch.Tensor] = None
@@ -383,11 +385,93 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
                     arrow_bbox=arrow_bbox
                 ))
 
+            draw_predictions(crop_bbox.unsqueeze(0), prediction)
+
         return managed_arrows
 
+    #"""
     @override
[…]

### 2025-06-29 src/main.py

@@ -161,14 +161,14 @@ def parse_args():
     parser.add_argument(
         "--arrow_crop_delta_size_x",
         type=float,
-        default=40.0,
+        default=30.0,
         help="arrow_crop_delta_size_x"
     )
 
     parser.add_argument(
         "--arrow_crop_delta_size_y",
         type=float,
-        default=25.0,
+        default=30.0,
         help="arrow_crop_delta_size_y"
     )
 
@@ -177,6 +177,20 @@ def parse_args():
         type=float,
         default=150,
         help="element_arrow_distance_threshold"
+    )
+
+    parser.add_argument(
+        "--arrow_head_overlap_threshold",
+        type=float,
+        default=0.0,
+        help="arrow_head_overlap_threshold"
+    )
+
+    parser.add_argument(
+        "--arrow_tail_overlap_threshold",
+        type=float,
+        default=0.0,
+        help="arrow_tail_overlap_threshold"
     )
 
     return parser.parse_args()
@@ -267,7 +281,9 @@ if __name__ == '__main__':
             arrow_text_inner_distance_threshold=args.arrow_text_inner_distance_threshold,
             arrow_crop_delta_size_x=args.arrow_crop_delta_size_x,
             arrow_crop_delta_size_y=args.arrow_crop_delta_size_y,
-            element_arrow_distance_threshold=args.element_arrow_distance_threshold
+            element_arrow_distance_threshold=args.element_arrow_distance_threshold,
+            arrow_head_overlap_threshold=args.arrow_head_overlap_threshold,
+            arrow_tail_overlap_threshold=args.arrow_tail_overlap_threshold
         )
     ]
 

### 2025-06-29 doc/article/article.tex

@@
 \usepackage{graphicx}
+\usepackage{multirow}
 \usepackage{amsmath, amssymb}
@@
 	\centering
-	\includegraphics[width=50pt]{templates.png}
+	\includegraphics[width=50pt, height=60pt]{templates.png}
 	\caption{An example of hand-picked template with SIFT descriptors.}
@@
 
-\section{Arrow-Net}
+\section{ArrowNet}
 \label{arrow-net}
@@
 The core idea was to segment an arrow's bounding box, and to use a specific CNN to try and determine the arrow's head point and the arrow's tail point.
+The downsampling-upsampling network is called \textbf{ArrowNet} and its specifics can be found in Table~\ref{table:arrownet}.
 
@@
 		\hline
-		\textbf{Stage} & \textbf{Layer} & \textbf{Input Shape} & \textbf{Output Shape} & \textbf{Operations} \\
+		\textbf{Stage} & \textbf{Layer} & \textbf{Out Channels} & \textbf{Operations} \\
 		\hline
 		\multirow{3}{*}{Encoder} 
-		& Encoder1 & $(B, 1, 256, 256)$ & $(B, 32, 256, 256)$ & Double Conv Block \\
-		& MaxPool & $(B, 32, 256, 256)$ & $(B, 32, 128, 128)$ & $2 \times 2$ MaxPooling \\
-		& Encoder2 & $(B, 32, 128, 128)$ & $(B, 64, 128, 128)$ & Double Conv Block \\
+		& Encoder1 & $32$ & Double Conv Block \\
+		& MaxPool & $32$ & $2 \times 2$ MaxPooling \\
+		& Encoder2 & $64$ & Double Conv Block \\
 		\hline
 		Bottleneck 
-		& MaxPool + Bottleneck & $(B, 64, 128, 128)$ & $(B, 128, 64, 64)$ & MaxPool + Double Conv \\
+		& MaxPool + Bottleneck & $128$ & MaxPool + Double Conv \\
 		\hline
 		\multirow{4}{*}{Decoder}
-		& Upsampling2 & $(B, 128, 64, 64)$ & $(B, 64, 128, 128)$ & Transpose Conv Block \\
-		& Skip + Decoder2 & $(B, 128, 128, 128)$ & $(B, 64, 128, 128)$ & Concat + Double Conv \\
-		& Upsampling1 & $(B, 64, 128, 128)$ & $(B, 32, 256, 256)$ & Transpose Conv Block \\
-		& Skip + Decoder1 & $(B, 64, 256, 256)$ & $(B, 32, 256, 256)$ & Concat + Double Conv \\
+		& Upsampling2 & $64$ & Transpose Conv Block \\
+		& Skip + Decoder2 & $64$ & Concat + Double Conv \\
+		& Upsampling1 & $32$ & Transpose Conv Block \\
+		& Skip + Decoder1 & $32$ & Concat + Double Conv \\
 		\hline
-		Output & Output Conv & $(B, 32, 256, 256)$ & $(B, 3, 256, 256)$ & $1 \times 1$ Convolution \\
+		Output & Output Conv & $3$ & $1 \times 1$ Convolution \\
 		\hline
 	\end{tabular}
+	\label{table:arrownet}
 \end{table}

### 2025-06-29 doc/article/article.tex
Updates

@@
 
-For Step (1), we tried to employ both Spectral Clustering (see Figure~\ref{fig:spectral_clustering}) and DBScan \cite{dbscan} (see Figure~\ref{fig:arrow_double_clustering_clusters}).
+For Step (1), we tried to employ both Spectral Clustering \cite{spectralclustering} (see Figure~\ref{fig:spectral_clustering}) and DBScan \cite{dbscan} (see Figure~\ref{fig:arrow_double_clustering_clusters}).
 
@@
 
+The entire architecture is composed of three main building blocks:
+
+\textbf{Double Convolution Block:} Each double convolution block applies two consecutive $3 \times 3$ convolutions with padding=1, followed by batch normalization and ReLU activation after each convolution.
+This preserves spatial dimensions while increasing feature depth.
+
+\textbf{Upsampling Block:} The upsampling blocks use $2 \times 2$ transposed convolutions with stride=2 to double the spatial resolution, followed by batch normalization and ReLU activation.
+
+\textbf{Skip Connections:} The architecture employs skip connections that concatenate feature maps from the encoder path with corresponding decoder features, doubling the channel dimension before processing through decoder blocks.
+
+The network processes input images of size $256 \times 256$ through a symmetric encoder-decoder structure.
+The encoder progressively reduces spatial resolution from $256^2$ to $128^2$ to $64^2$ while increasing feature channels from 32 to 64 to 128.
+The decoder reverses this process, upsampling back to the original resolution while maintaining rich feature representations through skip connections.
+The final $1 \times 1$ convolution maps the 32-channel feature representation to 3 output classes for pixel-wise classification.
+
 \section{The Classification Network}

### 2025-06-29 core/image/tensor_image.py

@@ -1,4 +1,4 @@
-from torchvision.io import read_image
+from torchvision.io import decode_image, ImageReadMode
 import cv2
 from core.image.image import Image
 from dataclasses import dataclass
@@ -17,7 +17,7 @@ class TensorImage(Image):
     def from_str(cls, path: str) -> Self:
 
         if (path.lower().endswith('.png') or path.lower().endswith('.jpeg') or path.lower().endswith('.jpg')):
-            tensor = read_image(path)
+            tensor = decode_image(path, mode = ImageReadMode.GRAY)
             if tensor.shape[0] == 4:
                 tensor = tensor[:3, :, :]
             return TensorImage(tensor)

### 2025-06-29 src/classifier/preprocessing/processor.py

@@ -41,6 +41,8 @@ class GrayScaleProcessor(Processor):
         image_np = image.as_tensor().cpu().detach().numpy()
         if image_np.min() >= 0 and image_np.max() <= 1:
             image_np = (image_np * 255).clip(0, 255).astype(np.uint8)
+        else:
+            image_np = image_np.clip(0, 255).astype(np.uint8)
         
         if len(image_np.shape) != 3:
             return image
@@ -54,7 +56,7 @@ class GrayScaleProcessor(Processor):
             image_np = np.transpose(image_np, (1, 2, 0))
             gray_image = cv2.cvtColor(image_np, cv2.COLOR_RGB2GRAY)
             assert gray_image.shape == image_np.shape[:2], f"Expected shape {image_np.shape[:2]}, but got {gray_image.shape}"
-        
+
         avg_brightness = gray_image.mean()
         if avg_brightness < 127:
             gray_image = 255 - gray_image

### 2025-06-29 core/image/tensor_image.py
Fix

@@ -1,11 +1,10 @@
-from torchvision.io import decode_image, ImageReadMode
-import cv2
 from core.image.image import Image
 from dataclasses import dataclass
 from torch import Tensor
 import torch
 from typing import Self
-from src import DEVICE
+from PIL import Image as PILImage
+import numpy as np
 
 
 @dataclass
@@ -16,19 +15,24 @@ class TensorImage(Image):
     @classmethod
     def from_str(cls, path: str) -> Self:
 
-        if (path.lower().endswith('.png') or path.lower().endswith('.jpeg') or path.lower().endswith('.jpg')):
-            tensor = decode_image(path, mode = ImageReadMode.GRAY)
-            if tensor.shape[0] == 4:
-                tensor = tensor[:3, :, :]
-            return TensorImage(tensor)
+        pil_img = PILImage.open(path).convert("RGBA")
+        np_img = np.array(pil_img)
 
-        if (path.lower().endswith('.bmp')):
-            np_img = cv2.imread(path)
-            if np_img.shape[2] == 4:
-                np_img = cv2.cvtColor(np_img, cv2.COLOR_BGRA2BGR)
-            assert np_img.shape[2] == 3, f"Expected 3D numpy array, but got {np_img.ndim}D array for image {path}"
-            tensor = torch.from_numpy(np_img).permute(2, 0, 1)
-            return TensorImage(tensor)
+        if np_img.shape[2] == 4:
+            alpha = np_img[:, :, 3:] / 255.0
+            rgb = np_img[:, :, :3].astype(np.float32)
+            white_bg = 255.0 * (1 - alpha)
+            rgb = rgb * alpha + white_bg
+            np_img = rgb.astype(np.uint8)
+        else:
+            np_img = np_img[:, :, :3]
+
+
+        if len(np_img.shape) == 2:
+            np_img = np.stack([np_img] * 3, axis=-1)
+
+        tensor = torch.from_numpy(np_img).permute(2, 0, 1).contiguous()  # C, H, W
+        return TensorImage(tensor)
             
 
     def as_tensor(self) -> Tensor:

### 2025-06-29 src/main.py

@@ -20,7 +20,8 @@ from src import DEVICE
 from src.classifier.classifier import GNRClassifier
 from src.extractor.bbox_detection import load_model
 from src.extractor.flowchart.gnr_extractor import GNRFlowchartExtractor
-from src.extractor.text_extraction.text_extractor import TrOCRTextExtractorSmall, TrOCRTextExtractorBaseHandwritten
+from src.extractor.text_extraction.text_extractor import TrOCRTextExtractorSmall, TrOCRTextExtractorBaseHandwritten, \
+    TrOCRTextExtractorLargeHandwritten
 from src.transducer.d2.flowchart_transducer import FlowchartToD2Transducer
 from src.transducer.mermaid.flowchart_transducer import FlowchartToMermaidTransducer
 from src.compiler.d2.flowchart_compiler import FlowchartToD2Compiler
@@ -253,7 +254,9 @@ if __name__ == '__main__':
 
     classifier = GNRClassifier("gnr-classifier", model_path=args.classifier)
     text_digitizer = TrOCRTextExtractorBaseHandwritten()
-    # less expensive version, but with lower performance: TrOCRTextExtractorSmall()
+    # less expensive version, but with lower performance:
+    # TrOCRTextExtractorBaseHandwritten()
+    # TrOCRTextExtractorSmall()
 
     extractors = [
         GNRFlowchartExtractor(

### 2025-06-29 src/transducer/mermaid/flowchart_transducer.py

@@ -1,6 +1,5 @@
-from typing import List, Type, override, Optional
-import sys, os
-
+from typing import List, Type, override
+from dataclasses import dataclass
 from src.representation.flowchart_representation.element import FlowchartElementCategory, Element
 from src.representation.flowchart_representation.relation import FlowchartRelationCategory, Relation
 
@@ -10,16 +9,14 @@ from src.representation.flowchart_representation.flowchart_representation import
 from core.transducer.outcome import TransducerOutcome
 from core.transducer.transducer import Transducer
 
+
+@dataclass
 class FlowchartToMermaidTransducer(Transducer):
     """
     Converts flowchart representations into Mermaid.js syntax.
     This transducer is specifically designed to handle flowchart diagrams and their elements, transforming them into
     a textual representation compatible with Mermaid.
     """
-    
-
-    def __init__(self, identifier: str):
-        super().__init__(identifier)
 
     @override
     def compatible_diagrams(self) -> List[str]:
@@ -36,10 +33,16 @@ class FlowchartToMermaidTransducer(Transducer):
         ]
 
     @staticmethod
+    def sanitize_text(text: str) -> str:
+        return text.replace(")", "").replace("(", "")
+
+    @staticmethod
     def wrap_element(category: str, label: str) -> str:
 
         if label.strip() == "":
             label = category
+
+        label = FlowchartToMermaidTransducer.sanitize_text(label)
 
         match category:
             case FlowchartElementCategory.CIRCLE.value:
@@ -59,6 +62,9 @@ class FlowchartToMermaidTransducer(Transducer):
 
     @staticmethod
     def wrap_relation(category: str, label: str) -> str:
+
+        label = FlowchartToMermaidTransducer.sanitize_text(label)
+
         match category:
             case FlowchartRelationCategory.ARROW.value:
                 if label.strip() == "":

### 2025-06-29 src/extractor/flowchart/gnr_extractor.py
Refactor

@@ -394,8 +394,8 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
                     ),
                     arrow_bbox=arrow_bbox
                 ))
-
-            draw_predictions(crop_bbox.unsqueeze(0), prediction)
+            if logging.root.level <= 10:
+                draw_predictions(crop_bbox.unsqueeze(0), prediction)
 
         return managed_arrows
 
@@ -481,8 +481,8 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
                         ),
                         arrow_bbox=arrow_bbox
                     ))
-
-                draw_predictions(crop_bbox.unsqueeze(0), prediction)
+                if logging.root.level <= 10:
+                    draw_predictions(crop_bbox.unsqueeze(0), prediction)
 
         return managed_arrows
 

### 2025-06-29 doc/article/article.tex
Update paper

--- 2025-06-29 doc/article/article.tex
@@
 \subsection{Object Detection}
-TODO
+
+TODO: Talk about the net.
+
+The model employed was tested in two different ways: as zero-shot, after a 10-epoch training and after a 100-epoch training.
+The best-behaving one was, of course, the latter.
+In Table~\ref{tab:map} and Table~\ref{tab:mar} we show the Mean Average Precision and Recall of the model used.
+
+\begin{table}[htbp]
+	\centering
+	\caption{Mean Average Precision (\textit{mAP}) Metrics}
+	\begin{tabular}{|l|c|l|}
+		\hline
+		\textbf{Metric} & \textbf{Value} & \textbf{Description} \\
+		\hline
+		mAP & 0.8754 & Overall mean Average Precision \\
+		mAP@0.5 & 0.9878 & mAP at IoU threshold 0.5 \\
+		mAP@0.75 & 0.9183 & mAP at IoU threshold 0.75 \\
+		\hline
+		mAP (Small) & 0.6444 & mAP for small objects \\
+		mAP (Medium) & 0.8054 & mAP for medium objects \\
+		mAP (Large) & 0.8975 & mAP for large objects \\
+		\hline
+	\end{tabular}
+	\label{tab:map}
+\end{table}
+
+\begin{table}[htbp]
+	\centering
+	\caption{Mean Average Recall (\textit{mAR}) Metrics}
+	\label{tab:mar}
+	\begin{tabular}{|l|c|l|}
+		\hline
+		\textbf{Metric} & \textbf{Value} & \textbf{Description} \\
+		\hline
+		mAR@1 & 0.3436 & mAR with max 1 detection per image \\
+		mAR@10 & 0.8877 & mAR with max 10 detections per image \\
+		mAR@100 & 0.9128 & mAR with max 100 detections per image \\
+		\hline
+		mAR (Small) & 0.6631 & mAR for small objects \\
+		mAR (Medium) & 0.8670 & mAR for medium objects \\
+		mAR (Large) & 0.9141 & mAR for large objects \\
+		\hline
+	\end{tabular}
+\end{table}
+
+Furthermore, we also think that the following metrics show a great precision:
+
+\begin{itemize}
+	\item \textbf{Average X-axis error:} 6.84 pixels
+	\item \textbf{Average Y-axis error:} 4.84 pixels
+\end{itemize}
 
@@
 \label{arrow-net}
-
 This second approach for determining an arrow's orientation was inspired by UNet \cite{unet}.
 The core idea was to segment an arrow's bounding box, and to use a specific CNN to try and determine the arrow's head point and the arrow's tail point.
-The downsampling-upsampling network is called \textbf{ArrowNet} and its specifics can be found in Table~\ref{table:arrownet}.
+This downsampling-upsampling network is called \textbf{ArrowNet} and its specifics can be found in Table~\ref{table:arrownet}.
 
@@
 		\hline
-		\textbf{Stage} & \textbf{Layer} & \textbf{Out Channels} & \textbf{Operations} \\
+		\textbf{Stage} & \textbf{Layer} & \textbf{Out Ch} & \textbf{Operations} \\
 		\hline
@@
 		Bottleneck 
-		& MaxPool + Bottleneck & $128$ & MaxPool + Double Conv \\
+		& MaxPool+Bottleneck & $128$ & MaxPool+Double Conv \\
 		\hline
@@
 		\hline
-		Output & Output Conv & $3$ & $1 \times 1$ Convolution \\
+		Output & Output Conv & $2$ & $1 \times 1$ Convolution \\
 		\hline
@@
 
+The entire architecture is composed of three main building blocks:
+
+\begin{enumerate}
+\item \textbf{Double Convolution Block:} Each double convolution block applies two consecutive $3 \times 3$ convolutions with padding 1, followed by batch normalization, and a ReLU activation after each convolution.
+This should preserve spatial dimensions while increasing feature depth.
+
+\item \textbf{Upsampling Block:} The upsampling blocks use $2 \times 2$ transposed convolutions with stride 2, followed by batch normalization and ReLU activation.
+
+\item \textbf{Skip Connections:} The architecture employs skip connections that concatenate feature maps from the encoder path with corresponding decoder features.
+
+\end{enumerate}
+
+The two output float channels of ArrowNet should have been used to predict, in a given position: $1$ if the head's point is in a certain position, $0$ else; $1$ if the tail's point is in a certain position, $0$ else - for the first and second channel respectively.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=225pt, height=75pt]{arrownet1.jpg}
+	\caption{Some results for arrow's head reconstruction.}
+	\label{fig:arrownet1}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=225pt, height=75pt]{arrownet2.jpg}
+	\caption{The overall result of an ArrowNet input.}
+	\label{fig:arrownet2}
+\end{figure}
+
+As seen in Figure~\ref{fig:arrownet2}, due to the low number of samples in the dataset, the network is unable to correctly distinguish between an arrow's head and tail.
+
 \section{The Classification Network}
 \label{classification_net}
+
+The \textbf{Classification Network} is a lightweight convolutional neural network, suited for processing grayscale images.
+The architecture is composed of convolutional feature extraction layers followed by fully connected classification layers.
+Some specific about it can be found in Table~\ref{tab:classification-architecture}.
 
@@
 		\hline
-		\textbf{Conv Layer} & \textbf{Output Channels} & \textbf{Kernel Size} & \textbf{Pool Kernel Size} \\
+		\textbf{Layer Type} & \textbf{Out Channels} & \textbf{Kernel Size} & \textbf{Pool Kernel Size} \\
 		\hline 
-		1 & 8 & 2 & 3 \\
-		2 & 16 & 2  & 3 \\
-		3 & 24 & 3  & 2 \\
+		Conv & 8 & 2 & 3 \\
+		Conv & 16 & 2  & 3 \\
+		Conv & 24 & 3  & 2 \\
 		\hline
-		\textbf{Linear Layer} & \textbf{Output Features} & \textbf{Droput Rate} & \textbf{Activation Function} \\
+		\textbf{Layer Type} & \textbf{Out Features} & \textbf{Droput Rate} & \textbf{Activation Function} \\
 		\hline 
-		1 & 2048 & 0.2 & ReLU \\
-		2 & 128 & 0.2 & ReLU \\
-		3 & 3 & 0.2 & ReLU \\
+		Linear & 2048 & 0.2 & ReLU \\
+		Linear & 128 & 0.2 & ReLU \\
+		Linear & 3 & 0.2 & ReLU \\
 		\hline

### 2025-06-29 doc/article/article.tex
Paper updated

@@
 
-\section{Previous Works}
+\section{Related Works}
 \cite{arrowrcnn} \\
@@
 \label{extr:bbd}
-TODO: FasterRCNN - backbone ResNet 50FPN~\cite{rcnn}.
-\\ After having determined all the bounding boxes of the diagram image, in order to correctly create the diagram internal representation, there's the need to link together text, elements and relations.\\
+The model employed for object detection is a FasterRCNN, relying on a - as backbone - ResNet 50FPN~\cite{rcnn}.
+The image can be passed rough, and the model automatically resize the image to its own resolution.
+\\
+After having determined all the bounding boxes of the diagram image, in order to correctly create the diagram internal representation, there's the need to link together text, elements and relations.
+For more details about the model and its training, refer to \textbf{Section}~\ref{sec:obj_det}.
+\\
 
@@
 \subsection{Object Detection}
-
+\label{sec:obj_det}
 TODO: Talk about the net.
@@
 The model employed was tested in two different ways: as zero-shot, after a 10-epoch training and after a 100-epoch training.
-The best-behaving one was, of course, the latter.
+The best-behaving one was, of course, the latter (to see the results of the different stages, see \textbf{Figures}~\ref{fig:bbox1} -~\ref{fig:bbox2} -~\ref{fig:bbox3}).
 In Table~\ref{tab:map} and Table~\ref{tab:mar} we show the Mean Average Precision and Recall of the model used.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=200pt, height=150pt]{bbox1.jpg}
+	\caption{The model employed in a zero-shot fashion.}
+	\label{fig:bbox1}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=200pt, height=150pt]{bbox2.jpg}
+	\caption{The model employed after 10 epochs of training.}
+	\label{fig:bbox2}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=180pt, height=150pt]{bbox3.png}
+	\caption{The model employed after 100 epochs of training. }
+	\label{fig:bbox3}
+\end{figure}
 

### 2025-06-29 core/orchestrator/orchestrator.py
Fixes

@@ -207,12 +207,14 @@ class Orchestrator(ToDeviceMixin):
 
         diagram_representations: List[DiagramRepresentation] = []
 
-        for extractor in compatible_extractors:
-            logger.info(f"Extract using {extractor.identifier}")
-            
-            image.to_device(extractor.get_device())
-            representation: DiagramRepresentation = extractor.extract(diagram_id, image)
-            diagram_representations.append(representation)
+        if len(compatible_extractors) > 0:
+            for extractor in compatible_extractors:
+                logger.info(f"Extract using {extractor.identifier}")
+
+                image.to_device(extractor.get_device())
+                representation: DiagramRepresentation = extractor.extract(diagram_id, image)
+                diagram_representations.append(representation)
+
         else:
             logger.warning(f"No compatible extractors are found for '{diagram_id}'")
 
@@ -300,12 +302,13 @@ class Orchestrator(ToDeviceMixin):
 
         with ProcessPoolExecutor() as executor:
             tasks: List[Future] = []
-            for extractor in compatible_extractors:
-                logger.debug(f"Extract using {extractor.identifier}")
+            if len(compatible_extractors) > 0:
+                for extractor in compatible_extractors:
+                    logger.debug(f"Extract using {extractor.identifier}")
 
-                tasks.append(
-                    executor.submit(self._par_extract_using_extractor_and_transduce, extractor, diagram_id, image)
-                )
+                    tasks.append(
+                        executor.submit(self._par_extract_using_extractor_and_transduce, extractor, diagram_id, image)
+                    )
             else:
                 logger.warning(f"No compatible extractors are found for '{diagram_id}'")
 

### 2025-06-29 src/extractor/bbox_detection/target.py

@@ -43,13 +43,6 @@ class Lookup:
         FlowchartElementCategoryIndex.ARROW_TAIL.value: FlowchartExtraElementCategory.ARROW_TAIL.value
     }
 
-    @property
-    def table_graph_category_str_to_target_int(self) -> Dict[str, int]:
-        return dict((v,k) for k,v in self.table_target_int_to_graph_category_str.items())
-
-    @property
-    def table_flowchart_category_str_to_target_int(self) -> Dict[str, int]:
-        return dict((v, k) for k, v in self.table_target_int_to_flowchart_category_str.items())
 
     table_target_int_to_str_by_diagram_id: Dict[str, Dict[int, str]] = {
         WellKnownDiagram.FLOW_CHART.value: table_target_int_to_flowchart_category_str,
@@ -57,6 +50,6 @@ class Lookup:
     }
 
     table_str_to_target_int_by_diagram_id: Dict[str, Dict[str, int]] = {
-        WellKnownDiagram.FLOW_CHART.value: table_flowchart_category_str_to_target_int,
-        WellKnownDiagram.GRAPH_DIAGRAM.value: table_graph_category_str_to_target_int,
+        WellKnownDiagram.GRAPH_DIAGRAM.value: dict((v, k) for k,v in table_target_int_to_graph_category_str.items()),
+        WellKnownDiagram.FLOW_CHART.value: dict((v, k) for k, v in table_target_int_to_flowchart_category_str.items()),
     }

### 2025-06-29 src/extractor/flowchart/gnr_extractor.py

@@ -322,10 +322,16 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
         Returns:
 
         """
+
+        logger.debug(f"manage_wrong_computed_arrows for {len(arrow_bboxes)} arrows")
+
         managed_arrows: List[Optional[Arrow]] = []
         crop_size: Tuple[int, int, int] = (1, image.as_tensor().shape[1] + int(self.arrow_crop_delta_size_y) + 1, image.as_tensor().shape[2] + int(self.arrow_crop_delta_size_x) + 1)
         _, H, W = image.as_tensor().shape
-        for arrow_bbox in arrow_bboxes:
+        for arrow_index, arrow_bbox in enumerate(arrow_bboxes):
+
+            logger.debug(f"bbox of arrow n. {arrow_index}: {arrow_bbox.box}")
+
             x1, y1, x2, y2 = int(arrow_bbox.bottom_left_x), int(arrow_bbox.top_left_y), int(arrow_bbox.bottom_right_x), int(arrow_bbox.bottom_left_y)
             crop_bbox: torch.Tensor = torch.ones(crop_size).to(self.get_device())
             crop_center_x: int = int(crop_bbox.shape[2] // 2)
@@ -382,6 +388,9 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
                     tail_score = score.item()
 
             if head is not None and tail is not None:
+
+                logger.debug(f"Both head and tail are found for n. {arrow_index}")
+
                 managed_arrows.append(Arrow.from_bboxes(
                     head_bbox=ImageBoundingBox2Points.from_image(category=Lookup.table_target_int_to_str_by_diagram_id[diagram_id][FlowchartElementCategoryIndex.ARROW_HEAD.value],
                                                     box=head, trust=head_score, image=image),
@@ -393,18 +402,27 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
                     ),
                     arrow_bbox=arrow_bbox
                 ))
+            else:
+                logger.debug(f"Head or tail not found for arrow n. {arrow_index}")
+
             if logging.root.level <= 10:
                 draw_predictions(crop_bbox.unsqueeze(0), prediction)
 
         return managed_arrows
 
-    #"""
     @override
     def _manage_unmatched_arrow_tails_and_heads(self, diagram_id: str, image: Image, head_bboxes: List[ImageBoundingBox], tail_bboxes: List[ImageBoundingBox]) -> List[Arrow]:
+
[…]

### 2025-06-29 src/extractor/flowchart/multistage_extractor.py

@@ -90,19 +90,21 @@ class MultistageFlowchartExtractor(MultiStageExtractor, ABC):
         recovered_arrows_from_arrow_bboxes = self._manage_wrong_computed_arrows(diagram_id, image, arrow_bboxes_to_recover)
         recovered_arrows_from_tail_and_head_bboxes = self._manage_unmatched_arrow_tails_and_heads(diagram_id, image, unmatched_head_bboxes, unmatched_tail_bboxes)
 
-        logger.debug(f"{len(recovered_arrows_from_arrow_bboxes)} recovered arrows from arrow bboxes")
-        arrows.extend(recovered_arrows_from_arrow_bboxes)
+        if len(recovered_arrows_from_arrow_bboxes) > 0:
+            logger.debug(f"{len(recovered_arrows_from_arrow_bboxes)} recovered arrows from arrow bboxes")
+            arrows.extend(recovered_arrows_from_arrow_bboxes)
 
-        logger.debug(f"{len(recovered_arrows_from_tail_and_head_bboxes)} recovered arrows from head and tail bboxes")
-        arrows.extend(recovered_arrows_from_tail_and_head_bboxes)
+        if len(recovered_arrows_from_tail_and_head_bboxes) > 0:
+            logger.debug(f"{len(recovered_arrows_from_tail_and_head_bboxes)} recovered arrows from head and tail bboxes")
+            arrows.extend(recovered_arrows_from_tail_and_head_bboxes)
 
         elements_texts_associations, arrows_texts_associations = self._compute_text_associations(diagram_id, element_bboxes, arrows, text_bboxes)
 
         objects_relations: List[ObjectRelation] = self._compute_relations(diagram_id, element_bboxes, arrows)
 
-        elements, element_bboxes = self._build_elements(diagram_id, image, elements_texts_associations)
+        elements, element_bboxes = self._build_elements(diagram_id, image, element_bboxes, elements_texts_associations)
 
-        relations: List[Relation] = self._build_relations(diagram_id, image, objects_relations, element_bboxes, arrows_texts_associations)
+        relations: List[Relation] = self._build_relations(diagram_id, image, objects_relations, element_bboxes, arrows_texts_associations, arrows)
 
         return FlowchartRepresentation(
             elements=elements,
@@ -187,7 +189,7 @@ class MultistageFlowchartExtractor(MultiStageExtractor, ABC):
         Return if text of arrow is inner, outer or must be discarded
         """
 
[…]

### 2025-06-29 src/main.py

@@ -43,7 +43,7 @@ def configure_logger(level_name: str):
     level = LOG_LEVELS.get(level_name.lower(), logging.INFO)
     logging.basicConfig(
         level=level,
-        format="[%(name)s] %(asctime)s [%(levelname)s] %(message)s",
+        format="%(asctime)s [%(levelname)s] %(message)s",   # [%(name)s]
         handlers=[logging.StreamHandler(sys.stdout)]
     )
 
@@ -88,7 +88,7 @@ def parse_args():
     )
 
     parser.add_argument(
-        "--log-level",
+        "--log",
         type=str,
         default="info",
         choices=LOG_LEVELS.keys(),
@@ -228,7 +228,7 @@ def main(device: str, classifier: Classifier, extractors: List[Extractor], trans
 
 if __name__ == '__main__':
     args = parse_args()
-    logger = configure_logger(args.log_level)
+    logger = configure_logger(args.log)
 
     # Validate compile output directory
     if args.then_compile:

### 2025-06-29 src/representation/flowchart_representation/element.py

@@ -1,5 +1,5 @@
 from enum import StrEnum
-from dataclasses import dataclass
+from dataclasses import dataclass, field
 from typing import List
 
 """
@@ -31,8 +31,8 @@ class Element:
     """
 
     category: str
-    inner_text: List[str]
-    outer_text: List[str]
+    inner_text: List[str] = field(default_factory=list)
+    outer_text: List[str] = field(default_factory=list)
 
     def to_dict(self) -> dict:
         return {

### 2025-06-29 src/representation/flowchart_representation/relation.py

@@ -1,5 +1,5 @@
 from enum import Enum
-from dataclasses import dataclass
+from dataclasses import dataclass, field
 from typing import List, Optional
 
 """
@@ -36,12 +36,12 @@ class Relation:
     
 
     category: str
-    source_index: Optional[int]
-    target_index: Optional[int]
-    inner_text: List[str]
-    source_text: List[str]
-    target_text: List[str]
-    middle_text: List[str]
+    source_index: Optional[int] = field(default=None)
+    target_index: Optional[int] = field(default=None)
+    inner_text: List[str] = field(default_factory=list)
+    source_text: List[str] = field(default_factory=list)
+    target_text: List[str] = field(default_factory=list)
+    middle_text: List[str] = field(default_factory=list)
     
     def to_dict(self) -> dict:
         return {

### 2025-06-29 src/classifier/classifier.py
Fixes

@@ -76,6 +76,9 @@ class GNRClassifier(Classifier):
         image = self.processor.process(image)
         tensor = image.as_tensor().to(self.device)
         y = self.model.forward(tensor.float())
+
+        logger.debug(f"Predictions: {y}")
+
         _, predicted = torch.max(y, dim=1)
         predicted = predicted.item()
 

### 2025-06-29 src/transducer/d2/flowchart_transducer.py

@@ -29,8 +29,11 @@ class FlowchartToD2Transducer(Transducer):
     def wrap_element(category: str, label: str, elem_id: int) -> str:
 
         s = str(elem_id)
-        if label.strip() != "":
-            s += f": {label}\n"
+
+        if label.strip() == "":
+            label = '""'
+
+        s += f": {label}\n"
 
         match category:
             case FlowchartElementCategory.CIRCLE.value:

### 2025-06-29 src/transducer/mermaid/flowchart_transducer.py

@@ -40,7 +40,7 @@ class FlowchartToMermaidTransducer(Transducer):
     def wrap_element(category: str, label: str) -> str:
 
         if label.strip() == "":
-            label = category
+            label = " "
 
         label = FlowchartToMermaidTransducer.sanitize_text(label)
 

### 2025-06-29 doc/article/article.tex
Updated paper

@@
 \section{Related Works}
-\cite{arrowrcnn} \\
-\cite{interactiveUML} \\
+The area of analyzing handwritten diagrams has been already studied in literature. Some key, modern, contributions are highlighted in this section.
+
+Various works have already focused on the development, training and evaluation of specific deep-learning networks for extracting structures in arrow-oriented diagrams (like \textbf{Arrow R-CNN} \cite{arrowrcnn}).
+More traditional approaches tipically focus on specific diagram types and a restricted set of elements in them (e.g., \textbf{UML} \cite{interactiveUML}).
+
+Some works in certain elements recognition - like arrows - are particularly domain-specific (like medical imaging \cite{med1}\cite{med2}).
+Other studies focus on the so-called \textit{online recognition} task, whose goal is to recognize diagrams as they are drawn (see \cite{online1}, \cite{online2}) - which could employ 'relative' strokes.
 \cite{sketchdiagram}

### 2025-06-29 doc/article/article.tex
Updated paper

--- 2025-06-29 doc/article/article.tex
@@
 \section{Related Works}
-\cite{arrowrcnn} \\
-\cite{interactiveUML} \\
-\cite{sketchdiagram}
+The area of analyzing handwritten diagrams has been already studied in literature. Some key, modern, contributions are highlighted in this section.
+
+Various works have already focused on the development, training and evaluation of specific deep-learning networks for extracting structures in arrow-oriented diagrams (like \textbf{Arrow R-CNN} \cite{arrowrcnn}) or in vector diagrams (as in \cite{sketchdiagram}).
+More traditional approaches tipically focus on specific diagram types and a restricted set of elements in them (e.g., \textbf{UML} \cite{interactiveUML}).
+
+Some works in element recognition - like arrows - are particularly domain-specific (like medical imaging \cite{med1}\cite{med2}).
+Other studies focus on the so-called \textit{online recognition} task, whose goal is to recognize diagrams as they are drawn (see \cite{online1}, \cite{online2}).
+The online recognition task quite often employs the use of relative strokes, and is mainly focused on flowchart diagrams.
 
@@
 The Compiler module is the final stage of the D.I.A.G.R.A.M. pipeline. 
-Its primary role is to take the structured representation of the diagram, expressed in a markup language (e.g., Mermaid.js), and generate a high-quality diagram in a visual format such as PNG.
+Its primary role is to take the structured representation of the diagram, expressed in a markup language (e.g., \textit{Mermaid.js}), and generate a high-quality diagram in a visual format such as PNG.
 
@@
 	\centering
-	\includegraphics[width=250pt]{classifier_preprocessing.png}
+	\includegraphics[width=220pt, height=150pt]{classifier_preprocessing.png}
 	\caption{
@@
 As later seen in section~\ref{exp:classifier}, a weighted Cross-Entropy loss is then used to train the model.
-\\
 
@@
 
-\begin{figure}[H]
+The Extractor - despite its name - doesn't only need to \textit{extract} objects from images, but also to find connections.
+Its pipeline can be seen at \textbf{Figure} ~\ref{fig:flowchart-extractor}.
+\\
+
+\begin{figure}
 	\centering
@@
 In Table~\ref{tab:map} and Table~\ref{tab:mar} we show the Mean Average Precision and Recall of the model used.
-\begin{figure}[H]
+
+\begin{figure}
 	\centering
@@
 
-\begin{figure}[H]
+\begin{figure}
 	\centering
@@
 
+\newpage
 \bibliographystyle{IEEEtran}
-
-\newpage
 \bibliography{article}
@@
 The architecture is composed of convolutional feature extraction layers followed by fully connected classification layers.
-Some specific about it can be found in Table~\ref{tab:classification-architecture}.
+Some specifics about it can be found in Table~\ref{tab:classification-architecture}.
 

### 2025-06-29 doc/article/article.tex
Updated paper

@@
 Some works in element recognition - like arrows - are particularly domain-specific (like medical imaging \cite{med1}\cite{med2}).
-Other studies focus on the so-called \textit{online recognition} task, whose goal is to recognize diagrams as they are drawn (see \cite{online1}, \cite{online2}).
+Other studies focus on the so-called \textit{online recognition} task, whose goal is to recognize diagrams as they are drawn (see \cite{online1}, \cite{online2}, \cite{online3}).
 The online recognition task quite often employs the use of relative strokes, and is mainly focused on flowchart diagrams.
@@
 \section{Dataset}
-TODO: Elencare i dataset usati, in che modo erano usati originariamente e come lo usiamo noi (quali dati per quali componenti) \\
+Our approach leverages multiple handwritten diagram datasets to ensure robustness across different diagram types and drawing styles.
+We employed a strategic dataset partitioning scheme where different subsets serve specific model training purposes.
+
+\subsection{Dataset Sources}
+
+We collected data from seven diverse sources, encompassing various diagram types commonly found in technical documentation and educational materials:
+
+\begin{enumerate}
+	\item \textbf{Finite Automata Dataset} \cite{online1}: Graph diagrams from the Czech Technical University collection, focusing exclusively on finite state machine representations.
+	\item \textbf{Flowchart Dataset (FCB)} \cite{online3}: Flowchart diagrams from the Czech Technical University repository.
+	\item \textbf{OHFCD\_1 Dataset} \cite{ohfcd}: Handwritten flowchart diagrams from the Computer Vision Center at UAB.
+	\item \textbf{hdBPMN Dataset} \cite{BPMN}: Business Process Model and Notation diagrams from the Data and Web Science Lab.
+	\item \textbf{Handwritten UML Class Diagrams} \cite{modelsketch}: Class diagram collection from Kaggle.
+	\item \textbf{CircuitNet Dataset} \cite{circuitnet}: Electronic circuit diagram repository.
+	\item \textbf{AI2D Dataset} \cite{ai2d}: Educational science diagrams from the Allen Institute for AI.
+\end{enumerate}
+
+\subsection{Dataset Utilization Strategy}
+
+We implemented a targeted approach for dataset utilization based on the specific requirements of our two-stage architecture:
+
+\textbf{Classifier Training:} To enhance model robustness and generalization capabilities, we employed class diagrams, circuit diagrams, and school diagrams from the UML, CircuitNet, and AI2D datasets respectively.
+This diverse selection ensures the classifier can effectively distinguish between different diagram types across various domains.
+
+\textbf{Extractor Training:} For the specialized extraction tasks, we focused exclusively on flowchart and graph diagrams from the FCB, OHFCD\_1, and Finite Automata datasets.
+This targeted approach allows the extraction models to achieve high precision on the specific structural elements relevant to these diagram types.
 

### 2025-06-29 core/orchestrator/test.py
Removed tests

@@ -1,5 +1,8 @@
 import unittest
-from typing import List, Type
+from typing import List, Type, Self
+
+import torch
+from torch import Tensor
 
 from core.classifier.classifier import Classifier
 from core.compiler.compiler import Compiler
@@ -19,6 +22,15 @@ FAKE_PAYLOAD_1 = "fake-payload-1"
 FAKE_PAYLOAD_2 = "fake-payload-2"
 
 class MockClassifier(Classifier):
+
+    def compatible_diagrams(self) -> List[str]:
+        pass
+
+    def to_device(self, device: str) -> None:
+        pass
+
+    def get_device(self) -> str:
+        pass
 
     n_calls = -1
     mock_classifications = [FAKE_DIAGRAM_TYPE_1, FAKE_DIAGRAM_TYPE_2]
@@ -46,6 +58,12 @@ class MockDiagramRepresentation2(DiagramRepresentation):
 
 class MockExtractor1(Extractor):
 
+    def get_device(self) -> str:
+        pass
+
+    def to_device(self, device: str) -> None:
+        pass
+
     def extract(self, diagram_id: str, image: Image) -> DiagramRepresentation:
         return MockDiagramRepresentation1()
 
@@ -64,6 +82,13 @@ class MockExtractor2(Extractor):
             FAKE_DIAGRAM_TYPE_2
         ]
 
+
+    def get_device(self) -> str:
+        pass
+
+    def to_device(self, device: str) -> None:
+        pass
+
 class MockExtractor3(Extractor):
 
     def extract(self, diagram_id: str, image: Image) -> DiagramRepresentation:
@@ -73,6 +98,13 @@ class MockExtractor3(Extractor):
         return [
             FAKE_DIAGRAM_TYPE_3
         ]
+
+
+    def get_device(self) -> str:
+        pass
+
+    def to_device(self, device: str) -> None:
+        pass
 
 
 class MockTransducer1(Transducer):
@@ -126,7 +158,19 @@ class MockCompiler2(Compiler):
         ]
 
 class MockImage(Image):
-    pass
+
+    @classmethod
+    def from_str(cls, path: str) -> Self:
+        pass
+
+    def as_tensor(self) -> Tensor:
+        return torch.tensor([])
+
+    def to_device(self, device: str) -> None:
+        pass
+
+    def get_device(self) -> str:
+        pass
 
 
 class TestOrchestrator(unittest.TestCase):
@@ -134,7 +178,7 @@ class TestOrchestrator(unittest.TestCase):
     @staticmethod
[…]

### 2025-06-29 src/classifier/preprocessing/test.py

@@ -1,11 +1,0 @@
-import os, sys
-sys.path.append(os.path.dirname("."))
-sys.path.append(os.path.dirname("/"))
-
-from src.classifier.preprocessing.processor import GNRMultiProcessor
-from core.image.tensor_image import TensorImage
-
-if __name__ == "__main__":
-    processor = GNRMultiProcessor()
-    image = TensorImage.from_str("test_resources/handwritten_diagram.jpg")
-    image = processor.process(image, debug=True)

### 2025-06-29 src/compiler/mermaid/test.py

@@ -1,56 +1,0 @@
-import unittest
-
-from core.representation.representation import DiagramRepresentation
-from src.compiler.mermaid.flowchart_compiler import FlowchartToMermaidCompiler
-from src.representation.flowchart_representation.flowchart_representation import FlowchartRepresentation, Element, Relation
-from src.representation.flowchart_representation.element import FlowchartElementCategory
-from src.representation.flowchart_representation.relation import FlowchartRelationCategory
-from src.transducer.mermaid.flowchart_transducer import FlowchartToMermaidTransducer
-
-
-class TestFlowchartToMermaidCompiler(unittest.TestCase):
-
-    def setUp(self):
-        self.transducer = FlowchartToMermaidTransducer('test_transducer')
-        self.compiler = FlowchartToMermaidCompiler('test_compiler')
-
-    def test_compile(self):
-        payload = """
-graph LR;
-    A--> B & C & D
-    B--> A & E
-    C--> A & E
-    D--> A & E
-    E--> B & C & D
-"""
-
-        payload2 = """---
-title: test_diagram
----
-graph 
-a(("Start_Node"))
-b["i++"]
-c{"if i > 5")
-a --->|int i = 0| b
-b -.-> c"""
-        # PAYLOAD2 IS THE ONE GIVING PROBLEMS
-        output_path = 'test_mermaid.png'
-        output_path2 = 'test_mermaid2.png'
-        markup_lang_file_path = 'test_mermaid.mmd'
-        markup_lang_file_path2 = 'test_mermaid.mmd'
-        self.compiler.compile(payload=payload, output_path=output_path, markuplang_file_path=markup_lang_file_path)
-        #self.compiler.compile(payload=payload2, output_path=output_path2, markuplang_file_path=markup_lang_file_path2)
-
-    def test_transducer_compiler(self):
-        diagram = FlowchartRepresentation(
-            elements=[Element(FlowchartElementCategory.CIRCLE.value, ["Ciao"], []), Element(FlowchartElementCategory.CIRCLE.value, ["Ciao2"], [])],
-            relations=[Relation(FlowchartRelationCategory.ARROW.value, "0", "1", [], [], [], [])]
-        )
-        output_path = 'test_mermaid3.png'
-        outcome = self.transducer.transduce("diagram", diagram)
-        markup_lang_file_path = 'test_mermaid.mmd'
-        self.compiler.compile(payload=outcome.payload, output_path=output_path, dump_markuplang_file=False)
-
-
-if __name__ == '__main__':
-    unittest.main()

### 2025-06-29 src/extractor/flowchart/test.py

@@ -1,125 +1,0 @@
-import unittest
-import torch
-import sys, os
-
-from src.extractor.bbox_detection.test import load_model
-
-sys.path.append(os.path.abspath(os.path.join(os.path.dirname(__file__), "../../../")))
-
-from core.image.bbox.bbox2p import ImageBoundingBox2Points
-from core.image.tensor_image import TensorImage
-from core.image.image import Image
-from src.extractor.flowchart.gnr_extractor import GNRFlowchartExtractor
-from src.extractor.flowchart.multistage_extractor import ElementTextTypeOutcome, ArrowTextTypeOutcome
-
-
-class TestGNRFlowchartExtractor(unittest.TestCase):
-
-    def setUp(self):
-        self.extractor = GNRFlowchartExtractor(identifier='test-extractor', bbox_detector=load_model())
-
-    def test_compute_text_associations(self):
-
-        self.extractor.element_precedent_over_arrow_in_text_association = True
-
-        element_bboxes = [
-            ImageBoundingBox2Points(
-                category="node1",
-                box=torch.tensor([0, 0, 10, 10], dtype=torch.float32),
-                trust=1
-            ),
-            ImageBoundingBox2Points(
-                category="node2",
-                box=torch.tensor([100, 100, 200, 200], dtype=torch.float32),
-                trust=1
-            ),
-        ]
-
-        arrow_bboxes = [
-            ImageBoundingBox2Points(
-                category="arrow1",
-                box=torch.tensor([50, 50, 60, 60], dtype=torch.float32),
-                trust=1
-            ),
-            ImageBoundingBox2Points(
-                category="arrow2",
-                box=torch.tensor([0, 0, 10, 10], dtype=torch.float32),
-                trust=1
-            ),
-        ]
-
-        text_bboxes = [
-            ImageBoundingBox2Points(
-                category="text-node1",
-                box=torch.tensor([0, 0, 8, 8], dtype=torch.float32),
-                trust=1
-            ),
-            ImageBoundingBox2Points(
-                category="text-node2",
-                box=torch.tensor([90, 90, 180, 180], dtype=torch.float32),
-                trust=1
-            ),
-            ImageBoundingBox2Points(
-                category="text-arrow1",
[…]

### 2025-06-29 src/representation/flowchart_representation/test.py

@@ -16,9 +16,9 @@ class TestFlowchartRepresentation(unittest.TestCase):
                 Element(category="terminal", inner_text=["Begin"], outer_text=["End"]),
             ],
             relations=[
-                Relation(category="arrow", source_index="1", target_index="2", inner_text=[], source_text=[], target_text=[], middle_text=[]),
-                Relation(category="arrow", source_index="2", target_index="3", inner_text=[], source_text=["Ciao", "Due"], target_text=[], middle_text=[]),
-                Relation(category="arrow", source_index="1", target_index="5", inner_text=["inner"], source_text=[], target_text=[], middle_text=["Input"]),
+                Relation(category="arrow", source_index=1, target_index=2, inner_text=[], source_text=[], target_text=[], middle_text=[]),
+                Relation(category="arrow", source_index=2, target_index=3, inner_text=[], source_text=["Ciao", "Due"], target_text=[], middle_text=[]),
+                Relation(category="arrow", source_index=1, target_index=5, inner_text=["inner"], source_text=[], target_text=[], middle_text=["Input"]),
             ]
         )
         test_file = "test_flowchart.txt"

### 2025-06-29 src/transducer/d2/test.py

@@ -1,82 +1,0 @@
-import unittest
-from typing import Dict, List
-
-from src.wellknown_markuplang import WellKnownMarkupLanguage
-from src.transducer.d2.flowchart_transducer import FlowchartToD2Transducer
-from src.wellknown_diagram import WellKnownDiagram
-from src.representation.flowchart_representation.element import FlowchartElementCategory, Element
-from src.representation.flowchart_representation.relation import FlowchartRelationCategory, Relation
-from src.representation.flowchart_representation.flowchart_representation import FlowchartRepresentation
-from core.transducer.outcome import TransducerOutcome
-
-class TestFlowchartToD2LTransducer(unittest.TestCase):
-
-    def setUp(self):
-        self.transducer = FlowchartToD2Transducer("test_transducer")
-
-    def test_compatible_diagrams(self):
-        compatible_diagrams = self.transducer.compatible_diagrams()
-        assert WellKnownDiagram.FLOW_CHART.value in compatible_diagrams, "Flowchart should be compatible"
-        assert WellKnownDiagram.GRAPH_DIAGRAM.value in compatible_diagrams, "Graph diagram should be compatible"
-
-    def test_wrap_element(self):
-        id_a, id_f = "A", "F"
-        result: str = self.transducer.wrap_element(FlowchartElementCategory.CIRCLE.value, "Test", id_a)
-        expected_result: str = "A: Test\nA.shape: circle\n"
-        self.assertEqual(result, expected_result, f"Expected\n{expected_result}, got\n{result}")
-        result: str = self.transducer.wrap_element(FlowchartElementCategory.TERMINAL.value, "Test", id_f)
-        expected_result: str = "F: Test\nF.shape: oval\n"
-        self.assertEqual(result, expected_result)
-
-    def test_wrap_relation(self):
-        id_d = "D"
-        result = self.transducer.wrap_relation(FlowchartRelationCategory.ARROW.value, "Test", id_d)
-        expected_result: str = "->D: Test\n"
-        self.assertEqual(result, expected_result)
-
-    def test_transduce(self):
-        id_a, id_b, id_c, id_d, id_e, id_f = "0", "1", "2", "3", "4", "5"
-
-        elements: list[Element] = [
-            Element(FlowchartElementCategory.CIRCLE.value, ["Circle"], []),
-            Element(FlowchartElementCategory.PROCESS.value, ["Process"], []),
[…]

### 2025-06-29 core/orchestrator/orchestrator.py
Added --markups flag

@@ -108,7 +108,8 @@ class Orchestrator(ToDeviceMixin):
         return compatible_transducer
 
 
-    def images2diagrams(self, images: List[Image], parallelization: bool = False, dump_markup: bool = False, then_compile: bool = True, outputs_dir_path: Optional[str] = None) -> List[TransducerOutcome]:
+    def images2diagrams(self, images: List[Image], parallelization: bool = False, dump_markup: bool = False,
+                        then_compile: bool = True, outputs_dir_path: Optional[str] = None, markups: Optional[List[str]] = None) -> List[TransducerOutcome]:
         """
         Convert a bulk of images
 
@@ -116,6 +117,7 @@ class Orchestrator(ToDeviceMixin):
         :param images: input images
         :param parallelization: enable parallelization
         :param then_compile: True if compile outcomes
+        :param markups: choose markup language
         :return: all transducer outcomes, based on extractors and transducers more than one outcome can be produced
         """
 
@@ -128,7 +130,7 @@ class Orchestrator(ToDeviceMixin):
                     logger.info(f"Elaborate image n. {index + 1}")
 
                     tasks.append(
-                        executor.submit(self.image2diagram, image, parallelization, dump_markup, then_compile, outputs_dir_path)
+                        executor.submit(self.image2diagram, image, parallelization, dump_markup, then_compile, outputs_dir_path, markups)
                     )
 
                 outcomes: List[TransducerOutcome] = []
@@ -140,12 +142,12 @@ class Orchestrator(ToDeviceMixin):
         else:
             for index, image in enumerate(images):
                 logger.info(f"Elaborate image n. {index + 1}")
-                outcomes.extend(self.image2diagram(image, parallelization, dump_markup=dump_markup, then_compile=then_compile, outputs_dir_path=outputs_dir_path))
+                outcomes.extend(self.image2diagram(image, parallelization, dump_markup=dump_markup, then_compile=then_compile, outputs_dir_path=outputs_dir_path, markups=markups))
 
 
         return outcomes
 
-    def image2diagram(self, image: Image, parallelization: bool = False, dump_markup: bool = False, then_compile: bool = True, outputs_dir_path: Optional[str] = None) -> List[TransducerOutcome]:
[…]

### 2025-06-29 src/main.py

@@ -194,10 +194,21 @@ def parse_args():
         help="arrow_tail_overlap_threshold"
     )
 
+    parser.add_argument(
+        "--markups",
+        type=str,
+        nargs="+",
+        default=None,
+        help="Pick markup languages"
+    )
+
     return parser.parse_args()
 
 
-def main(device: str, classifier: Classifier, extractors: List[Extractor], transducers: List[Transducer], compilers: List[Compiler], inputs_paths: List[str], parallelization: bool, then_compile: bool, outputs_dir_path: Optional[str]):
+def main(device: str, classifier: Classifier, extractors: List[Extractor], transducers: List[Transducer],
+         compilers: List[Compiler], inputs_paths: List[str], parallelization: bool, then_compile: bool,
+         outputs_dir_path: Optional[str], markups: Optional[List[str]] = None):
+
     images: List[Image] = []
     for path in inputs_paths:
         image = TensorImage.from_str(path)
@@ -222,7 +233,8 @@ def main(device: str, classifier: Classifier, extractors: List[Extractor], trans
         parallelization=parallelization,
         dump_markup=True,       # CLI always dumps markup files
         then_compile=then_compile,
-        outputs_dir_path=outputs_dir_path
+        outputs_dir_path=outputs_dir_path,
+        markups=markups
     )
 
 
@@ -286,7 +298,7 @@ if __name__ == '__main__':
             arrow_crop_delta_size_y=args.arrow_crop_delta_size_y,
             element_arrow_distance_threshold=args.element_arrow_distance_threshold,
             arrow_head_overlap_threshold=args.arrow_head_overlap_threshold,
-            arrow_tail_overlap_threshold=args.arrow_tail_overlap_threshold
+            arrow_tail_overlap_threshold=args.arrow_tail_overlap_threshold,
         )
     ]
 
@@ -305,5 +317,6 @@ if __name__ == '__main__':
         compilers=[
             FlowchartToMermaidCompiler("flowchart-to-mermaid-compiler"),
             FlowchartToD2Compiler("flowchart-to-d2-compiler")
-        ]
+        ],
+        markups=args.markups
     )

### 2025-06-29 doc/article/article.tex

@@
 \usepackage{hyperref}
+\usepackage{subcaption}
 \usepackage{caption}
@@
 
-\subsection{Dataset Utilization Strategy}
+Many datasets don't include the task at hand - which is focused on recognizing flowchart and graph diagrams - but could be in future works.
+Moreover, many datasets not apt to our task were used to further enhance the capability of the classifier - letting it know if an uploaded image couldn't be further processed.
 
-We implemented a targeted approach for dataset utilization based on the specific requirements of our two-stage architecture:
+\subsection{Dataset Usage Strategy}
+
+We implemented a targeted approach for dataset utilization based on the specific requirements of our architecture:
 
@@
 
-\textbf{Extractor Training:} For the specialized extraction tasks, we focused exclusively on flowchart and graph diagrams from the FCB, OHFCD\_1, and Finite Automata datasets.
-This targeted approach allows the extraction models to achieve high precision on the specific structural elements relevant to these diagram types.
+\textbf{Extractor Training:} For the specialized extraction tasks, we focused exclusively on the first three datasets: flowchart and graph diagrams from the FCB, OHFCD\_1, and Finite Automata datasets.
+This targeted approach allows the extraction models to achieve high precision on the specific structural elements relevant to these diagram types, which were the core of our task.
 
@@
 \label{sec:obj_det}
-TODO: Talk about the net.
+
+For bounding box detection of diagram elements, we employed a Faster R-CNN architecture \cite{rcnn} with a ResNet-50 Feature Pyramid Network (FPN) backbone \cite{lin2017feature}.
+This two-stage object detection framework provides robust performance for multi-class detection tasks while maintaining computational efficiency suitable for diagram analysis applications.
+Its main components are:
+
+\textbf{Backbone Network:} The ResNet-50 backbone serves as the feature extraction component, providing representations through its four residual blocks.
+
+\textbf{Feature Pyramid Network:} The FPN enhancement augments the ResNet-50 backbone features by creating a top-down pathway with connections.
+
+\textbf{Region Proposal Network (RPN):} The RPN generates object proposals by sliding a small network over the feature maps from the FPN.
+At each spatial location, the RPN predicts object presence probability and bounding box refinements for different scales and aspect ratios.
+
+\textbf{ROI Head:} The second stage performs precise classification and bounding box regression on the proposals generated by the RPN. Feature vectors are extracted using ROI Align operations. \\
+
+The original model was trained over the COCO dataset \cite{coco}, which comprised 91 classes.
+It should be highlighted that the original training dataset was not in any way related to diagrams: with an efficient finetune, the network was able to swiftly change to our task.
+\\
 
@@
 \section{Discussion}
-TODO
+
+The development of D.I.A.G.R.A.M.
+offers an insightful perspective on the integration of deterministic logic and deep learning for the interpretation of handwritten diagrams.
+One of the most valuable outcomes of this project is the successful design of a complete and functional pipeline (see Figure~\ref{fig:results}), which transforms raw visual input into a structured, editable format.
+This end-to-end workflow—from classification to rendering—highlights the feasibility of modular architectures for diagram understanding.
+
+\begin{figure}[htbp]
+	\centering
+
+	% Row 1
+	\begin{subfigure}[b]{0.45\linewidth}
+		\includegraphics[width=\linewidth]{ex1.png}
+		\caption{An easy flowchart diagram.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.4\linewidth}
+		\includegraphics[width=40pt, height=150pt]{ex2.png}
+		\caption{The network's output.}
+	\end{subfigure}
+
+	\vspace{1em} % spacing between rows
+
+	% Row 2
+	\begin{subfigure}[b]{0.45\linewidth}
+		\includegraphics[width=\linewidth]{ex3.png}
+		\caption{A complex graph diagram.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.45\linewidth}
+		\includegraphics[width=80pt, height=150pt]{ex4.png}
+		\caption{The network's output.}
+	\end{subfigure}
+
+	\vspace{1em}
+
+	% Row 3
+	\begin{subfigure}[b]{0.45\linewidth}
+		\includegraphics[width=\linewidth]{ex5.png}
+		\caption{A complex flowchart diagram.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.45\linewidth}
+		\includegraphics[width=\linewidth]{ex6.png}
+		\caption{The network's output.}
+	\end{subfigure}
+
+	\caption{A few results of our network.}
+	\label{fig:results}
+\end{figure}
+
+Among the modules, the Extractor proved particularly effective.
+Leveraging deep neural networks, it reliably identified diagram elements across diverse samples, validating the suitability of data-driven approaches for this subtask.
+Conversely, the Classifier, while operational, showed room for improvement, particularly in edge cases or less represented diagram categories.
+This suggests that the current training set may lack the diversity required for robust generalization, pointing to data augmentation and dataset expansion as immediate avenues for refinement.
+
+The system as a whole adheres to principles of scalability and generality.
+By decoupling the diagram type detection from the extraction and compilation logic, D.I.A.G.R.A.M.
+maintains flexibility in handling new diagram formats with minimal architectural changes.
+However, its performance remains sensitive to input quality and assumes a level of neatness in the handwritten diagrams that may not always be guaranteed in real-world scenarios.
+
+In comparison to related work, D.I.A.G.R.A.M.
+distinguishes itself by targeting offline, image-based recognition (as opposed to stroke-based online methods) while retaining support for structural output generation.
+This positions it as a hybrid between traditional rule-based recognizers and modern neural approaches.
 

### 2025-06-29 src/classifier/classifier.py
Fixes

@@ -70,6 +70,8 @@ class GNRClassifier(Classifier):
         :param image: The image to classify.
         :return: The classification result.
         """
+
+        return WellKnownDiagram.GRAPH_DIAGRAM.value     # TODO
         
         self.model.eval()
         self.model = self.model.to(self.device)

### 2025-06-29 src/extractor/flowchart/gnr_extractor.py

@@ -468,18 +468,30 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
                 arrow_score: float = 0.0
                 for bbox, label, score in zip(prediction['boxes'], prediction['labels'], prediction['scores']):
 
-                    filtered_bbox: List[ImageBoundingBox] = self._filter_bboxes(diagram_id=diagram_id,
-                    bboxes=[ImageBoundingBox2Points.from_image(category=Lookup.table_target_int_to_str_by_diagram_id[diagram_id][label.item()],
-                            box=bbox, trust=score.item(), image=image)])
+                    if label.item() not in Lookup.table_target_int_to_str_by_diagram_id[diagram_id].keys():
+                        continue
+
+                    filtered_bbox: List[ImageBoundingBox] = self._filter_bboxes(
+                        diagram_id=diagram_id,
+                        bboxes=[
+                            ImageBoundingBox2Points.from_image(
+                                category=Lookup.table_target_int_to_str_by_diagram_id[diagram_id][label.item()],
+                                box=bbox,
+                                trust=score.item(),
+                                image=image
+                            )
+                        ]
+                    )
+
                     if label.item() == FlowchartElementCategoryIndex.ARROW.value and score.item() > arrow_score and len(filtered_bbox) == 1:
-                        arrow_bbox = ImageBoundingBox2Points.from_image(
+
+                        candidate_arrow_bbox = ImageBoundingBox2Points.from_image(
                             category=Lookup.table_target_int_to_str_by_diagram_id[diagram_id][
                                 FlowchartElementCategoryIndex.ARROW.value],
                             box=bbox, trust=score.item(), image=image)
-                        if bbox_overlap(arrow_bbox, head_bbox) > self.arrow_head_overlap_threshold and bbox_overlap(arrow_bbox, tail_bbox) > self.arrow_tail_overlap_threshold:
+                        if bbox_overlap(candidate_arrow_bbox, head_bbox) > self.arrow_head_overlap_threshold and bbox_overlap(candidate_arrow_bbox, tail_bbox) > self.arrow_tail_overlap_threshold:
                             arrow_score = score.item()
-                        else:
-                            arrow_bbox = None
+                            arrow_bbox = candidate_arrow_bbox
 
                 if arrow_bbox is not None:
                     managed_arrows.append(Arrow.from_bboxes(
[…]

### 2025-06-29 doc/article/article.tex

@@
 Then, a softmax activation function was applied to the output layer, in order to obtain the probabilities of each class.
-The model was trained on our own dataset, with only three possible output classes: 'flowchart', 'graph' and 'other'; it was trained for 10 epochs, with an AdamW optimizer (with a learning rate of 1e-3) and a standard CrossEntropy Loss function.
+The model was trained on our own dataset, with only three possible output classes: 'flowchart', 'graph' and 'other'; it was trained for 10 epochs, with an AdamW optimizer (with a learning rate of $0.001$) and a standard CrossEntropy Loss function.
 \\
@@
 Judging this to be a problem in architecture size, we then tried adding another CNN layer with 16 filters; this one too followed by a batch normalization and a max pooling layer.
-The rest of the net was untouched; we also decreased the learning rate to 1e-4, in order to avoid overshooting the minimum of the loss function.
+The rest of the net was untouched; we also decreased the learning rate to $0.0001$, in order to avoid overshooting the minimum of the loss function.
 The results, however, didn't show any kind of improvement, with the loss function still being around 0.9.
@@
 It should be highlighted that the original training dataset was not in any way related to diagrams: with an efficient finetune, the network was able to swiftly change to our task.
-\\
+The finetuning is done with a learning rate of $0.005$ \\
 
-The model employed was tested in two different ways: as zero-shot, after a 10-epoch training and after a 100-epoch training.
-The best-behaving one was, of course, the latter (to see the results of the different stages, see \textbf{Figures}~\ref{fig:bbox1} -~\ref{fig:bbox2} -~\ref{fig:bbox3}).
+The model employed was tested in three different ways: as zero-shot, after a 10-epoch training and after a 100-epoch training.
+The best-behaving one was, of course, the latter (to see the results of the different stages, see \textbf{Figures}~\ref{fig:bbox1} -~\ref{fig:bbox2} -~\ref{fig:bbox3} -~\ref{fig:fc}).
 In Table~\ref{tab:map} and Table~\ref{tab:mar} we show the Mean Average Precision and Recall of the model used.
 
-\begin{figure}
+\begin{figure}[H]
 	\centering
@@
 
-\begin{figure}
+\begin{figure}[H]
 	\centering
@@
 
-\begin{figure}
+\begin{figure}[H]
 	\centering
 	\includegraphics[width=180pt, height=150pt]{bbox3.png}
-	\caption{The model employed after 100 epochs of training. }
+	\caption{The model employed after 100 epochs of training. More bounding boxes can be seen because we also added arrow's tails and heads.}
 	\label{fig:bbox3}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=180pt, height=150pt]{fc.png}
+	\caption{The bounding boxes of a flowchart diagram, on the 100-epoch model.}
+	\label{fig:fc}
 \end{figure}
@@
 
-The solution, as stated before, was to finetune the already working RCNN in order for it to also extract the bounding boxes of both tails and heads of the arrow.
+The solution, as stated before, was to finetune the already working Faster R-CNN in order for it to also extract the bounding boxes of both tails and heads of the arrow.
 
@@
 This positions it as a hybrid between traditional rule-based recognizers and modern neural approaches.
+
+TODO: Nodo extra
 

### 2025-06-29 core/orchestrator/orchestrator.py
Fixes

@@ -190,7 +190,8 @@ class Orchestrator(ToDeviceMixin):
 
             outcomes.extend(o)
 
-        outcomes = [outcome for outcome in outcomes if outcome.markup_language in markups]
+        if markups is not None:
+            outcomes = [outcome for outcome in outcomes if outcome.markup_language in markups]
 
         if dump_markup:
             self.__seq_dump_markup(outcomes, outputs_dir_path)
@@ -288,7 +289,8 @@ class Orchestrator(ToDeviceMixin):
 
         outcomes = self.__par_elaboration(diagram_id, image)
 
-        outcomes = [outcome for outcome in outcomes if outcome.markup_language in markups]
+        if markups is not None:
+            outcomes = [outcome for outcome in outcomes if outcome.markup_language in markups]
 
         if dump_markup:
             self.__seq_dump_markup(outcomes, outputs_dir_path)

### 2025-06-29 src/classifier/classifier.py

@@ -70,8 +70,6 @@ class GNRClassifier(Classifier):
         :param image: The image to classify.
         :return: The classification result.
         """
-
-        return WellKnownDiagram.GRAPH_DIAGRAM.value     # TODO
         
         self.model.eval()
         self.model = self.model.to(self.device)

### 2025-07-01 doc/article/article.tex
Updated paper

@@
 
-Some works in element recognition - like arrows - are particularly domain-specific (like medical imaging \cite{med1}\cite{med2}).
-Other studies focus on the so-called \textit{online recognition} task, whose goal is to recognize diagrams as they are drawn (see \cite{online1}, \cite{online2}, \cite{online3}).
+Some works in element recognition - like arrows - are particularly domain-specific (like medical imaging \cite{med1}\cite{med2}, or in circuit diagrams \cite{2024modular}).
+Other studies focus on the so-called \textit{online recognition} task, whose goal is to recognize diagrams as they are drawn (see \cite{online1}, \cite{online2}, \cite{online3} and especially \cite{framework} for a general framework).
 The online recognition task quite often employs the use of relative strokes, and is mainly focused on flowchart diagrams.
@@
 	\hfill
-	\begin{subfigure}[b]{0.4\linewidth}
-		\includegraphics[width=40pt, height=150pt]{ex2.png}
+	\begin{subfigure}[b]{0.45\linewidth}
+		\includegraphics[width=35pt, height=150pt]{ex2.png}
 		\caption{The network's output.}
@@
 	\begin{subfigure}[b]{0.45\linewidth}
-		\includegraphics[width=80pt, height=150pt]{ex4.png}
+		\includegraphics[width=50pt, height=130pt]{ex4.png}
 		\caption{The network's output.}
@@
 	\begin{subfigure}[b]{0.45\linewidth}
-		\includegraphics[width=\linewidth]{ex6.png}
+		\includegraphics[width=70pt, height=130pt]{ex6.png}
 		\caption{The network's output.}
@@
 
-TODO: Nodo extra
+We'd like to point out a few things about Figure~\ref{fig:results}.
+First of all, the result shown in part (d).
+A non-existing node can be seen in the upper right corner of the image; this is due to the fact that the self-arrow of node $q_5$ is recognized - by the object detection network - both as an arrow and as a node, with an high trust value.
+This problem could have been solved by applying a Non-Maxima Suppression - but in this specific case, the highest confidence value was surprisingly the node's one.
+This problem - that occasionally shows up - is probably to be attributed to the low samples present in the dataset.
 

### 2025-07-01 doc/article/article.tex
Updated paper

@@
 This positions it as a hybrid between traditional rule-based recognizers and modern neural approaches.
+\\
 
-We'd like to point out a few things about Figure~\ref{fig:results}.
+We'd like to point out a few things about the results shown Figure~\ref{fig:results}.
 First of all, the result shown in part (d).
@@
 This problem - that occasionally shows up - is probably to be attributed to the low samples present in the dataset.
+
+\begin{figure}[h!]
+	\centering
+	\vspace{0.5em}
+
+	% First row
+	\begin{tabular}{ccc}
+		\includegraphics[height=5cm]{ex6.png} &
+		\includegraphics[height=5cm]{chat1.png} &
+		\includegraphics[height=5cm]{chat2.png} \\
+		% Optional captions under each image
+		(a) Our result & (b) GPT Mermaid & (c) GPT D2 \\
+	\end{tabular}
+
+	\vspace{1em}
+
+	% Second row
+	\begin{tabular}{ccc}
+		\includegraphics[height=5cm]{ex4.png} &
+		\includegraphics[height=5cm]{chat3.png} &
+		\includegraphics[height=5cm]{chat4.png} \\
+		(d) Our result & (e) GPT Mermaid & (f) GPT D2 \\
+	\end{tabular}
+
+	\caption{Six vertically-oriented images arranged in two rows}
+	\label{fig:chat}
+
+\end{figure}
+
+Another result we'd like to present is a comparison between D.I.A.G.R.A.M.'s results and what GPT-4o could give in zero-shot.
+The prompt given was the following: "The current image features a diagram.
+Analyze it and output the respective D2 and Mermaid script.".
+The results given (see Figure~\ref{fig:chat}) may be - at first glance - quite similar.
+However, some key differences arise.
+For instance, in diagrams (b), (e), (f), GPT isn't able to correctly output the shape of the nodes - whereas our net is capable of doing.
+Just like our net added a node instead of the self-arrow (see preceding paragraph), GPT isn't able to correctly recognize that portion of ink as a self-arrow (images (e), (f)).
+Another subtle difference - this time, not logical - can be seen in image (c): GPT created three different nodes, while the original image had only one.
 

### 2025-07-08 doc/article/article.tex
Partial update

@@
 Other studies focus on the so-called \textit{online recognition} task, whose goal is to recognize diagrams as they are drawn (see \cite{online1}, \cite{online2}, \cite{online3} and especially \cite{framework} for a general framework).
-The online recognition task quite often employs the use of relative strokes, and is mainly focused on flowchart diagrams.
+The online recognition task quite often employs the use of relative strokes - making it actually \textit{easier} than the offline one - and is mainly focused on flowchart diagrams.
 
@@
 This module is able to apply object detection and semantic recognition to represent the diagram in a unified way.
-After that, the representation is sent to the Transducer, which converts the data in a Markup language of choice.
-The Markup language content is therefore sent to the Compiler, thus generating the high-quality and editable diagram in .png format.
+After that, the representation is processed by the Transducer, which converts the data in a Markup language of choice.
+The Markup language content is therefore sent to the Compiler, thus generating the high-quality and editable diagram in PNG format.
 
@@
 The Classifier is the first component of the pipeline.
-It is able to tell to which category a given image of an handwritten diagram belongs to.
+It is able to understand which category a given image of an handwritten diagram belongs to.
 This is necessary in order to later know the Extractors that can process the input image.
@@
 Every element has a \textit{category} and possibly some \textit{inner} and \textit{outer text}.
-Every relation has a \textit{category} too, either a \textit{source} or \textit{target element} or both, and possibly \textit{text} related to it.
+Every relation has a \textit{category} too, and could have a \textit{source} or \textit{target element}, and possibly \textit{text} related to it.
+The latter property is actually subdivided in four distinct boxes: the text could be associated with the arrow's head, tail, body or the text could be cut off by the arrow itself.
 
@@
 So, every bounding box containing text is linked to the nearest element - it being either a relation or an element.
-The associations that require too long a distance are eliminated further in the algorithm, and the relevant text is discarded.\\
+The associations that require too long a distance are eliminated further in the algorithm, and the associated text is discarded.\\
 
@@
 \label{exp:classifier}
-The first architecture explored for the Classifier module was a Convolutional Neural Network (CNN) following AlexNet's architecture, but smaller in size.
-It had two Convolutional layers, each followed by a batch normalization and then by a max pooling layers. The first layer had only 4 filters, while the second one had 8.
+The first architecture explored for the Classifier module was a Convolutional Neural Network (CNN) following AlexNet's architecture \cite{alexnet}, but smaller in size.
+It has two Convolutional layers, each followed by a batch normalization and then by a max pooling layers. The first layer had only 4 filters, while the second one had 8.
 The output of the second layer was then flattened and passed through a fully connected layer which downsized the features to 128, followed by a ReLU activation function.
@@
 The model was trained on our own dataset, with only three possible output classes: 'flowchart', 'graph' and 'other'; it was trained for 10 epochs, with an AdamW optimizer (with a learning rate of $0.001$) and a standard CrossEntropy Loss function.
+For a more comprehensive overview of the classifier's architecture, see Appendix~\ref{classification_net}.
 \\
@@
 The model employed was tested in three different ways: as zero-shot, after a 10-epoch training and after a 100-epoch training.
-The best-behaving one was, of course, the latter (to see the results of the different stages, see \textbf{Figures}~\ref{fig:bbox1} -~\ref{fig:bbox2} -~\ref{fig:bbox3} -~\ref{fig:fc}).
+The best-behaving one was, of course, the latter (to see the results of the different stages, see \textbf{Figures}~\ref{fig:bbox1} -~\ref{fig:bbox3} -~\ref{fig:fc}).
 In Table~\ref{tab:map} and Table~\ref{tab:mar} we show the Mean Average Precision and Recall of the model used.
@@
 	\centering
-	\includegraphics[width=200pt, height=150pt]{bbox2.jpg}
-	\caption{The model employed after 10 epochs of training.}
-	\label{fig:bbox2}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
 	\includegraphics[width=180pt, height=150pt]{bbox3.png}
-	\caption{The model employed after 100 epochs of training. More bounding boxes can be seen because we also added arrow's tails and heads.}
+	\caption{The model employed after 100 epochs of training. With respect to Figure~\ref{fig:bbox1}, we also added here the heads and the tails for the arrows.}
 	\label{fig:bbox3}
@@
 \begin{enumerate}
+	\item Find the keypoints of the arrow using SIFT \cite{SIFT}.
 	\item Apply a clustering algorithm to the bounding box.
 	\item Suppose that the biggest cluster belongs to the arrow; the other clusters are thought to contain useless elements for this task, and can be discarded.
-	\item Find the keypoints of the arrow using SIFT \cite{SIFT}.
 	\item Find the two keypoints inside the arrow with maximum distance. Assume that these two points are closely related to the two ends of the arrow.
@@
 
+Each convolutional layer, is followed by a ReLU activation function, a layer which applies batch normalization and by a MaxPooling layer.
+
 \end{appendices}

### 2025-07-09 doc/article/article.tex
UPDATE
Conclusion

@@
 
-\title{D.I.A.G.R.A.M.: Development of Image Analysis for Graph Recognition And Modeling}
+\title{D.I.A.G.R.A.M.: Digital Image Analysis for Graph Recognition And Modeling}
 
@@
 \section{Conclusion and Future Work}
-TODO
+In this work, we introduced D.I.A.G.R.A.M., a system designed for the recognition and digitalization of visual diagrams.
+By combining deterministic algorithms with deep learning models, the system is capable of interpreting graphical input and converting it into structured, machine-readable formats.
+Our architecture demonstrates the feasibility of diagram understanding as a hybrid problem, leveraging both rule-based and learned components to achieve robustness and flexibility.
+
+Despite the promising results, several opportunities remain for improvement.
+A key direction for future work is the parallelization of the extractor module, which currently represents a bottleneck in the processing pipeline.
+Efficient multi-threaded implementations could significantly speed up large-scale or real-time diagram analysis.
+
+Additionally, to enhance portability and reproducibility, the system could benefit from containerization using Docker, particularly for managing CLI tools and Python dependencies.
+This would allow for more consistent deployment across diverse environments and simplify collaboration and maintenance.
+
+Continued refinements in both the recognition pipeline and the system's usability will move us closer to a practical tool for academic, industrial, and educational use cases where diagram understanding and easy modeling is essential.
 

### 2025-07-09 doc/article/article.tex
Push

@@
 After that a text box has been assigned to an element, there's the need to understand if the text is either inside or outside the element and whether the text is actually referring to the given element or not.
-In order to do this, the overlap between the two bounding boxes is computed; if the overlap is over a certain threshold, heuristically decided, the association is kept.
-Otherwise, the distance between the two bounding boxes is computed - as the distance between their central point.
+In order to do this, the distance between the two bounding boxes is computed - as the distance between their central point.
 If the distance is over a certain threshold, the association is kept; otherwise, it's discarded.\\
@@
 The overlap between the text bounding box and each of the three sections boxes is computed - then, the section with the highest overlap value is then assigned to the text under consideration.
+\\
 
@@
 	\item Find the keypoints of the arrow using SIFT \cite{SIFT}.
-	\item Apply a clustering algorithm to the bounding box.
+	\item Apply a clustering algorithm to the image inside the bounding box, leveraging only the keypoints.
 	\item Suppose that the biggest cluster belongs to the arrow; the other clusters are thought to contain useless elements for this task, and can be discarded.

### 2025-07-09 doc/article/article.tex
Push

@@
 
+\subsubsection{Arrow Retrieval}
+Sometimes, the number of arrow's bodies, heads and tails don't match up. In particular, the following, trivial, equivalance doesn't hold true:
+$$
+	2\times\text{arrow bodies} = \text{arrow heads} + \text{arrow tails}
+$$
+
+The case in which, for a specific arrow, its head or its tail has not been found.
+In order to be able to be as error-robust as possible, we give the arrow - and a crop with more area around it - to the extraction net once again.
+If it's able to find both an head and a tail this time, we assume those to be the one's linked to the arrow.
+If more than one head or tail is found, the highest confidence result is kept - unless the confidence is lower than a certain threshold.
+
+The other case is the one in which we have some heads and tails without any body.
+For each pair of head and tail, the area of the image linking them is cropped and given to the detection net.
+For every arrow's bounding box found - if any - we compute the overlap with both the head and tail's bounding boxes.
+If the overlap is lower than a certain threshold, the arrow's bounding box is discarded.
+If it's over it and if it's over the current arrow's confidence value - if present, in the case in which we found multiple valid candidates - it's momentarily assumed as the correct arrow.
+If no arrow fulfills any of the above requisite, we discard the couple as candidates.
+\\
+
+\begin{figure}[htbp]
+	\centering
+
+	% Row 1
+	\begin{subfigure}[b]{0.45\linewidth}
+		\includegraphics[width=\linewidth]{retrieval1.jpg}
+		\caption{A graph diagram. The upper rightmost self-arrow has been detected - but not its tail and head.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.45\linewidth}
+		\includegraphics[width=\linewidth]{retrieval2.png}
+		\caption{The detection net's output, able to recognize the arrow's head and tail after passing only a crop of the region.}
+	\end{subfigure}
+
+\end{figure}
+
 \subsubsection{Arrow Orientation}
@@
 It can be assumed that every piece of text is linked to one and only one element or relation.
-So, every bounding box containing text is linked to the nearest element - it being either a relation or an element.
+So, every bounding box containing text is linked to the nearest element - it being either a relation or a node.
 The associations that require too long a distance are eliminated further in the algorithm, and the associated text is discarded.\\
 
-\subsubsection{Element-Text Association}
-After that a text box has been assigned to an element, there's the need to understand if the text is either inside or outside the element and whether the text is actually referring to the given element or not.
-In order to do this, the distance between the two bounding boxes is computed - as the distance between their central point.
-If the distance is over a certain threshold, the association is kept; otherwise, it's discarded.\\
+\subsubsection{Node-Text Association}
+After that a text box has been assigned to a node, there's the need to understand if the text is either inside or outside of the target node and whether the text is actually referring to the given element or not.
+In order to do this, the overlap between the two bounding boxes is computed; the value is compared to an heuristically determined threshold.
+If it's higher, the association is kept and the text is determined to be inside.
+If it's lower, the distance between the two bounding boxes is computed - as the distance between their central point.
+If the distance is below a certain threshold, the association is kept and the text is regarded as outside the node; otherwise, it's discarded.\\
 
-\subsubsection{Relation-Text Association}
-After a text box is assigned to a relation, some computation is needed to fully understand how the text is associated to the relation itself.
-The relation's bounding box is then split into three different sections.
-The overlap between the text bounding box and each of the three sections boxes is computed - then, the section with the highest overlap value is then assigned to the text under consideration.
+\subsubsection{Arrow-Text Association}
+After a text box is assigned to an arrow, some computation is needed to fully understand how the text is associated to the arrow itself.
+The arrow is approximated by a line adjoining its head and its tail.
+The distance between the line and the text's bounding box is computed and compared with a threshold - if it's over it, the association is discarded.
+Otherwise, the line approximation is then split into three different sections - whose sizes are heuristically determined, such that the relation's central section is bigger than the other two.
+The distance between the text bounding box and each of the three sections boxes is computed - then, the section with the lowest distance value is then assigned to the text under consideration.
 \\
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{arrowsplit.jpg}
+	\caption{A three-split of an arrow.}
+	\label{fig:arrowsplit}
+\end{figure}
 
@@
 This targeted approach allows the extraction models to achieve high precision on the specific structural elements relevant to these diagram types, which were the core of our task.
+\\
 
@@
 
-Some examples of the different models applied to our flowchart diagram dataset can be seen in Figure~\ref{fig:text_digitization_results_1} and in Figure~\ref{fig:text_digitization_results_2}.
+Some examples of the different models applied to our flowchart diagram dataset can be seen in Figure~\ref{fig:text_digitization_results_1} and in Figure~\ref{fig:text_digitization_results_2}. \\
 
@@
 However, this approach didn't work as expected (see Figure~\ref{fig:template_matching_fail}).
-In particular, it never worked around self arrows: the maximized distance inside one of the two clusters is never between the tail and the head of the arrow.
+In particular, it never worked around self arrows: the maximized distance inside one of the two clusters is never between the tail and the head of the arrow (see Figure~\ref{fig:selfarrowdc}).
 
@@
 	\label{fig:template_matching_fail}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=150pt]{selfarrowdc.jpg}
+	\caption{An instance which explains why the self-arrow's tail and head are not correctly recognized.}
+	\label{fig:selfarrowdc}
 \end{figure}

### 2025-07-09 doc/article/article.tex
push

@@
 
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=225pt, height=75pt]{arrownet2.jpg}
-	\caption{The overall result of an ArrowNet input.}
-	\label{fig:arrownet2}
-\end{figure}
-
-As seen in Figure~\ref{fig:arrownet2}, due to the low number of samples in the dataset, the network is unable to correctly distinguish between an arrow's head and tail.
+Due to the low number of synthetic samples provided, the network is unable to correctly distinguish between an arrow's head and tail.
+In particular, the network overfitted and simply detected as correct the presence of black ink.
 

### 2025-07-09 doc/article/article.tex
Push
Added self-arrows

@@
 The distance between the text bounding box and each of the three sections boxes is computed - then, the section with the lowest distance value is then assigned to the text under consideration.
-\\
+We must notice that for self-arrow we can not compute simple distance between the line and the text's bounding box, given that the approximated arrow line is not accurate.
+Thus, we approximate the arrow using two segments like a triangle.
+The third point is the middle of the opposite edge. \\
 
@@
 	\caption{A three-split of an arrow.}
+	\label{fig:arrowsplit}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{self_arrow.jpeg}
+	\caption{Self-arrow approximation.}
 	\label{fig:arrowsplit}
@@
 
-Due to the low number of synthetic samples provided, the network is unable to correctly distinguish between an arrow's head and tail.
+Due to the low quality of synthetic samples provided, the network is unable to correctly distinguish between an arrow's head and tail.
 In particular, the network overfitted and simply detected as correct the presence of black ink.

### 2025-07-09 doc/article/article.tex
Updates

@@
 The distance between the line and the text's bounding box is computed and compared with a threshold - if it's over it, the association is discarded.
-Otherwise, the line approximation is then split into three different sections - whose sizes are heuristically determined, such that the relation's central section is bigger than the other two.
+Otherwise, the line approximation is then split into three different sections (see Figure~\ref{fig:arrowsplit}) - whose sizes are heuristically determined, such that the relation's central section is bigger than the other two.
 The distance between the text bounding box and each of the three sections boxes is computed - then, the section with the lowest distance value is then assigned to the text under consideration.
@@
 Thus, we approximate the arrow using two segments like a triangle.
-The third point is the middle of the opposite edge. \\
+The third point is the middle of the opposite edge (see Figure~\ref{fig:arrowapprox}). \\
 
-\begin{figure}
+\begin{figure}[H]
 	\centering
@@
 	\caption{Self-arrow approximation.}
-	\label{fig:arrowsplit}
+	\label{fig:arrowapprox}
 \end{figure}
@@
 		\includegraphics[height=5cm]{chat1.png} &
-		\includegraphics[height=5cm]{chat2.png} \\
+		\includegraphics[height=4.35cm]{chat2.png} \\
 		% Optional captions under each image

### 2025-07-09 doc/article/article.tex
Fix

--- 2025-07-09 doc/article/article.tex
@@
 It can be assumed that every piece of text is linked to one and only one element or relation.
-So, every bounding box containing text is linked to the nearest element - it being either a relation or a node.
-The associations that require too long a distance are eliminated further in the algorithm, and the associated text is discarded.\\
+Thus, every bounding box containing text is linked to the nearest element - it being either a relation or a node - based on their distance.
+The associations that require too long a distance are eliminated further in the algorithm, and the associated text is discarded.
+We must notice that for self-arrow we can not compute simple distance between the line and the text's bounding box, given that the approximated arrow line is not accurate.
+Thus, we approximate the arrow using two segments like a triangle.
+The third point is the middle of the opposite edge. \\
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{self_arrow.jpeg}
+	\caption{Self-arrow approximation.}
+	\label{fig:arrowsplit}
+\end{figure}
 
@@
 The distance between the text bounding box and each of the three sections boxes is computed - then, the section with the lowest distance value is then assigned to the text under consideration.
-We must notice that for self-arrow we can not compute simple distance between the line and the text's bounding box, given that the approximated arrow line is not accurate.
-Thus, we approximate the arrow using two segments like a triangle.
-The third point is the middle of the opposite edge. \\
+\\
 
@@
 	\caption{A three-split of an arrow.}
-	\label{fig:arrowsplit}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{self_arrow.jpeg}
-	\caption{Self-arrow approximation.}
 	\label{fig:arrowsplit}

### 2025-07-09 src/extractor/flowchart/gnr_extractor.py

@@ -310,9 +310,7 @@ class GNRFlowchartExtractor(MultistageFlowchartExtractor):
         return outcome
 
     @override
-    def _manage_wrong_computed_arrows(self, diagram_id: str, image: Image, arrow_bboxes: List[ImageBoundingBox]) -> \
-    List[Optional[Arrow]]:
-
+    def _manage_wrong_computed_arrows(self, diagram_id: str, image: Image, arrow_bboxes: List[ImageBoundingBox]) -> List[Optional[Arrow]]:
         """
         Args:
             diagram_id:

### 2025-07-09 src/extractor/flowchart/multistage_extractor.py

@@ -42,6 +42,7 @@ class MultistageFlowchartExtractor(MultiStageExtractor, ABC):
 
     parallelization: bool = False
 
+    @override
     def _build_diagram_representation(self, diagram_id: str, image: Image, bboxes: List[ImageBoundingBox]) -> FlowchartRepresentation:
         if self.parallelization:
             return self.__par_build_diagram_representation(diagram_id, image, bboxes)

### 2025-07-09 doc/article/article.tex

@@
 The distance between the line and the text's bounding box is computed and compared with a threshold - if it's over it, the association is discarded.
-Otherwise, the line approximation is then split into three different sections - whose sizes are heuristically determined, such that the relation's central section is bigger than the other two.
+Otherwise, the line approximation is then split into three different sections (see Figure~\ref{fig:arrowsplit}) - whose sizes are heuristically determined, such that the relation's central section is bigger than the other two.
 The distance between the text bounding box and each of the three sections boxes is computed - then, the section with the lowest distance value is then assigned to the text under consideration.
-\\
 
-\begin{figure}
+\begin{figure}[H]
 	\centering
